### 2017-04-06 tables/predictedFertility.tex
Further edits of theory part; moved tables out of doc; moved comparison back to appendix

\begin{table}[htbp]
\begin{center}
\begin{scriptsize}
\begin{threeparttable}
\caption{Predicted Fertility Behavior, Sex Ratio, and \protect\linebreak
Abortions for Women with 8 or More Years of Education}
\label{tab:predicted}
\begin{tabular} {@{} l D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1}  D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1} D{.}{.}{2.1}  @{}} \toprule
                 & \multicolumn{4}{c}{1972-1984}                                                               & \multicolumn{4}{c}{1985-1994}                                                               & \multicolumn{4}{c}{1995-2006} \\ \cmidrule(lr){2-5} \cmidrule(lr){6-9} \cmidrule(lr){10-13}
                 & \mco{Births (\%)}          & \mct{Abortions per 100}                      & \mco{Boys}      & \mco{Births (\%)}          & \mct{Abortions per 100}                      & \mco{Boys}      & \mco{Births (\%)}          & \mct{Abortions per 100}                      & \mco{Boys}      \\
                 & \mco{/ fertility\tnote{a}} & \mco{births\tnote{b}} & \mco{women\tnote{c}} & \mco{(\%)}      & \mco{/ fertility\tnote{a}} & \mco{births\tnote{b}} & \mco{women\tnote{c}} & \mco{(\%)}      & \mco{/ fertility\tnote{a}} & \mco{births\tnote{b}} & \mco{women\tnote{c}} & \mco{(\%)}      \\
\midrule
\bf Urban        & \multicolumn{12}{c}{} \\                                                                    
Spell 2          &       85.9   &  2.0    &  1.8    &  52.3   &       78.2   &  4.7    &  3.7  &  53.7   &       68.0   &  8.0    &  5.4   &  55.3      \\
Spell 3          &       64.4   &  5.1    &  3.3    &  53.9   &       49.9   &  9.8    &  4.9  &  56.2   &       38.6   &  7.8    &  3.0   &  55.3      \\
Spell 4          &       57.0   &  2.8    &  1.6    &  52.6   &       44.8   &  2.6    &  1.1  &  52.6   &       33.7   &  21.8   &  7.3   &  62.3      \\
Overall\tnote{d} &       2.7    &  2.0    &  5.5    &  52.3   &       2.3    &  3.4    &  8.0  &  53.1   &       2.0    &  4.6    &  9.4   &  53.7      \\
\addlinespace 
\bf Rural        & \multicolumn{12}{c}{} \\                                                                                  
Spell 2          &       92.5   &  -1.4   &  -1.3   &  50.5   &       87.6   &  3.1    &  2.7  &  52.8   &       83.1   &  4.4    &  3.6   &  53.5      \\
Spell 3          &       80.9   &  2.6    &  2.1    &  52.5   &       70.6   &  6.5    &  4.6  &  54.5   &       59.6   &  10.6   &  6.3   &  56.7      \\
Spell 4          &       75.7   &  -3.1   &  -2.4   &  49.6   &       62.0   &  0.2    &  0.1  &  51.3   &       51.9   &  6.1    &  3.2   &  54.4      \\
Overall\tnote{d} &       3.2    &  -0.3   &  -1.1   &  51.2   &       2.9    &  2.4    &  6.8  &  52.8   &       2.6    &  4.1    &  10.5  &  53.4      \\
\bottomrule
\end{tabular}                        
\begin{tablenotes} \tiny
\item \hspace*{-0.5em} \textbf{Note.} Predictions are based on estimates of equations 
(\ref{eq:probability_boy}) and (\ref{eq:probability_girl}).
The samples consist of all women from the data with 8 or more years of education who
married during the indicated period using the estimation results for that periods 
and the women's characteristics at the start of their marriage. 
All numbers presented are for women predicted to have one or more births.
Birth/progression rates are based on the ratio of women predicted to have a birth 
by the end of each spell.
Abortion rates within each spell are based on the predicted number of abortions each
woman would have if she went through the entire spell.
To generate conditions such as sex composition and fecundity for second and higher spells 
each woman is randomly assigned outcomes based on the estimation results and
her characteristics as follows.
For each quarter, her probabilities of having a boy, a girl, or no birth is calculated.
She is assigned one of these outcomes based on a random draw from a uniform distribution.
The first quarter where she has a draw that results in a birth she leaves the spell
and the quarter reached is used to calculate her starting age for the subsequent spell.
If she has no draw that result in a birth within a spell her childbearing stops and
she does not enter the next spell.
Results are averaged over 1,000 repetitions.
\item[a] For individual spells the column shows the percent of women who started the
spell and ended up with a birth by the end of the spell period covered.
For the ``Overall'' row, the number is the expected number of children born per woman over 
the first four spells, conditional on having at least one birth.
\item[b] For individual spells the column shows the number of abortions per 100
births predicted to take place by the end of the spell.
For the ``Overall'' row, the number is the total number of abortion per 100 births 
predicted to take place over all four spells.
\item[c]  For individual spells the column shows the number of abortions 
predicted to take place by the end of the spell per 100 women who started the spell .
For the ``Overall'' row, the number is the total number of abortion 
predicted to take place over all four spells per 100 women who had at least one child.
\item[d] Covers first through fourth spells. 
\end{tablenotes}
\end{threeparttable}
\end{scriptsize}
\end{center}
\end{table}