### 2017-10-13 paper/sexSelectionSpacingAppendix-ver1.tex
Initial edits of Appendix for this version

% Begun.: 2017-04-02
% Edited: 2017-04-02

\documentclass[12pt,letterpaper]{article}
% \documentclass[AER,draftmode]{AEA}

\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}
\usepackage[title]{appendix}
\usepackage[margin=1.0in]{geometry}
\usepackage[figuresleft]{rotating}
\usepackage[longnamesfirst]{natbib}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}
\usepackage{setspace}
\usepackage[justification=centering]{caption}
\usepackage[font=scriptsize]{subfig}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{adjustbox}
% \usepackage[nolists,tablesfirst]{endfloat}

% \bibpunct{(}{)}{;}{a}{}{,}
\newcommand{\mco}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\mct}[1]{\multicolumn{2}{c}{#1}}
\newcommand{\X}{$\times$ }
\newcommand{\hs}{\hspace{15pt}}

% Attempt to squeeze more floats in
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

% \renewcommand\subfigtopskip{0pt}
% \renewcommand\subfigcapskip{0pt}
%\renewcommand\subfigbottomskip{2pt}

\newcommand{\bkv}{\Bkk{5.2,0}{0,0}{d}}
\newcommand{\bkh}{\Bkk{5.2,0}{0,0}{dr}}
\newcommand{\bkl}{\Bkk{5.2,0}{5.2,0}{dl}}
\newcommand{\bkd}{\Bkk{5.2,0}{5.2,0}{dd}}
\newcommand{\bkr}{\Bkk{5.2,0}{5.2,0}{dr}}

%
% If tree is not showing remove divps in line 60 in the following file
% /Library/TeX/Root/texmf-dist/tex/latex/xyling/xyling.sty
%

% For estout tables
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}

%------------------------------------------------------------------------

\title{Sex Selective Abortions, Fertility, and Birth Spacing
% \protect\thanks{%
% I am grateful to Andrew Foster and Darryl Holman for discussions about the method.
% I owe thanks to Shelly Lundberg, Daniel Rees, David Ribar, Hendrik Wolff, 
% seminar participants at University of Copenhagen, University of Michigan, 
% University of Washington, University of \AA{}rhus, the Fourth 
% Annual Conference on Population, Reproductive Health, 
% and Economic Development, and the Economic Demography Workshop for helpful suggestions and comments.
% I would also like to thank Nalina Varanasi for research assistance.
% Support from the University of Washington Royalty Research Fund and the 
% Development Research Group of the World Bank is gratefully acknowledged.
% The views and findings expressed here are those of the author and
% should not be attributed to the World Bank or any of its member countries.
% Partial support for this research came from a Eunice Kennedy Shriver National
% Institute of Child Health and Human Development research infrastructure grant,
% 5R24HD042828, to the Center for Studies in Demography and Ecology at the
% University of Washington.
% Prior versions of this paper were presented under the title ``The 
% Determinants of Sex Selective Abortions.''
% }
}

\author{}

% \author{Claus C P\"ortner\\
%     Department of Economics\\
%     Albers School of Business and Economics\\
%     Seattle University, P.O. Box 222000\\
%     Seattle, WA 98122\\
%     \href{mailto:cportner@seattleu.edu}{\texttt{cportner@seattleu.edu}}\\
%     \href{http://www.portner.dk}{\texttt{www.portner.dk}}\\
%     \& \\
%     Center for Studies in Demography and Ecology \\
%     University of Washington\\ \vspace{2cm}
%     }

% \date{May 2010\\ \bigskip PRELIMINARY}
\date{October 2017}

%------------------------------------------------------------------------

\begin{document}

\graphicspath{{../figures/}}
\DeclareGraphicsExtensions{.eps,.jpg,.pdf,.mps,.png}
\doublespacing

\appendix

% \section{Recall Error}

% CHANGING NUMBERING OF FIGURES AND TABLES FOR APPENDIX
\renewcommand\thefigure{\thesection.\arabic{figure}}    
% \setcounter{figure}{0}    
% \setcounter{postfig}{0}

\renewcommand\thetable{\thesection.\arabic{table}}    
\setcounter{table}{0}
% \setcounter{posttbl}{0}

% \section{Why Bundle Scan and Abortion?}
% 
% The easiest way to show why a bundled service makes sense is to model 
% the provider's problem first and then show that both provider and
% buyer will be better off if a bundled service is offered.
% Take a provider with with cost, $c_a$, of providing an abortion,
% and cost, $c_s$, of providing a scan.
% These costs include expenses for equipment, labor,
% and possibly a premium for getting caught and prosecuted. 
% The provider can either charge separately for the two services,
% in which case he will charge $p_s$ for the scan and $p_a$ for
% the abortion, or he can charge $p_b$ for the bundle.
% 
% The provider's profit if he offers the two services separately, $\pi_{sep}$,
% to $n$ buyers, who conceive girls with the probability $p(g)$, will be
% \begin{equation}
% \pi_{sep} = n (p_s - c_s) + p(g) n (p_a - c_a).
% \end{equation}
% The first part is profit from the customers who want a scan
% and the second part is the profit from the fraction of the
% original customers who conceived a girl.
% If the provider instead offered the same $n$ buyers a bundle service
% the profit, $\pi_{bun}$, would be
% \begin{equation}
% \pi_{bun} = n p_b - n c_s - p(g) n c_a,
% \end{equation}
% since all customers pays $p_b$, all customers get a scan 
% at cost $c_s$, and the fraction $p(g)$ who conceived a girl
% would receive an abortion at cost $c_a$.
% A profit maximizing provider then picks bundled pricing if 
% $\pi_{bun} > \pi_{sep}$, or
% \begin{equation}
%  p_b > p_s + p(g) p_a.
% \end{equation}
% 
% Assume that buyers are risk averse and that the only motivation
% for getting a prenatal sex determination test is to abort 
% female fetuses and that parents will abort the fetus if the test 
% show a female fetus (I will relax the last assumptions below).
% Furthermore, to ease exposition assume that buyers can only
% spend money on two things: consumption $c$ and prenatal sex
% determinations/abortions.
% For a buyer facing separate services the expected utility
% is then
% \begin{equation}
% U = (1-p(g)) U(c-p_s) + p(g) U(c - p_s - p_a),
% \end{equation}
% where the first part is the utility if the fetus is a boy
% and the buyers therefore only spend on the scan, which
% happens with probability $1-p(g)$, and the
% second part is the utility if the fetus is a female
% and they therefore also has to pay for the abortion,
% which happens with probability $p(g)$.
% For a buyer facing a bundled service the utility is simply
% \begin{equation}
% U = U(c-p_b).
% \end{equation}
% 
% If follows from the standard assumptions about utility functions
% and the risk aversion assumption that
% \begin{equation}
% U(c - p_s - p(g) p_a) > (1-p(g)) U(c-p_s) + p(g) U(c - p_s - p_a).
% \end{equation}
% In other words risk averse buyers prefer a sure thing---where they pay 
% upfront---over the gamble that they either have a boy and pay only for the
% scanning or have a girl and pay for both the scan and the abortion.
% In fact, risk aversion implies that they are willing to pay 
% \emph{more} than the expected cost.
% In other words,
% \begin{equation}
% U(c - p_s - p(g) p_a - \delta) = (1 - p(g)) U(c-p_s) + p(g) U(c - p_s - p_a),
% \end{equation}
% where $\delta$ is a positive number that depends on the
% degree of risk aversion (the more risk averse the larger
% $\delta$ is).
% 
% The provider realizes that buyers are risk averse
% and therefore offer a bundle where the price is in 
% the range
% \begin{equation}
% p_b \in [ p_s + p(g) p_a; p_s + p(g) p_a + \delta ].
% \end{equation}
% Hence, the optimal pricing scheme for both provider
% and buyer is bundled pricing, where you pay a flat fee
% and get the pre-natal sex determination and an abortion
% if needed.
% % Where the provider ends up in the possible range of
% % prices depends on the nature of the provider market.
% 
% I implicitly assumed that the provider is risk-neutral.
% The same result hold, however, if the provider
% is risk-averse as long as there are enough potential clients.
% In that case, the law of large numbers will ensure that 
% the risk-neutral and risk-averse choices are the same
% for the provider---essentially there is no gamble if you see 
% enough clients, $p(g)$ of them will present with a female fetus 
% and $1-p(g)$ with a male fetus.
% In fact, it is possible for the provider to be \emph{more}
% risk averse than the buyers and still prefer the bundled 
% pricing scheme.
% 
% Now imagine a situation where there there is some 
% uncertainty about whether a buyer will go through with
% the abortion after the prenatal sex determination.
% This can be represented by the probability $\rho$ that
% a buyer who had a prenatal sex determination test that
% showed a female fetus will go through with the abortion.
% If parents know that there is probability $1-\rho$ that they
% will not go through with the abortion their utility is
% \begin{equation}
% U = (1-p(g)) U(c-p_s) + p(g) U(c - p_s - \rho p_a),
% \end{equation}
% where I have made the assumption that they
% are equally happy with a boy or a girl if they
% are in the $1-\rho$ portion of the population that
% have a prenatal sex determination which show a female
% fetus but do not go through with the abortion (obviously
% this cannot be the case because if they were there
% would be no point in having the prenatal sex determination,
% but if I can show that the bundle price arise here it
% would trivially hold if there were some lower utility
% associated with having a girl than a boy).
% 
% Because the buyers are still risk averse they would continue
% to prefer the ``sure'' thing over the gamble
% \begin{equation}
% U(c - p_s - p(g) \rho p_a) > p(b) U(c-p_s) + p(g) U(c - p_s - \rho p_a).
% \end{equation}
% The provider still offers a bundle, but the price is now 
% in  the range
% \begin{equation}
% p_b \in [ p_s + p(g) \rho p_a; p_s + p(g) \rho p_a + \delta ],
% \end{equation}
% where $\delta$ still depends on the degree of risk aversion.
% 
% If parents on the other hand initially believe that they will always
% go through with the abortion if the fetus is female but
% still have probability $1-\rho$ that they will not, they would
% be back to the original comparison of the utility.
% \begin{equation}
% U(c - p_s - p(g) p_a) > p(b) U(c-p_s) + p(g) U(c - p_s - p_a).
% \end{equation}
% All that would do is \emph{increase} the profit for the
% provider because the law of large numbers would tell them
% exactly how many would not go through with the abortion.
% In other words, they would offer the original
% bundled price because the buyers did not know that they
% will change their mind
% \begin{equation}
% p_b \in [ p_s + p(g) p_a; p_s + p(g) p_a + \delta ],
% \end{equation}
% but the provider will have profit of
% \begin{equation}
% p_b - c_s - \rho c_a,
% \end{equation}
% which is strictly larger than the profit in either of
% the other cases.
% Also, they do not need to know which buyers are more
% or less likely to use abortion, just that some
% fraction will not use their purchased abortion.
% 
% \clearpage
% \newpage
%

% \section{Simulation Results for 20 Periods}
% 
% \begin{table}[!htp]
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Fertility and Percent Boys by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_fertility_100}
% \begin{tabular}{@{} c D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} @{}}
% \toprule
%                & \multicolumn{10}{c}{Discount rate ( $ d = 1.0 $ ) }                         \\ 
%                & \multicolumn{10}{c}{Cost of prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{2}{c}{12} & \multicolumn{2}{c}{16} & \multicolumn{2}{c}{20} & \multicolumn{2}{c}{24} & \multicolumn{2}{c}{200} \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9} \cmidrule(lr){10-11}
% \multirow{2}{12mm}{Cost of\\children}  & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       \\
%                                        & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} \\
% \midrule
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &  7.46 &  50.4 &  7.46 &  50.1 &  7.47 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 \\
%  k = 14        &  5.40 &  52.0 &  5.40 &  50.5 &  5.40 &  50.3 &  5.41 &  50.2 &  5.43 &  50.0 \\
%  k = 16        &  4.10 &  51.5 &  4.10 &  51.5 &  4.10 &  51.4 &  4.13 &  51.1 &  4.19 &  49.9 \\
%  k = 18        &  3.13 &  54.0 &  3.13 &  54.0 &  3.13 &  54.0 &  3.14 &  53.7 &  3.25 &  50.0 \\
%  k = 20        &  2.26 &  61.0 &  2.38 &  57.9 &  2.50 &  55.1 &  2.50 &  55.1 &  2.50 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &  7.15 &  53.3 &  7.27 &  51.2 &  7.29 &  50.4 &  7.30 &  50.1 &  7.31 &  50.0 \\
%  k = 14        &  5.21 &  55.3 &  5.22 &  54.0 &  5.30 &  52.9 &  5.37 &  51.0 &  5.42 &  50.0 \\
%  k = 16        &  4.10 &  59.1 &  4.10 &  58.7 &  4.10 &  56.8 &  4.10 &  52.6 &  4.19 &  49.9 \\
%  k = 18        &  3.13 &  69.0 &  3.13 &  57.5 &  3.13 &  54.0 &  3.13 &  54.0 &  3.32 &  50.0 \\
%  k = 20        &  2.25 &  61.2 &  2.25 &  61.2 &  2.25 &  61.2 &  2.25 &  61.1 &  2.50 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &  7.07 &  61.7 &  7.10 &  57.2 &  7.14 &  54.3 &  7.24 &  52.0 &  7.34 &  50.0 \\
%  k = 14        &  5.20 &  65.2 &  5.05 &  63.3 &  5.05 &  60.6 &  5.06 &  56.3 &  5.26 &  50.0 \\
%  k = 16        &  4.06 &  75.0 &  4.06 &  71.5 &  4.06 &  63.3 &  4.06 &  60.3 &  4.16 &  49.9 \\
%  k = 18        &  3.13 &  70.0 &  3.13 &  69.9 &  3.13 &  69.7 &  3.13 &  69.0 &  3.32 &  50.0 \\
%  k = 20        &  2.98 &  70.7 &  2.81 &  68.9 &  2.25 &  61.2 &  2.25 &  61.2 &  2.63 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.25 &  49.9 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  7.23 &  72.0 &  7.12 &  67.9 &  7.10 &  63.2 &  7.09 &  59.2 &  7.34 &  50.0 \\
%  k = 14        &  5.38 &  80.8 &  5.05 &  77.4 &  5.05 &  72.8 &  5.05 &  67.6 &  5.27 &  50.0 \\
%  k = 16        &  4.06 &  76.1 &  4.06 &  75.9 &  4.06 &  75.4 &  4.06 &  73.2 &  4.17 &  49.9 \\
%  k = 18        &  3.77 &  75.2 &  3.13 &  70.0 &  3.13 &  70.0 &  3.13 &  69.9 &  3.32 &  50.0 \\
%  k = 20        &  2.75 &  72.7 &  2.75 &  72.7 &  2.74 &  72.6 &  2.71 &  72.4 &  2.38 &  50.0 \\
%  k = 22        &  2.00 &  99.8 &  2.00 &  99.1 &  2.00 &  95.1 &  2.00 &  74.1 &  2.25 &  49.9 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a]Percent boys is the percent boys of all births.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% 
% \begin{table}[!htp]
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Fertility and Percent Boys by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_fertility_95}
% \begin{tabular}{@{} c D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} @{}}
% \toprule
%                & \multicolumn{10}{c}{Discount rate ( $ d = 0.95 $ ) }                         \\ 
%                & \multicolumn{10}{c}{Cost of prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{2}{c}{12} & \multicolumn{2}{c}{16} & \multicolumn{2}{c}{20} & \multicolumn{2}{c}{24} & \multicolumn{2}{c}{200} \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9} \cmidrule(lr){10-11}
% \multirow{2}{12mm}{Cost of\\children}  & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       \\
%                                        & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} \\
% \midrule
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &  7.46 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 \\
%  k = 14        &  5.40 &  50.3 &  5.41 &  50.3 &  5.42 &  50.1 &  5.42 &  50.0 &  5.43 &  50.0 \\
%  k = 16        &  4.10 &  51.5 &  4.13 &  51.1 &  4.16 &  50.7 &  4.16 &  50.7 &  4.19 &  49.9 \\
%  k = 18        &  3.13 &  54.0 &  3.14 &  53.7 &  3.25 &  51.9 &  3.25 &  51.9 &  3.25 &  50.0 \\
%  k = 20        &  2.50 &  55.1 &  2.50 &  55.1 &  2.50 &  55.1 &  2.50 &  55.0 &  2.50 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.4 &  2.00 &  61.0 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &  7.30 &  50.8 &  7.30 &  50.3 &  7.30 &  50.0 &  7.30 &  50.0 &  7.31 &  50.0 \\
%  k = 14        &  5.24 &  53.8 &  5.40 &  51.7 &  5.40 &  50.4 &  5.40 &  50.2 &  5.42 &  50.0 \\
%  k = 16        &  4.10 &  58.9 &  4.10 &  52.6 &  4.10 &  51.5 &  4.11 &  51.3 &  4.19 &  49.9 \\
%  k = 18        &  3.13 &  54.0 &  3.13 &  54.0 &  3.13 &  54.0 &  3.13 &  53.9 &  3.32 &  50.0 \\
%  k = 20        &  2.25 &  61.2 &  2.25 &  61.2 &  2.27 &  60.8 &  2.50 &  55.1 &  2.50 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &  7.09 &  56.9 &  7.19 &  52.4 &  7.30 &  50.5 &  7.33 &  50.1 &  7.34 &  50.0 \\
%  k = 14        &  5.05 &  63.7 &  5.05 &  56.4 &  5.10 &  53.6 &  5.21 &  51.0 &  5.26 &  50.0 \\
%  k = 16        &  4.06 &  69.1 &  4.06 &  59.2 &  4.06 &  58.8 &  4.06 &  55.3 &  4.16 &  49.9 \\
%  k = 18        &  3.13 &  70.0 &  3.13 &  69.7 &  3.13 &  64.9 &  3.13 &  56.0 &  3.32 &  50.0 \\
%  k = 20        &  2.50 &  65.1 &  2.25 &  61.2 &  2.25 &  61.2 &  2.25 &  61.2 &  2.63 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.25 &  49.9 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  7.07 &  67.6 &  7.08 &  60.9 &  7.13 &  54.8 &  7.26 &  51.4 &  7.34 &  50.0 \\
%  k = 14        &  5.05 &  78.3 &  5.05 &  67.8 &  5.05 &  62.7 &  5.06 &  56.7 &  5.27 &  50.0 \\
%  k = 16        &  4.06 &  76.0 &  4.06 &  75.4 &  4.06 &  69.7 &  4.06 &  63.3 &  4.17 &  49.9 \\
%  k = 18        &  3.13 &  70.1 &  3.13 &  70.0 &  3.13 &  69.9 &  3.13 &  69.7 &  3.32 &  50.0 \\
%  k = 20        &  2.75 &  72.7 &  2.74 &  72.6 &  2.64 &  71.6 &  2.00 &  62.5 &  2.38 &  50.0 \\
%  k = 22        &  2.00 &  99.1 &  2.00 &  71.8 &  2.00 &  62.5 &  2.00 &  62.5 &  2.25 &  49.9 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a]Percent boys is the percent boys of all births.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% \begin{table}[!htp]
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Fertility and Percent Boys by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_fertility_90}
% \begin{tabular}{@{} c D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} @{}}
% \toprule
%                & \multicolumn{10}{c}{Discount rate ( $ d = 0.9 $ ) }                         \\ 
%                & \multicolumn{10}{c}{Cost of prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{2}{c}{12} & \multicolumn{2}{c}{16} & \multicolumn{2}{c}{20} & \multicolumn{2}{c}{24} & \multicolumn{2}{c}{200} \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9} \cmidrule(lr){10-11}
% \multirow{2}{12mm}{Cost of\\children}  & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       \\
%                                        & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} \\
% \midrule
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &  7.47 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 &  7.47 &  50.0 \\
%  k = 14        &  5.42 &  50.1 &  5.43 &  50.0 &  5.43 &  50.0 &  5.43 &  50.0 &  5.43 &  50.0 \\
%  k = 16        &  4.16 &  50.7 &  4.19 &  50.1 &  4.19 &  49.9 &  4.19 &  49.9 &  4.19 &  49.9 \\
%  k = 18        &  3.25 &  51.9 &  3.25 &  51.4 &  3.25 &  50.0 &  3.25 &  50.0 &  3.25 &  50.0 \\
%  k = 20        &  2.50 &  55.1 &  2.50 &  50.0 &  2.50 &  50.0 &  2.50 &  50.0 &  2.50 &  50.0 \\
%  k = 22        &  2.01 &  62.2 &  2.00 &  50.0 &  2.00 &  50.0 &  2.00 &  50.0 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &  7.30 &  50.0 &  7.31 &  50.0 &  7.31 &  50.0 &  7.31 &  50.0 &  7.31 &  50.0 \\
%  k = 14        &  5.40 &  50.3 &  5.41 &  50.1 &  5.42 &  50.0 &  5.42 &  50.0 &  5.42 &  50.0 \\
%  k = 16        &  4.10 &  51.4 &  4.16 &  50.7 &  4.19 &  50.1 &  4.19 &  49.9 &  4.19 &  49.9 \\
%  k = 18        &  3.13 &  54.0 &  3.25 &  51.9 &  3.26 &  51.9 &  3.32 &  50.0 &  3.32 &  50.0 \\
%  k = 20        &  2.50 &  55.1 &  2.50 &  55.1 &  2.50 &  55.0 &  2.50 &  50.0 &  2.50 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.24 &  55.4 &  2.00 &  50.0 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &  7.30 &  50.5 &  7.33 &  50.0 &  7.34 &  50.0 &  7.34 &  50.0 &  7.34 &  50.0 \\
%  k = 14        &  5.24 &  52.3 &  5.24 &  50.3 &  5.25 &  50.1 &  5.26 &  50.0 &  5.26 &  50.0 \\
%  k = 16        &  4.06 &  58.8 &  4.07 &  51.5 &  4.13 &  50.7 &  4.16 &  50.1 &  4.16 &  49.9 \\
%  k = 18        &  3.13 &  54.0 &  3.13 &  54.0 &  3.25 &  51.9 &  3.26 &  51.9 &  3.32 &  50.0 \\
%  k = 20        &  2.25 &  61.2 &  2.25 &  61.1 &  2.50 &  55.1 &  2.50 &  55.0 &  2.63 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.4 &  2.25 &  49.9 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  7.12 &  53.6 &  7.31 &  50.3 &  7.33 &  50.0 &  7.34 &  50.0 &  7.34 &  50.0 \\
%  k = 14        &  5.05 &  63.2 &  5.21 &  52.6 &  5.24 &  50.3 &  5.26 &  50.0 &  5.27 &  50.0 \\
%  k = 16        &  4.06 &  61.5 &  4.06 &  59.0 &  4.07 &  51.5 &  4.13 &  50.7 &  4.17 &  49.9 \\
%  k = 18        &  3.13 &  70.0 &  3.13 &  68.3 &  3.13 &  54.0 &  3.16 &  53.5 &  3.32 &  50.0 \\
%  k = 20        &  2.43 &  69.2 &  2.00 &  62.5 &  2.00 &  62.5 &  2.25 &  55.6 &  2.38 &  50.0 \\
%  k = 22        &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.00 &  62.5 &  2.25 &  49.9 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a]Percent boys is the percent boys of all births.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% \begin{table}[!htp]
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Fertility and Percent Boys by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_fertility_85}
% \begin{tabular}{@{} c D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} D{.}{.}{2.2} D{.}{.}{1.2} D{.}{.}{1.2} @{}}
% \toprule
%                & \multicolumn{10}{c}{Discount rate ( $ d = 0.85 $ ) }                         \\ 
%                & \multicolumn{10}{c}{Cost of prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{2}{c}{12} & \multicolumn{2}{c}{16} & \multicolumn{2}{c}{20} & \multicolumn{2}{c}{24} & \multicolumn{2}{c}{200} \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9} \cmidrule(lr){10-11}
% \multirow{2}{12mm}{Cost of\\children}  & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       & \mco{Avg}       & \mco{Percent}       \\
%                                        & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} & \mco{Fertility} & \mco{Boys\tnote{a}} \\
% \midrule
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &  7.46 &  50.0 &  7.46 &  50.0 &  7.46 &  50.0 &  7.46 &  50.0 &  7.46 &  50.0 \\
%  k = 14        &  5.42 &  50.1 &  5.42 &  50.0 &  5.42 &  50.0 &  5.42 &  50.0 &  5.42 &  50.0 \\
%  k = 16        &  4.16 &  50.7 &  4.18 &  50.0 &  4.18 &  50.0 &  4.18 &  50.0 &  4.18 &  50.0 \\
%  k = 18        &  3.25 &  51.9 &  3.25 &  50.1 &  3.25 &  50.1 &  3.25 &  50.1 &  3.25 &  50.1 \\
%  k = 20        &  2.50 &  54.9 &  2.50 &  50.0 &  2.50 &  50.0 &  2.50 &  50.0 &  2.50 &  50.0 \\
%  k = 22        &  2.19 &  54.3 &  2.00 &  50.0 &  2.00 &  50.0 &  2.00 &  50.0 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &  7.30 &  50.0 &  7.30 &  50.0 &  7.30 &  50.0 &  7.30 &  50.0 &  7.30 &  50.0 \\
%  k = 14        &  5.40 &  50.2 &  5.41 &  50.0 &  5.41 &  50.0 &  5.41 &  50.0 &  5.41 &  50.0 \\
%  k = 16        &  4.15 &  50.7 &  4.16 &  50.6 &  4.18 &  50.0 &  4.18 &  50.0 &  4.18 &  50.0 \\
%  k = 18        &  3.14 &  53.8 &  3.25 &  51.9 &  3.31 &  50.1 &  3.31 &  50.1 &  3.31 &  50.1 \\
%  k = 20        &  2.50 &  55.0 &  2.50 &  54.9 &  2.50 &  52.5 &  2.50 &  50.0 &  2.50 &  50.0 \\
%  k = 22        &  2.00 &  62.4 &  2.00 &  62.1 &  2.00 &  50.0 &  2.00 &  50.0 &  2.00 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &  7.33 &  50.1 &  7.33 &  50.0 &  7.33 &  50.0 &  7.33 &  50.0 &  7.33 &  50.0 \\
%  k = 14        &  5.23 &  51.3 &  5.25 &  50.2 &  5.25 &  50.0 &  5.25 &  50.0 &  5.25 &  50.0 \\
%  k = 16        &  4.06 &  52.3 &  4.12 &  50.7 &  4.14 &  50.5 &  4.15 &  50.0 &  4.15 &  50.0 \\
%  k = 18        &  3.12 &  54.0 &  3.13 &  53.9 &  3.25 &  51.9 &  3.31 &  50.1 &  3.31 &  50.1 \\
%  k = 20        &  2.25 &  61.0 &  2.31 &  59.4 &  2.50 &  54.9 &  2.51 &  54.3 &  2.62 &  50.0 \\
%  k = 22        &  2.00 &  62.4 &  2.00 &  62.4 &  2.00 &  62.3 &  2.12 &  58.0 &  2.25 &  50.0 \\
%                &  \multicolumn{10}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  7.30 &  50.7 &  7.33 &  50.0 &  7.33 &  50.0 &  7.33 &  50.0 &  7.33 &  50.0 \\
%  k = 14        &  5.06 &  55.0 &  5.23 &  50.6 &  5.25 &  50.1 &  5.26 &  50.0 &  5.26 &  50.0 \\
%  k = 16        &  4.06 &  58.9 &  4.06 &  55.2 &  4.10 &  51.1 &  4.16 &  50.2 &  4.17 &  50.0 \\
%  k = 18        &  3.12 &  69.7 &  3.13 &  57.1 &  3.13 &  53.9 &  3.25 &  51.8 &  3.31 &  50.1 \\
%  k = 20        &  2.00 &  62.4 &  2.00 &  62.4 &  2.01 &  62.1 &  2.25 &  55.4 &  2.37 &  50.0 \\
%  k = 22        &  2.00 &  62.4 &  2.00 &  62.4 &  2.00 &  62.4 &  2.00 &  62.3 &  2.25 &  50.0 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a]Percent boys is the percent boys of all births.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% 
% \begin{table}[!htp]
% \centering
% \scriptsize
% \begin{threeparttable}
% \caption{Abortions and Change in Abortion Use by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_abortion_100}
% \begin{tabular}{@{} c D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} @{}}
% \toprule
%                & \multicolumn{12}{c}{Discount rate ( $ d = 1.0 $ ) }                         \\ 
%                & \multicolumn{12}{c}{Cost of Prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{3}{c}{12}    & \multicolumn{3}{c}{16}    & \multicolumn{3}{c}{20}    & \multicolumn{3}{c}{24}    \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13} 
%                                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       \\
% \multirow{2}{9mm}{Cost of\\children}  & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         \\
%                                       & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} \\
% \midrule
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &    6.0 &    0.1 &   22.3 &    1.4 &    0.0 &   28.2 &    0.6 &    0.0 &   49.1 &    0.2 &    0.0 &   48.2 \\
%  k = 14        &   21.6 &    0.2 &   12.3 &    5.6 &    0.1 &   23.2 &    3.2 &    0.0 &    4.8 &    3.0 &    0.0 &   15.0 \\
%  k = 16        &   12.7 &    0.1 &    0.2 &   12.6 &    0.1 &    1.0 &   12.2 &    0.1 &    6.4 &    9.5 &    0.2 &   40.1 \\
%  k = 18        &   25.3 &    0.3 &    0.1 &   25.3 &    0.3 &    0.5 &   25.1 &    0.3 &    1.8 &   23.7 &    0.3 &   12.0 \\
%  k = 20        &   49.6 &    0.5 &    3.2 &   37.8 &    0.6 &   40.1 &   25.3 &    0.3 &    0.1 &   25.3 &    0.3 &    0.2 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.3 &    0.5 &    0.1 &   50.1 &    0.5 &    0.4 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &   47.0 &    0.6 &   21.7 &   18.1 &    0.3 &   37.6 &    6.5 &    0.1 &   47.3 &    1.3 &    0.0 &   29.1 \\
%  k = 14        &   56.5 &    0.9 &   36.7 &   41.9 &    0.5 &    7.6 &   31.2 &    0.5 &   34.0 &   10.8 &    0.1 &   25.1 \\
%  k = 16        &   75.2 &    0.8 &    0.8 &   72.1 &    0.7 &    4.0 &   56.4 &    0.7 &   19.4 &   22.1 &    0.3 &   22.7 \\
%  k = 18        &  119.3 &    1.2 &    4.1 &   47.3 &    0.5 &   11.9 &   25.3 &    0.3 &    0.1 &   25.3 &    0.3 &    0.3 \\
%  k = 20        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 &   50.3 &    0.5 &    0.2 &   50.2 &    0.5 &    0.9 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &  165.3 &    2.1 &   22.6 &  102.2 &    1.4 &   27.9 &   61.3 &    0.9 &   32.2 &   29.3 &    0.5 &   40.6 \\
%  k = 14        &  158.7 &    1.6 &    1.0 &  135.6 &    1.4 &    4.8 &  108.0 &    1.3 &   15.4 &   64.1 &    0.8 &   19.1 \\
%  k = 16        &  204.5 &    2.1 &    3.2 &  175.9 &    1.9 &   10.7 &  109.5 &    1.3 &   15.8 &   84.6 &    0.9 &    8.6 \\
%  k = 18        &  126.0 &    1.2 &    0.1 &  125.4 &    1.2 &    0.5 &  123.9 &    1.2 &    1.3 &  119.4 &    1.2 &    4.0 \\
%  k = 20        &  123.9 &    1.2 &    0.0 &  107.0 &    1.1 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  319.6 &    3.8 &   15.6 &  256.2 &    3.3 &   21.9 &  188.1 &    2.5 &   26.3 &  131.1 &    1.6 &   17.7 \\
%  k = 14        &  331.2 &    3.4 &    1.5 &  277.3 &    2.9 &    5.4 &  230.2 &    2.6 &   10.7 &  179.1 &    2.0 &   13.1 \\
%  k = 16        &  213.6 &    2.1 &    0.2 &  212.2 &    2.1 &    0.7 &  208.2 &    2.1 &    2.0 &  190.3 &    2.0 &    6.6 \\
%  k = 18        &  191.3 &    1.9 &    0.0 &  126.0 &    1.2 &    0.1 &  125.9 &    1.2 &    0.2 &  125.4 &    1.2 &    0.5 \\
%  k = 20        &  126.0 &    1.2 &    0.0 &  125.7 &    1.2 &    0.0 &  124.9 &    1.2 &    0.0 &  122.3 &    1.2 &    0.0 \\
%  k = 22        &  200.3 &    2.0 &    0.3 &  197.2 &    2.0 &    1.5 &  181.2 &    1.9 &    7.3 &   97.1 &    1.0 &    6.1 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \tiny
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a] Average number of spells per woman where scan has been used at least once.
% \item[b] Percent of spells where prenatal scan was initially used but birth is the result
% of pregnancy without prenatal scan.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% 
% \begin{table}[!htp]
% \centering
% \scriptsize
% \begin{threeparttable}
% \caption{Abortions and Change in Abortion Use by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_abortion_95}
% \begin{tabular}{@{} c D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} @{}}
% \toprule
%                & \multicolumn{12}{c}{Discount rate ( $ d = 0.95 $ ) }                         \\ 
%                & \multicolumn{12}{c}{Cost of Prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{3}{c}{12}    & \multicolumn{3}{c}{16}    & \multicolumn{3}{c}{20}    & \multicolumn{3}{c}{24}    \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13} 
%                                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       \\
% \multirow{2}{9mm}{Cost of\\children}  & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         \\
%                                       & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} \\
% \midrule
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &    0.9 &    0.0 &    8.3 &    0.4 &    0.0 &    7.7 &    0.2 &    0.0 &   48.2 &    0.0 &    0.0 &      . \\
%  k = 14        &    3.3 &    0.0 &    0.8 &    3.2 &    0.0 &    7.2 &    1.6 &    0.0 &    7.4 &    0.8 &    0.0 &   50.3 \\
%  k = 16        &   12.6 &    0.1 &    1.0 &    9.6 &    0.2 &   39.9 &    6.3 &    0.1 &    1.9 &    6.0 &    0.1 &    7.0 \\
%  k = 18        &   25.3 &    0.3 &    0.3 &   23.7 &    0.3 &   11.8 &   12.6 &    0.1 &    0.5 &   12.2 &    0.1 &    3.5 \\
%  k = 20        &   25.3 &    0.3 &    0.0 &   25.3 &    0.3 &    0.1 &   25.3 &    0.3 &    0.2 &   24.9 &    0.3 &    1.8 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.3 &    0.5 &    0.1 &   49.9 &    0.5 &    0.9 &   44.0 &    0.5 &   12.6 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &   12.3 &    0.2 &   21.2 &    4.3 &    0.1 &   42.6 &    0.6 &    0.0 &   49.1 &    0.2 &    0.0 &   48.2 \\
%  k = 14        &   40.5 &    0.5 &   11.8 &   18.8 &    0.2 &   20.7 &    4.8 &    0.1 &   41.2 &    2.8 &    0.0 &   24.3 \\
%  k = 16        &   73.7 &    0.7 &    2.4 &   22.2 &    0.3 &   22.4 &   12.6 &    0.1 &    1.8 &   11.0 &    0.1 &   22.4 \\
%  k = 18        &   25.3 &    0.3 &    0.0 &   25.3 &    0.3 &    0.1 &   25.3 &    0.3 &    0.5 &   24.9 &    0.3 &    3.3 \\
%  k = 20        &   50.4 &    0.5 &    0.0 &   50.3 &    0.5 &    0.2 &   48.8 &    0.5 &    6.1 &   25.3 &    0.3 &    0.1 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.3 &    0.5 &    0.1 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &   98.2 &    1.3 &   23.7 &   34.6 &    0.5 &   32.0 &    8.0 &    0.1 &   39.3 &    1.3 &    0.0 &   29.1 \\
%  k = 14        &  139.2 &    1.4 &    3.3 &   65.2 &    0.8 &   19.4 &   36.8 &    0.5 &   20.7 &   10.8 &    0.1 &   25.1 \\
%  k = 16        &  156.3 &    1.9 &   17.6 &   75.2 &    0.8 &    0.8 &   72.1 &    0.7 &    4.1 &   43.4 &    0.6 &   32.5 \\
%  k = 18        &  125.7 &    1.2 &    0.3 &  123.9 &    1.2 &    1.4 &   93.6 &    1.1 &   18.2 &   37.7 &    0.6 &   40.2 \\
%  k = 20        &   75.5 &    0.7 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 &   50.3 &    0.5 &    0.3 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  248.9 &    2.8 &   10.0 &  155.2 &    2.0 &   24.1 &   69.0 &    0.8 &   19.1 &   21.0 &    0.3 &   26.7 \\
%  k = 14        &  285.7 &    3.0 &    4.5 &  180.6 &    2.2 &   18.4 &  129.1 &    1.4 &    7.6 &   68.2 &    0.8 &   12.8 \\
%  k = 16        &  213.0 &    2.1 &    0.4 &  208.3 &    2.1 &    2.0 &  161.3 &    1.8 &   12.1 &  108.9 &    1.3 &   16.2 \\
%  k = 18        &  126.0 &    1.2 &    0.0 &  126.0 &    1.2 &    0.1 &  125.4 &    1.2 &    0.5 &  123.9 &    1.2 &    1.3 \\
%  k = 20        &  125.9 &    1.2 &    0.0 &  124.9 &    1.2 &    0.0 &  114.9 &    1.1 &    0.0 &   50.4 &    0.5 &    0.0 \\
%  k = 22        &  197.1 &    2.0 &    1.6 &   87.6 &    1.1 &   22.2 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \tiny
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a] Average number of spells per woman where scan has been used at least once.
% \item[b] Percent of spells where prenatal scan was initially used but birth is the result
% of pregnancy without prenatal scan.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% \begin{table}[!htp]
% \centering
% \scriptsize
% \begin{threeparttable}
% \caption{Abortions and Change in Abortion Use by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_abortion_90}
% \begin{tabular}{@{} c D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} @{}}
% \toprule
%                & \multicolumn{12}{c}{Discount rate ( $ d = 0.9 $ ) }                         \\ 
%                & \multicolumn{12}{c}{Cost of Prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{3}{c}{12}    & \multicolumn{3}{c}{16}    & \multicolumn{3}{c}{20}    & \multicolumn{3}{c}{24}    \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13} 
%                                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       \\
% \multirow{2}{9mm}{Cost of\\children}  & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         \\
%                                       & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} \\
% \midrule
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &    0.4 &    0.0 &    5.0 &    0.1 &    0.0 &   55.7 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 14        &    3.2 &    0.0 &    4.0 &    1.6 &    0.0 &    3.7 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 16        &    6.4 &    0.1 &    0.2 &    6.4 &    0.1 &    1.0 &    2.4 &    0.0 &   25.8 &    0.0 &    0.0 &      . \\
%  k = 18        &   19.1 &    0.3 &   40.4 &   12.7 &    0.1 &    0.3 &   11.8 &    0.1 &    6.4 &    0.0 &    0.0 &      . \\
%  k = 20        &   25.3 &    0.3 &    0.0 &   25.3 &    0.3 &    0.2 &   22.0 &    0.3 &   12.5 &    0.0 &    0.0 &      . \\
%  k = 22        &   50.4 &    0.5 &    0.1 &   49.5 &    0.5 &    1.6 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &    4.4 &    0.1 &   41.3 &    0.4 &    0.0 &    7.7 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 14        &   21.6 &    0.2 &   12.3 &    3.2 &    0.0 &    4.4 &    1.6 &    0.0 &    7.4 &    0.0 &    0.0 &      . \\
%  k = 16        &   12.7 &    0.1 &    0.2 &   12.3 &    0.1 &    6.2 &    6.4 &    0.1 &    1.0 &    5.2 &    0.1 &   28.9 \\
%  k = 18        &   25.3 &    0.3 &    0.1 &   25.2 &    0.3 &    0.9 &   12.7 &    0.1 &    0.3 &   12.5 &    0.1 &    1.7 \\
%  k = 20        &   50.3 &    0.5 &    0.2 &   25.3 &    0.3 &    0.0 &   25.3 &    0.3 &    0.1 &   25.1 &    0.3 &    0.8 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.3 &    0.5 &    0.1 &   49.5 &    0.5 &    1.6 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &   33.0 &    0.4 &   22.5 &    4.8 &    0.1 &   47.7 &    0.4 &    0.0 &   13.6 &    0.0 &    0.0 &      . \\
%  k = 14        &   43.9 &    0.4 &    1.7 &   26.1 &    0.3 &   11.2 &    4.8 &    0.1 &   41.2 &    1.6 &    0.0 &    7.4 \\
%  k = 16        &   75.6 &    0.8 &    0.4 &   69.2 &    0.7 &    7.0 &   12.6 &    0.1 &    1.8 &    6.4 &    0.1 &    1.0 \\
%  k = 18        &  124.8 &    1.2 &    0.9 &   25.3 &    0.3 &    0.1 &   25.3 &    0.3 &    0.3 &   24.9 &    0.3 &    3.3 \\
%  k = 20        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 &   50.3 &    0.5 &    0.4 &   37.7 &    0.6 &   40.1 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &  139.8 &    1.5 &    7.8 &   44.0 &    0.5 &   19.3 &    5.4 &    0.1 &   36.8 &    0.2 &    0.0 &   48.2 \\
%  k = 14        &  144.2 &    1.4 &    0.7 &  120.3 &    1.3 &   11.4 &   39.1 &    0.5 &   14.0 &    5.9 &    0.1 &   14.2 \\
%  k = 16        &  210.7 &    2.1 &    1.2 &  116.1 &    1.3 &   11.2 &   73.7 &    0.7 &    2.2 &   24.5 &    0.3 &    6.7 \\
%  k = 18        &  126.0 &    1.2 &    0.0 &  125.7 &    1.2 &    0.3 &  119.4 &    1.2 &    4.0 &   44.1 &    0.6 &   22.6 \\
%  k = 20        &  125.4 &    1.2 &    0.0 &   75.5 &    0.7 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \tiny
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a] Average number of spells per woman where scan has been used at least once.
% \item[b] Percent of spells where prenatal scan was initially used but birth is the result
% of pregnancy without prenatal scan.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% \begin{table}[!htp]
% \centering
% \scriptsize
% \begin{threeparttable}
% \caption{Abortions and Change in Abortion Use by Cost of \protect\linebreak Children and prenatal Sex  Determination}
% \label{tab:simul_abortion_85}
% \begin{tabular}{@{} c D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} D{.}{.}{3.1} D{.}{.}{1.1} D{.}{.}{1.1} @{}}
% \toprule
%                & \multicolumn{12}{c}{Discount rate ( $ d = 0.85 $ ) }                         \\ 
%                & \multicolumn{12}{c}{Cost of Prenatal Sex Determination ($ p_s $)}                         \\ 
%                & \multicolumn{3}{c}{12}    & \multicolumn{3}{c}{16}    & \multicolumn{3}{c}{20}    & \multicolumn{3}{c}{24}    \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10} \cmidrule(lr){11-13} 
%                                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       & \mco{Abortions}  & \mco{Average}           &                       \\
% \multirow{2}{9mm}{Cost of\\children}  & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         & \mco{per 100}    & \mco{Spells}            & \mco{Percent}         \\
%                                       & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} & \mco{Women}      & \mco{w.\ Scan\tnote{a}} & \mco{Change\tnote{b}} \\
% \midrule
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.5 $)}                         \\ 
%  k = 12        &    0.2 &    0.0 &   16.0 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 14        &    1.7 &    0.0 &    2.1 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 16        &    6.4 &    0.1 &    0.7 &    1.6 &    0.0 &   51.2 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 18        &   12.7 &    0.1 &    0.2 &    9.4 &    0.1 &   24.8 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 20        &   25.3 &    0.3 &    0.1 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 22        &   49.3 &    0.5 &    3.2 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.55 $)}                         \\ 
%  k = 12        &    0.4 &    0.0 &    5.8 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 14        &    3.3 &    0.0 &    2.0 &    1.6 &    0.0 &    3.7 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 16        &   12.3 &    0.1 &    6.2 &    6.4 &    0.1 &    1.3 &    1.6 &    0.0 &   51.2 &    0.0 &    0.0 &      . \\
%  k = 18        &   25.3 &    0.3 &    0.5 &   12.7 &    0.1 &    0.2 &   12.2 &    0.1 &    3.5 &    0.0 &    0.0 &      . \\
%  k = 20        &   25.3 &    0.3 &    0.0 &   25.3 &    0.3 &    0.1 &   24.9 &    0.3 &    1.8 &    0.0 &    0.0 &      . \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.3 &    0.5 &    0.1 &   24.4 &    0.2 &    0.0 &    0.0 &    0.0 &      . \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.6 $)}                         \\ 
%  k = 12        &    7.0 &    0.1 &   13.3 &    0.4 &    0.0 &   13.6 &    0.0 &    0.0 &      . &    0.0 &    0.0 &      . \\
%  k = 14        &   24.8 &    0.3 &    2.3 &    3.2 &    0.0 &    4.0 &    1.5 &    0.0 &   13.2 &    0.0 &    0.0 &      . \\
%  k = 16        &   72.0 &    0.7 &    4.3 &   12.5 &    0.1 &    3.5 &    6.3 &    0.1 &    2.4 &    1.6 &    0.0 &   51.2 \\
%  k = 18        &   25.3 &    0.3 &    0.0 &   25.3 &    0.3 &    0.3 &   12.7 &    0.1 &    0.2 &   12.2 &    0.1 &    3.5 \\
%  k = 20        &   50.4 &    0.5 &    0.0 &   50.2 &    0.5 &    0.9 &   25.3 &    0.3 &    0.1 &   25.2 &    0.3 &    0.5 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   49.8 &    0.5 &    1.8 \\
%                &  \multicolumn{12}{c}{Son Preference ($\alpha = 0.65 $)}                         \\ 
%  k = 12        &   52.3 &    0.6 &   14.5 &    4.4 &    0.1 &   41.3 &    0.1 &    0.0 &   55.7 &    0.0 &    0.0 &      . \\
%  k = 14        &  134.3 &    1.4 &    7.0 &   27.3 &    0.3 &    6.2 &    3.1 &    0.0 &   14.2 &    0.8 &    0.0 &   50.3 \\
%  k = 16        &   94.6 &    1.1 &   12.0 &   73.8 &    0.7 &    2.2 &   12.6 &    0.1 &    1.8 &    6.2 &    0.1 &    4.3 \\
%  k = 18        &  126.0 &    1.2 &    0.1 &  114.7 &    1.2 &    6.8 &   25.3 &    0.3 &    0.2 &   22.2 &    0.3 &   22.4 \\
%  k = 20        &   94.4 &    0.9 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.1 &   25.3 &    0.3 &    0.1 \\
%  k = 22        &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 &   50.4 &    0.5 &    0.0 \\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} \tiny
% \item \hspace*{-0.5em} \textbf{Note.} Simulations based on  50,000 individuals, 20 periods, $ I = 200 $,
%  and a probability of a son of 0.5.
% \item[a] Average number of spells per woman where scan has been used at least once.
% \item[b] Percent of spells where prenatal scan was initially used but birth is the result
% of pregnancy without prenatal scan.
% \end{tablenotes}
% \end{threeparttable}
% \normalsize
% \end{table}
% 
% 
% \clearpage
% \newpage
%

\section{Recall Error and the Sex Ratio}

\setcounter{figure}{0}
\setcounter{table}{0}

% \footnote{%
% This also has implications for analyses of child mortality and fertility.
% % Relatively little attention has been paid to this problem despite the
% % potential for substantially underestimating mortality, especially for girls.
% A companion paper examines fertility and mortality trends in India when
% account is taken of recall error.
% % The method suggested here can also be used to establish the extent of
% % recall error and what factors are associated with recall error.
% }
% This is important because the estimation method proposed
% rely on good quality data on both the sex of children born and the spacing between these 
% children.
% Recall error makes an analysis of the spread of sex determinations techniques
% less precise, because there may appear to be little change in the sex
% ratio over time even though the actual sex ratio at birth has increased.
% It also makes it more difficult to identify which families are using sex selective abortions.
%

Since the proposed method relies both on spacing between births and  birth outcomes, the 
quality of the results depends on the reliability of the birth 
histories provided by the respondents.
An important concern here is child mortality.
A respondent may not provide information about deceased children because of
the painful nature of the topic.
% First, deceased children may not be listed in a respondent's birth history because
% the sensitivity of the topic.
Furthermore, the likelihood of reporting a deceased child may depend on sex of the child. 
Systematic recall error, where children of a specific sex are missed non-randomly, 
is especially problematic because it directly biases the sex ratios results.

NFHS enumerators probe for any missed births, although the method depends on the survey.%
\footnote{%
NFHS-1 probe for each calendar birth interval that is 4 or more years.
NFHS-2 asked for stillbirths, spontaneous and induced abortions and also probed 
for each calendar birth interval 4 or more years.
NFHS-3 did not directly use birth intervals, but asked whether were there any other live births 
between (name of previous birth) and (name), including any children who died after birth. 
It did ask for births before the birth listed as first birth.
}
Probing catches many originally missed births but systematic recall error based on son
preference may still be a problem.
First, strong son preference lead to significantly higher mortality for girls than boys.
Secondly, for children who died son preference makes it more likely that parents will
remember boys than girls.
Finally, in the absence of sex selective abortions, son preference lead parents to have 
the next birth sooner if the last child was a girl than if it was a boy.
If this girl subsequently dies she is more likely to be missed because probing for missed 
births is only done for long intervals.%
\footnote{
As an example, imagine a family whose first child is a girl.
Because of son preference this girl is followed quickly by another child.
If this child is also a girl the parents will quickly try for a third child.
This can lead to three children in three years and
if the middle child dies and is not listed the enumerators will not probe for
missing births because the interval between the first and last child is only two years.
}

To assess the extent of systematic recall error Table \ref{tab:recallBirth} 
shows sex ratios of children recorded as first-born and second-born
by year of birth together with tests for whether the observed sex ratio is
significantly higher than the natural sex ratio.
Births are combined into five-year cohorts to achieve sufficient power for a test of 
differences in sex ratios.
Prenatal sex determination techniques did not become widely available until the 
mid-1980s and any significant deviation from the natural sex ratio before that time
is therefore likely the result of recall error.
Furthermore, births and deaths that took place longer ago may be more 
likely to be subject to recall error than more recent events.
The three NFHS surveys make it possible to examine whether recall error increases with 
time by comparing sex ratios for births that took place during the same years but were 
surveyed at different times.
Table \ref{tab:recallBirth} therefore also shows tests for whether more recent surveys 
have a higher sex ratio for the cohort than older surveys.

% Main points:
% - births that took place further away from the survey show a higher sex ratio
% - there is evidence of a U-shaped relationship in sex ratio over time for second
% and especially the three parities combined with cohorts furthest away from the
% survey showing the highest sex ratios. The sex ratio then declines and only begins
% claiming again when sex selective abortions became available in the mid-1980s.
% The exception to this pattern is children listed as first-born for whom the
% sex ratios declines steadily over time leveling off at the natural sex ratio.
% - lack of power for individual parities means that relatively few cohorts show
% significantly different sex ratios across surveys; there is more power for
% the three first parities combined

% [Is this worth including?]
% the reason that sex ratio for second born can be higher in more recent surveys
% (NFHS-3 compared to NFHS-1) is that a higher proportion of women in the
% recent survey have lower desired fertility and are better educated.
% They should also be younger, as in born more recently.

% The decline in sex ratio over time followed by the increase after sex selection
% became available may be best argument against referee 1's concern about whether 
% this is simply infanticide.

\begin{table}[!htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{Observed sex ratios by year of birth in five-year cohorts }
\label{tab:recallBirth}
\begin{tabular} {@{} l D{.}{.}{1.6} D{.}{.}{1.6} D{.}{.}{1.6} l D{.}{.}{1.6} D{.}{.}{1.6} D{.}{.}{1.6} l  @{}} \toprule
                   & \multicolumn{4}{c}{Listed as first-born}                          & \multicolumn{4}{c}{Listed as second-born} \\ \cmidrule(lr){2-5} \cmidrule(lr){6-9}
                   & \mco{NFHS-1}      & \mco{NFHS-2}      & \mco{NFHS-3}      & Diff.         & \mco{NFHS-1}      & \mco{NFHS-2}      & \mco{NFHS-3}      & Diff.              \\
                   & \mco{1992–-1993}  & \mco{1998–-1999}  & \mco{2005–-2006}  & test\tnote{a} & \mco{1992–-1993}  & \mco{1998–-1999}  & \mco{2005–-2006}  & test\tnote{a} \\ \midrule
1960--1964         & 0.5473\sym{***}   &      .\sym{}      &      .\sym{}      &       & 0.5172\sym{}      &      .\sym{}      &      .\sym{}      &           \\
                   & (0.0004)          & (     .)          & (     .)          &       & (0.4382)          & (     .)          & (     .)          &           \\
                   & \mco{[2,673]}     & \mco{[.]}         & \mco{[.]}         &       & \mco{[1,189]}     & \mco{[.]}         & \mco{[.]}         &           \\
1965--1969         & 0.5274\sym{**}    & 0.5538\sym{***}   &      .\sym{}      & A     & 0.5327\sym{**}    & 0.5244\sym{}      &      .\sym{}      &           \\
                   & (0.0306)          & (0.0002)          & (     .)          &       & (0.0124)          & (0.3310)          & (     .)          &           \\
                   & \mco{[5,429]}     & \mco{[1,997]}     & \mco{[.]}         &       & \mco{[3,888]}     & \mco{[595]}       & \mco{[.]}         &           \\
1970--1974         & 0.5354\sym{***}   & 0.5341\sym{***}   & 0.5411\sym{}      &       & 0.5212\sym{}      & 0.5313\sym{**}    & 0.6160\sym{**}    & BC        \\
                   & (0.0001)          & (0.0019)          & (0.1208)          &       & (0.1497)          & (0.0248)          & (0.0143)          &           \\
                   & \mco{[7,763]}     & \mco{[5,546]}     & \mco{[523]}       &       & \mco{[6,249]}     & \mco{[3,499]}     & \mco{[125]}       &           \\
1975--1979         & 0.5207\sym{}      & 0.5151\sym{}      & 0.5255\sym{*}     &       & 0.5185\sym{}      & 0.5262\sym{**}    & 0.5384\sym{**}    & B         \\
                   & (0.1258)          & (0.4669)          & (0.0920)          &       & (0.2477)          & (0.0340)          & (0.0225)          &           \\
                   & \mco{[8,749]}     & \mco{[7,478]}     & \mco{[3,762]}     &       & \mco{[7,819]}     & \mco{[6,220]}     & \mco{[1,811]}     &           \\
1980--1984         & 0.5210\sym{*}     & 0.5244\sym{**}    & 0.5280\sym{***}   &       & 0.5236\sym{**}    & 0.5236\sym{**}    & 0.5249\sym{*}     &           \\
                   & (0.0872)          & (0.0273)          & (0.0094)          &       & (0.0361)          & (0.0492)          & (0.0628)          &           \\
                   & \mco{[11,147]}    & \mco{[9,709]}     & \mco{[7,718]}     &       & \mco{[9,950]}     & \mco{[8,478]}     & \mco{[5,532]}     &           \\
1985--1989         & 0.5187\sym{}      & 0.5132\sym{}      & 0.5118\sym{}      &       & 0.5157\sym{}      & 0.5194\sym{}      & 0.5165\sym{}      &           \\
                   & (0.1916)          & (0.6153)          & (0.7091)          &       & (0.4099)          & (0.1695)          & (0.3716)          &           \\
                   & \mco{[11,276]}    & \mco{[11,048]}    & \mco{[9,422]}     &       & \mco{[10,296]}    & \mco{[9,857]}     & \mco{[8,022]}     &           \\
1990--1994         & 0.5174\sym{}      & 0.5185\sym{}      & 0.5183\sym{}      &       & 0.5120\sym{}      & 0.5166\sym{}      & 0.5289\sym{***}   & BC        \\
                   & (0.3263)          & (0.2006)          & (0.2207)          &       & (0.6570)          & (0.3430)          & (0.0030)          &           \\
                   & \mco{[6,544]}     & \mco{[11,653]}    & \mco{[10,601]}    &       & \mco{[5,869]}     & \mco{[10,653]}    & \mco{[9,299]}     &           \\
1995--1999         &      .\sym{}      & 0.5222\sym{*}     & 0.4980\sym{}      &       &      .\sym{}      & 0.5230\sym{*}     & 0.5262\sym{**}    &           \\
                   & (     .)          & (0.0788)          & (0.9998)          &       & (     .)          & (0.0664)          & (0.0103)          &           \\
                   & \mco{[.]}         & \mco{[8,642]}     & \mco{[11,111]}    &       & \mco{[.]}         & \mco{[7,990]}     & \mco{[9,956]}     &           \\
2000--2006         &      .\sym{}      &      .\sym{}      & 0.5135\sym{}      &       &      .\sym{}      &      .\sym{}      & 0.5292\sym{***}   &           \\
                   & (     .)          & (     .)          & (0.5978)          &       & (     .)          & (     .)          & (0.0008)          &           \\
                   & \mco{[.]}         & \mco{[.]}         & \mco{[13,413]}    &       & \mco{[.]}         & \mco{[.]}         & \mco{[11,645]}    &           \\
\bottomrule
\end{tabular}                        
\begin{tablenotes} \scriptsize
\item \hspace*{-0.6em} \textbf{Note.} 
Sample consists of Hindu women only.
First number in cell is ratio of boys to girls. 
Second number in parentheses is p-value for the hypothesis that observed sex ratio is 
greater than 105/205 using a binomial probability test (bitest in Stata 13)
with asterisks indicating significance as follows: 
* sign.\ at 10\%; ** sign.\ at 5\%; *** sign.\ at 1\%.
Third number in square brackets is number of observations.
\item[a] 
Test (prtest in Stata 13) whether recall error increases 
with time passed, which would manifest itself in a higher sex ratio for a more recent
survey than an earlier for the same cohort.
A: Cohort sex ratio significantly larger in NFHS-2 than NFHS-1 at the 10 percent level. 
B: Cohort sex ratio significantly larger in NFHS-3 than NFHS-1 at the 10 percent level. 
C: Cohort sex ratio significantly larger in NFHS-3 than NFHS-2 at the 10 percent level. 
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

The ``first-born'' sex ratios illustrate the systematic recall error problem well. 
According to the birth histories in NFHS-1 almost 55 percent of children born between 1960 
and 1964 were boys.
Given that this is more than two decades before sex selection techniques became available
in India the most likely explanation for the skewed sex ratio is that some children
listed as first-borns were not, in fact, the first children born in their families.
Instead, for a substantial proportion of families the first-born was a girl who died and was 
skipped when enumerators asked about birth history.

% Recall error is less of an issue for more recent births than for births that took place 
% longer ago.
As expected, the difference between the observed sex ratio and the natural sex ratio is 
less pronounced the closer to the survey date the cohort is.
The observed sex ratio for children born just prior to the NFHS-1 survey and listed as 
first-born is 0.517, which is not statistically significant different from the
natural sex ratio.
The same general pattern holds for NFHS-2 and NFHS-3, with cohorts further away
from the survey date more likely to have a sex ratio skewed male.

Second births show a pattern very similar to that for first births.
An interesting difference is that there is evidence of a U-shaped relationship between
time and sex ratio for second births.
Cohorts furthest away from the survey year show the highest sex ratios, but declines to the
natural sex ratio in the mid-1980s, and the sex ratio is then significantly higher again 
for more recent births.
This is in line with the results in the paper that show that sex selective abortions take 
place on lower parity births as desired fertility declined.

Finally, across surveys the same cohort tends to show a higher sex ratio the more recent 
the survey (births in the cohort took place longer ago), especially for second-borns.%
\footnote{
Births in the 1970-1974 cohort, for example, took place
between 18 and 23 years ago for NFHS-1, 
between 24 and 29 years ago for NFHS-2,
and  between 31 and 36 years ago for NFHS-3.
}
Despite this, few cohorts show significantly different sex ratios across surveys, most 
likely because of a lack of power.

Using year of birth to analyze recall error is problematic because year of birth for a 
given parity is affected by recall error; for example, a second born child listed as first 
born will naturally be born later than the real first born child.
Recorded year of marriage should, however, be affected neither by parental recall error 
nor use of sex selective abortions.
Table \ref{tab:recallMarriage} therefore shows sex ratios of children recorded as 
first-born and second-born by year of parents'
marriage, together with tests for whether the observed sex ratio is
significantly higher than the natural sex ratio and whether more recent surveys show
a higher sex ratio for the cohort than older surveys.

The basic recall error pattern remains with women married longer ago more
likely to report that their first-born is a boy.
If anything the pattern is more consistent when using year of marriage than 
year of birth. 
Before the introduction of sex selection there are only two instances  across the two 
parities where the observed sex ratio is lower for women married longer ago than for women 
married closer to the survey date.%
\footnote{
For first-borns it is between 1970--1974 and 1975-1979 for NFHS-3 and for second-borns
it is between 1975--1979 and 1980--1984 for NFHS-1.
}
Similarly, comparing women married in the same five-year period across surveys shows
that women married longer ago are more likely to report having a son.

\begin{table}[htbp]
\begin{center}
\begin{small}
\begin{threeparttable}
\caption{Observed sex ratios by year of parents' marriage in five-year cohorts }
\label{tab:recallMarriage}
\begin{tabular} {@{} l D{.}{.}{1.6} D{.}{.}{1.6} D{.}{.}{1.6} l D{.}{.}{1.6} D{.}{.}{1.6} D{.}{.}{1.6} l  @{}} \toprule
                   & \multicolumn{4}{c}{Listed as first-born}                          & \multicolumn{4}{c}{Listed as second-born} \\ \cmidrule(lr){2-5} \cmidrule(lr){6-9}
                   & \mco{NFHS-1}      & \mco{NFHS-2}      & \mco{NFHS-3}      & Diff.         & \mco{NFHS-1}      & \mco{NFHS-2}      & \mco{NFHS-3}      & Diff.              \\
                   & \mco{1992–-1993}  & \mco{1998–-1999}  & \mco{2005–-2006}  & test\tnote{a} & \mco{1992–-1993}  & \mco{1998–-1999}  & \mco{2005–-2006}  & test\tnote{a} \\ \midrule
1960--1964         & 0.5368\sym{***}   &      .\sym{}      &      .\sym{}      &        & 0.5274\sym{**}    &      .\sym{}      &      .\sym{}      &          \\
                   & (0.0002)          & (     .)          & (     .)          &        & (0.0239)          & (     .)          & (     .)          &          \\
                   & \mco{[6,200]}     & \mco{[.]}         & \mco{[.]}         &        & \mco{[6,001]}     & \mco{[.]}         & \mco{[.]}         &          \\
1965--1969         & 0.5333\sym{***}   & 0.5450\sym{***}   &      .\sym{}      &        & 0.5277\sym{**}    & 0.5401\sym{***}   &      .\sym{}      &          \\
                   & (0.0011)          & (0.0000)          & (     .)          &        & (0.0181)          & (0.0005)          & (     .)          &          \\
                   & \mco{[6,707]}     & \mco{[4,268]}     & \mco{[.]}         &        & \mco{[6,449]}     & \mco{[4,142]}     & \mco{[.]}         &          \\
1970--1974         & 0.5244\sym{**}    & 0.5222\sym{}      & 0.5271\sym{}      &        & 0.5182\sym{}      & 0.5235\sym{*}     & 0.5384\sym{**}    & B        \\
                   & (0.0386)          & (0.1120)          & (0.1372)          &        & (0.2651)          & (0.0805)          & (0.0194)          &          \\
                   & \mco{[8,154]}     & \mco{[6,546]}     & \mco{[1,973]}     &        & \mco{[7,843]}     & \mco{[6,308]}     & \mco{[1,913]}     &          \\
1975--1979         & 0.5272\sym{***}   & 0.5202\sym{}      & 0.5316\sym{***}   & C      & 0.5157\sym{}      & 0.5183\sym{}      & 0.5278\sym{**}    & B        \\
                   & (0.0062)          & (0.1477)          & (0.0050)          &        & (0.4208)          & (0.2499)          & (0.0242)          &          \\
                   & \mco{[9,829]}     & \mco{[8,644]}     & \mco{[5,798]}     &        & \mco{[9,295]}     & \mco{[8,294]}     & \mco{[5,608]}     &          \\
1980--1984         & 0.5151\sym{}      & 0.5142\sym{}      & 0.5195\sym{}      &        & 0.5198\sym{}      & 0.5171\sym{}      & 0.5187\sym{}      &          \\
                   & (0.4616)          & (0.5322)          & (0.1877)          &        & (0.1514)          & (0.3178)          & (0.2332)          &          \\
                   & \mco{[10,814]}    & \mco{[9,924]}     & \mco{[8,316]}     &        & \mco{[9,788]}     & \mco{[9,409]}     & \mco{[7,906]}     &          \\
1985--1989         & 0.5176\sym{}      & 0.5209\sym{*}     & 0.5095\sym{}      &        & 0.5131\sym{}      & 0.5194\sym{}      & 0.5267\sym{**}    & B        \\
                   & (0.2753)          & (0.0925)          & (0.8425)          &        & (0.5918)          & (0.1651)          & (0.0105)          &          \\
                   & \mco{[9,971]}     & \mco{[10,990]}    & \mco{[9,717]}     &        & \mco{[5,790]}     & \mco{[10,130]}    & \mco{[9,077]}     &          \\
1990--1994         & 0.5180\sym{}      & 0.5182\sym{}      & 0.5125\sym{}      &        & 0.4366\sym{}      & 0.5191\sym{}      & 0.5251\sym{**}    & AB       \\
                   & (0.3836)          & (0.2282)          & (0.6692)          &        & (0.9740)          & (0.2097)          & (0.0199)          &          \\
                   & \mco{[2,170]}     & \mco{[10,619]}    & \mco{[10,574]}    &        & \mco{[142]}       & \mco{[7,994]}     & \mco{[9,600]}     &          \\
1995--1999         &      .\sym{}      & 0.5243\sym{*}     & 0.5017\sym{}      &        &      .\sym{}      & 0.5613\sym{***}   & 0.5312\sym{***}   &          \\
                   & (     .)          & (0.0845)          & (0.9965)          &        & (     .)          & (0.0015)          & (0.0009)          &          \\
                   & \mco{[.]}         & \mco{[5,081]}     & \mco{[10,960]}    &        & \mco{[.]}         & \mco{[1,019]}     & \mco{[8,969]}     &          \\
2000--2006         &      .\sym{}      &      .\sym{}      & 0.5169\sym{}      &        &      .\sym{}      &      .\sym{}      & 0.5247\sym{}      &          \\
                   & (     .)          & (     .)          & (0.3280)          &        & (     .)          & (     .)          & (0.1243)          &          \\
                   & \mco{[.]}         & \mco{[.]}         & \mco{[9,210]}     &        & \mco{[.]}         & \mco{[.]}         & \mco{[3,316]}     &          \\
\bottomrule
\end{tabular}                        
\begin{tablenotes} \scriptsize
\item \hspace*{-0.6em} \textbf{Note.} 
Sample consists of Hindu women only.
First number in cell is ratio of boys to girls. 
Second number in parentheses is p-value for the hypothesis that observed sex ratio is 
greater than 105/205 using a binomial probability test (bitest in Stata 13)
with asterisks indicating significance as follows: 
* sign.\ at 10\%; ** sign.\ at 5\%; *** sign.\ at 1\%.
Third number in square brackets is number of observations.
\item[a] 
Test (prtest in Stata 13) whether recall error increases 
with time passed, which would manifest itself in a higher sex ratio for a more recent
survey than an earlier for the same cohort.
A: Cohort sex ratio significantly larger in NFHS-2 than NFHS-1 at the 10 percent level. 
B: Cohort sex ratio significantly larger in NFHS-3 than NFHS-1 at the 10 percent level. 
C: Cohort sex ratio significantly larger in NFHS-3 than NFHS-2 at the 10 percent level. 
\end{tablenotes}
\end{threeparttable}
\end{small}
\end{center}
\end{table}

The relationship between length of marriage and recall error can also seen in 
Figure \ref{fig:sexRatioMarriage}, which shows the observed sex ratio for children reported
as first born as a function of duration of marriage at the time of the survey combining 
all three surveys.
The solid line is the sex ratio of children reported 
as first born by the number of years between the survey and marriage, 
the dashed lines indicate the 95 percent confidence interval 
and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes the years are grouped in twos.
In line with the results from Table \ref{tab:recallMarriage}, the observed ratio of boys 
is increasingly above the expected value the longer ago the parents were married.%
\footnote{
The graph for second births shows a similar pattern with the
likelihood of the second child being a boy going up with increasing marriage duration.
The graphs for the second births and the individual survey rounds are available upon request.
% It is interesting to note that the graphs of ``first'' births for NFHS-1 and 2 
% very clearly show the effect of the 1965-1967 drought in India.
% There is a sharp drop in the percentage boys corresponding to the drought.
% For the surrounding years the percentage of boys among ``first-born'' is around
% 54 percent and increasing, but for the drought years the percentage is slightly
% \emph{below} the natural.
% See also the discussion in Section \ref{sec:review} on nutrition and sex ratios.
}

[Figure sex ratio marriage here!]

% \begin{figure}[htp]%
% \centering
% \includegraphics[width=0.9\textwidth]{sex_ratio_marriage}
% \caption{Sex ratio of ``first'' births by duration of marriage}
% \label{fig:sexRatioMarriage}}
% \end{figure}

The increasingly unequal sex ratio with increasing marriage duration suggests that
a solution to the recall error problem is to drop observations for 
women who were married ``too far'' from the survey year.
The main problem is establishing what the best cut-off point should be.
As Table \ref{tab:recallMarriage} shows there are differences in recall error across the
three surveys and between parities.
One possibility is simply to drop all women married in cohorts where the observed
sex ratio is substantially higher than the natural sex ratio, but even this is not
straightforward.
For ``first'' born this would imply dropping all women married before 1980, but
for ``second'' born it would be imply dropping all married before 1970 for NFHS-1, before 1975 
for NFHS-2, and before 1980 NFHS-3.
Because of the differences between ``first'' and ``second'' born in the different surveys
I vary the cut-off point by survey round.

NFHS-3 is the most straightforward in that the implied cut-off point is the same for
both first and second born.
Based on Table \ref{tab:recallMarriage}, women in NFHS-3 who were married 26 years or
more before the survey date are dropped.
For NFHS-1 and NFHS-2 the trade-off is between ensuring enough information and not bias 
the results unduly.
Given that sex selective abortions are more likely in higher parity birth I base the
cut-off point on the second births instead of the first births.
Hence, for NFHS-1 women who were married 22 years or more before the survey date 
were dropped and for NFHS-2 women who were married 23 year or more before the survey date 
were dropped.
The final sample consists of 146,096 women, with 332,951 
parity one through four births.

% \begin{figure}
% \centering
% \subfloat[NFHS-1]{\includegraphics[width=.49\textwidth]{des_b1_m_age_r1}}
% \subfloat[NFHS-2]{\includegraphics[width=.49\textwidth]{des_b1_m_age_r2}} \\
% \subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{des_b1_m_age_r3}} 
% \caption{Ratio of Boys for ``First'' Births by Survey Round}
% \label{fig:sex_ratio_recall_rounds}
% \end{figure}

% This indicates the problem with using older data for analyzing sex selective abortions.
% Even more important it shows why using year of birth as a cut-off point
% for addressing recall error is problematic.

\subsection{Sensitivity of Results to Recall Error}

% [referee 2 wants a discussion of how results are impacted by changes in number of
% women dropped because of recall error]

The last three sections in this Appendix show results for the three education groups for 
three samples: the preferred sample used in the paper (Figures \ref{fig:results_spell1_low}
through \ref{fig:results_spell4_high_rural}), 
a more restricted sample where only women married less than 19 years are 
included (Figures \ref{fig:results_limit_spell1_low} through \ref{fig:results_limit_spell3_high_rural}),
and the sample of women dropped because of potential recall error bias, i.e.\ all women 
dropped from the data to reach the preferred sample (Figures \ref{fig:results_dropped_spell1_low}
through \ref{fig:results_dropped_spell3_high_rural}).
For the ``dropped'' sample it is only possible to estimate sex ratios for the period
before sex-selective abortions became available.
The period covered is different from the other two samples because the births for 
the sample takes place over a much longer period; the sample covers spells that begin
between 1950 and 1984.

The different results for women with no education illustrates the potential problems 
with recall error.
Figures \ref{fig:results_spell1_low}, \ref{fig:results_limit_spell1_low}, and 
\ref{fig:results_dropped_spell1_low} show the different results for women with no
education across the three samples.
The preferred and the restricted samples show almost identical results with the main
difference that the restricted sample has wider confidence intervals because of the
smaller sample size.
The results from the dropped sample, on the other hand, show sex ratios that are 
statistically significantly above the natural sex ratio in many periods, which is
consistent with recall error.
Furthermore, the survival curves are substantially closer to being linear indicating
that observed births took place further away from the time of marriage as expected if 
the couple had a girl first but counted a subsequent son as their first-born.

The second spell results demonstrate the trade-off inherent in dealing with recall error.
While the results for women with no education are similar across the preferred and the
restricted samples in rural areas (Figures \ref{fig:results_spell2_low_rural} and
\ref{fig:results_limit_spell2_low_rural}), the results for urban women with no 
education in the restricted sample are much less stable and have substantially wider
confidence intervals than the preferred sample (Figures \ref{fig:results_spell2_low_urban}
and \ref{fig:results_limit_spell2_low_urban}).
This happens because the restricted sample is down to less than 900 urban women, compared
to close to 1,600 urban women in the preferred sample.

An interesting difference to this pattern is for the third spell for women with 8 or more
years of education.
This group shows no evidence on an unequal sex ratio for the ``dropped'' sample in 
either urban or rural areas (Figures \ref{fig:results_dropped_spell3_high_urban} and
\ref{fig:results_dropped_spell3_high_rural}).
However, for the preferred sample there are consistently more boys than girls for urban
women when the first two children are either two girls or one girl and one boy (Figure 
\ref{fig:results_spell3_high_urban}).
Furthermore, the sex ratios become even more unequal when restricting the sample to
women married less than 19 years (Figure \ref{fig:results_limit_spell3_high_urban}), 
although the sample sizes also become substantially smaller.
This is because spell periods are based on the starting year of the spell.
Hence, some of the pregnancies in the 1972--1984 period will have been exposed to the  
availability of prenatal sex determination.
For Figure \ref{fig:results_spell3_high_urban}(a) 25 percent of observed births 
occur in  1986 or after, meaning that conception took place in the 1985--1994 period, 
while for Figure \ref{fig:results_spell3_high_urban}(d) the number is 29 percent.
Hence, it is possible that the unequal sex ratios observed are, indeed, evidence of 
sex-selective abortions.

\clearpage
\newpage

\subsection{Comparing the Hazard and Simple Models}

As detailed in the main paper, one of the main advantages of the hazard model 
approach is that it can be used to predict what completed fertility, number of abortions, 
and sex ratios will be once women are done with childbearing.
This cannot be done with the simple model because it does not predict fertility
progression, does not take into account censoring, and cannot capture if parents
change their use of sex selection within a spell.
We can, however, still compare differences in predicted sex ratios for 
\emph{individual} parities.%
\footnote{
The simple model uses only observed births and the predicted sex ratio is 
therefore simply what we observe for the relevant uncensored sample.
}

For each parity, the simple model is
\begin{equation}
Y_i = \gamma + \alpha' \mathbf{Z_i} + \beta' \mathbf{X}_{i} + \epsilon_i,
\end{equation}
where $Y$ is a dummy variable that takes the value 1 if the child born is a boy and zero 
if it is a girl, and the vectors of explanatory variables, $\mathbf{Z}_{i}$ and 
$\mathbf{X}_{i}$, are the same as for the proposed method, except that 
$\mathbf{Z}_{i}$ obviously does not include the baseline hazard and that there are 
no time dependence.%
\footnote{
To ease presentation, the indicator for parity number is not shown.
}
I estimate this using a logit model, although a linear probability model leads to
similar results.
This model is estimated on all observed births for a given parity in the relevant 
period to ensure that the results are as close to the standard method of 
examining sex selection \citep[see, for example, ][]{retherford03b,jha06,abrevaya09}.

\begin{table}[htbp]
\begin{center}
\begin{footnotesize}
\begin{threeparttable}
\caption{Comparison of Hazard and Simple Models for  \protect\linebreak
Women with 8 or More Years of Education}
\label{tab:compare2}
\begin{tabular} {@{} l D{.}{.}{5.2} D{.}{.}{5.2} D{.}{.}{5.2} D{.}{.}{5.2}   @{}} \toprule
                                  &  \multicolumn{2}{c}{Urban}        & \multicolumn{2}{c}{Rural}         \\ \cmidrule(lr){2-3} \cmidrule(lr){4-5}
                                  & \mco{Simple}     &  \mco{Hazard}   &  \mco{Simple}    & \mco{Hazard}      \\   \midrule
                                  & \multicolumn{4}{c}{1985--1994 - Second Birth} \\ \cmidrule(l){2-5}
                                  & \multicolumn{4}{c}{1 girl} \\
Number of women\tnote{a}          &             & 4,869   &         & 3,105   \\
Births\tnote{b}                   &     3,848   & 3,926   & 2,417   & 2,795   \\
Percentage boys                   &      54.0   &  54.7   &  52.6   &  53.5   \\
Sex ratio (boys per 100 girls)    &     117.4   & 120.7   & 110.9   & 115.1   \\
\addlinespace 
                                  & \multicolumn{4}{c}{1985--1994 - Third Birth}  \\ \cmidrule(l){2-5}
%                                   & \multicolumn{4}{c}{1 boy and 1 girl}        \\
% Number of women\tnote{a}          &             & 3,460   &         & 1,987   \\
% Births\tnote{c}                   &     1,431   & 1,558   & 1,107   & 1,336   \\
% Percentage boys                   &      53.0   &  52.5   &  53.5   &  54.6   \\
% Sex ratio (boys per 100 girls)    &     112.9   & 110.4   & 115.0   & 120.2   \\
% \addlinespace
                                  & \multicolumn{4}{c}{2 girls}        \\
Number of women\tnote{a}          &             & 1,654   &         &   982   \\
Births\tnote{c}                   &     1,034   & 1,092   &   682   &   853   \\
Percentage boys                   &      60.8   &  63.7   &  54.7   &  57.4   \\
Sex ratio (boys per 100 girls)    &     155.3   & 175.3   & 120.7   & 134.6   \\
\addlinespace 
                                  & \multicolumn{4}{c}{1995--2006 - Second Birth} \\ \cmidrule(l){2-5}
                                  & \multicolumn{4}{c}{1 girl} \\                                                  
Number of women\tnote{a}          &             & 3,774   &         & 2,823   \\
Births\tnote{b}                   &     3,031   & 2,704   & 2,489   & 2,407   \\
Percentage boys                   &      58.0   &  58.6   &  56.0   &  56.9   \\
Sex ratio (boys per 100 girls)    &     138.1   & 141.3   & 127.1   & 132.2   \\
\addlinespace 
                                  & \multicolumn{4}{c}{1995--2006 - Third Birth}  \\ \cmidrule(l){2-5}
%                                   & \multicolumn{4}{c}{1 boy and 1 girl}        \\
% Number of women\tnote{a}          &             & 2,357   &         & 1,819   \\
% Births\tnote{c}                   &       830   &   775   &   853   &   998   \\
% Percentage boys                   &      53.9   &  52.9   &  55.5   &  56.5   \\
% Sex ratio (boys per 100 girls)    &     116.7   & 112.1   & 124.5   & 129.8   \\
% \addlinespace
                                  & \multicolumn{4}{c}{2 girls}        \\
Number of women\tnote{a}          &             & 1,000   &         &   863   \\
Births\tnote{c}                   &       695   &   600   &   690   &   661   \\
Percentage boys                   &      65.5   &  62.4   &  62.0   &  61.7   \\
Sex ratio (boys per 100 girls)    &     189.6   & 166.0   & 163.4   & 161.0   \\
\bottomrule
\end{tabular}                        
\begin{tablenotes} \tiny
\item \hspace*{-0.5em} \textbf{Note.} 
The simple models are estimated using logit on all births for a given parity that
occurred during the relevant time period.
The hazard models are estimated using all spells that began in the relevant time period.
For the second birth the hazard model covers the period from beginning of spell to 6 
years (24 quarters) after the birth of the first child.
For the third birth it covers the period from beginning of spell to 7.25 years 
(29 quarters) after the birth of the second child for the period 1985--1994 and from 
beginning of spell to 5.75 years (23 quarters) after the birth of the second child for 
the period 1995--2006.
\item[a] Number of women in period who began the spell with given sex composition of prior 
child/children in the relevant period.
\item[b] For Logit model the number of women who are observed to have a second birth
in the period. 
For hazard model the predicted number of second births that will occur between beginning 
of the spell and 6 years (24 quarters) after the birth of the first child.
% \item[c] Number of women in period who had one boy and one girl as their first two children.
\item[c] For Logit model the number of women who are observed to have a third birth
in the period. 
For hazard model the predicted number of third births that will occur between beginning of 
the spell and 7.25 years (29 quarters) after the birth of the second child for the
1985--1994 period and between beginning of the spell and 5.75 years (23 quarters) after 
the birth of the second child for the 1995--2006 period.
\end{tablenotes}
\end{threeparttable}
\end{footnotesize}
\end{center}
\end{table}

I focus on those cases where there is evidence of sex-selective abortions: the spells
from parity one to two and two to three for women with eight or more years of education.%
\footnote{
The results for the spell from parity three to four are available on request.
They are not presented because the number of educated women who end up with three births 
and the probability of having a fourth birth are small, which makes
comparing the simple and hazard models with any precision difficult.
}
Table \ref{tab:compare2} shows the number of women who enter the second and third spells
for the hazard model,
the number of births, observed in the case of the simple model and predicted in the case 
of the hazard model, predicted percent boys, and the associated predicted sex ratio.
It is possible for the number of births used for the simple model to be larger than the
hazard models' predicted number of births because the simple model is based
on all births that occurred in the period---independently of when the prior birth
occurred---whereas the hazard model uses only spells that began in the period.

For the 1985--1994 period---which is when sex selection become available---the 
hazard model consistently predicts a higher sex ratio than the logit model.
Furthermore, some of these differences are substantial.
For urban women with 2 girls as their first two children, the hazard model
predicts a sex ratio of 175 boys per 100 girls, whereas the simple method
``only'' predicts 155 boys per 100 girls.
The difference is smaller in rural areas but still 14 boys per 100 girls
higher for the hazard model than for the simple model.
A similar, if less pronounced pattern, show up for the second birth,
where the hazard model predict 3 and 5 boys more per 100 girls for
urban and rural women.

For the latest period, 1995--2006, the hazard model still produces higher 
predicted sex ratios for second birth and the differences are of a 
similar magnitude to the 1985--1994 period.
The predicted sex ratios for the third birth are, however, 
higher for the simple model than for the hazard model.
This difference is substantially for urban women where the difference 
is 23 boys extra per 100 girls.

A possible explanation for these differences is that the simple model 
fails to capture the move towards lower fertility and the subsequent 
increase in the use of sex-selective abortions at earlier parities that 
occurred over the periods.
The results also suggest that the take-up of sex selection occurred 
rapidly as the new methods became available and that this take-up
is not fully captured by the simple model.
More broadly, a possible interpretation of the differences is that
the simple model will tend to underestimate when sex-selective abortions
are initially spreading in use and overestimate when the spread of sex 
selection is slowing.
This is most clearly seen for the third births for urban women where
the hazard model show a \emph{decline} is use of sex-selective abortions,
consistent with the lower fertility as discussed in the main paper,
while the simple model show a substantial increase.

\clearpage
\newpage

\section{Potential for Reverse Causation in Education?}

The data provides no direct way to tell whether a mother is currently in
school or have returned to school.
The questions in the survey simply ask whether the respondent has ever
attended school, and if yes, what the highest level (standard) completed is.
My best estimate of the age somebody left school is therefore their years of
schooling completed plus 6 years for school starting age.
Based on this, more than 75 percent exit school at age 16 or before,
while 96 percent have stopped school by age 21.

It is possible to get a sense of whether women might return to school after 
giving birth using this information.
To find the year when a woman would have completed school I take her birth
year and add the estimated age at school completion.
If the year the woman is estimated to have completed school is equal to or greater than 
the year of the first birth it is conceivable that the woman may have returned to 
school after giving birth.
This ignores the possibility of grade repetitions, but if somebody needs to
repeat grades it is unlikely that they will go on to secondary or higher education.
Out of 146,096 women in the main sample only 2,615, or less than 2 percent, fall 
into this category.
Furthermore, only 322 are possibly in school at the time of the second school.
Among women who appear to return to school after their first child the sex ratio of their
first birth is skewed somewhat more male than normal (54 percent boys), so their 
return may, indeed, be tied to the sex of their first child.

The estimates are conservative in the sense that it is possible to finish school 
first and then have the child later in the year.
Using if the estimated end of schooling comes \emph{after} the year the child 
is born the numbers fall to 1,234, or less than 1 percent, for
the first birth and 162 for the second birth.

How much of a problem is this then?
It is, indeed, possible that some women return to school after their first child,
but it appears to be relatively  uncommon.
Furthermore, almost nobody are in school after their second child.
Since there is little evidence of sex-selective abortions among first-born children
the possibility of return to school should not affect the estimates.

Furthermore, I use 8 years of education as the cut-off point and hence only women who
had their first child before age 14 would potentially be misclassified in terms of
education group for their first birth. 
There are only 56 women who end up with 8 or more years of education but had their
first birth at age 12 or 13.
None are possibly misclassified for second birth or above, which is where sex-selective 
abortions become important.
The upshot is that while it is possible that there might be some reverse causation of
education and sex of children this should not affect the estimates.

\clearpage
\newpage

\section{Sex Ratios at First Birth}

\setcounter{figure}{0}
\setcounter{table}{0}

% [WHY DO EVERYBODY - COMPARISON WITH JHA ET AL. - ESTABLISH THE ACTUAL SEX RATIO - 
% SHOW THAT LENGTH OF 1ST SPELL GENERALLY SHORT AND RELATION TO FECUNDITY]
% [what do the graphs show - representative woman - later use sample -
% regression results ]

I present a separate analysis of first spell births for three reasons.
First, previous research claims that the largest number of missing girls
is for first order births \citep{jha06}.
Secondly, there are substantially more first births than subsequent births, allowing
for a precise estimation of the ``natural'' percentage boys born in India if there
are no sex selection.
Finally, the results provide an indication of whether first spell length is
a good indicator for fecundity.

Figures \ref{fig:results_spell1_low}, \ref{fig:results_spell1_med}, and 
\ref{fig:results_spell1_high} show the predicted
percentage boys born by quarter from marriage to first birth and the associated
survival functions for the lowest, middle, and highest education groups for representative 
women.
For the first spell the representative woman is 16 years old at the beginning of the 
spell for the no education group, 17 years old for the middle education group, and 
20 years old for the high education group.
Each column represents a time period with the top panel showing urban results and 
the bottom panel rural results.
The graphs also show the expected natural rate of boys, approximately 51.2 percent.%
\footnote{
For comparison, if 55 percent of children born in a given quarter were boys,
approximately 14 percent of the female fetuses were aborted.
Assume 105 boys per 100 girls born, the expected natural sex ratio.
With $b$ boys,  $b\frac{100}{105}$ girls should be born.
If $g$ girls are observed the number of abortions is therefore $b\frac{100}{105}-g$
and the percent aborted female fetuses is 
$\frac{b\frac{100}{105}-g}{b\frac{100}{105}}\times 100$.
With 55 percent boys we get $\frac{55\frac{100}{105}-45}{55\frac{100}{105}}\times 100 = 14.09$.
The corresponding numbers for 60 percent and 65 percent boys are approximately 30 percent and
43 percent of the female fetuses aborted.
}

The most interesting result is how close to the natural sex ratio the predicted percentage
boys is for each group and for each period.
As Figure \ref{fig:results_spell1_low}(d) shows, for rural women without education before 1985, 
who also represent the biggest group, the predicted sex ratios align almost perfectly with 
the expected sex ratio.
For the other groups there is more volatility in the predicted percentage boys, but
nowhere is it statistically significantly larger than 51.2 percent.%
\footnote{
The urban no education group for the 1972--1984 period show two quarters where the predicted
percentage boys is just statistically significantly higher than 51.2 percent, but this
is likely due to recall error not perfectly caught by the method above
and the periods around those two quarters are below the natural percentage boys.
}
Furthermore, for quarters with more substantial deviations from the natural
sex ratio, the predictions are generally based on few births.
In other words, it appears that the probability of having a boy is exactly the same in 
India as it is in other places.

For the group most likely to use sex selection, highly
educated, urban women in the 1995--2006 period, the predicted percentage
boys is also almost perfectly aligned with the expected percentage boys, as shown
in Figure \ref{fig:results_spell1_high}(c).
Hence, there is no evidence that Hindus in India use sex selection on first births.
This cast serious doubts on the data used by \cite{jha06} and their results, as
also discussed by \cite{george06} and \cite{bhat06}.
% \footnote{%
% \cite{jha06} found that in both urban and rural areas there were 54.4 percent boys born 
% among first-borns.
% This translate into 0.17 to 0.23 million females missing among first-borns per year.
% The recall error problem is also noted by \cite{george06} and \cite{bhat06}.
% }

For all education groups and for all periods more than ninety percent of women had their 
first child within 21 quarters of being married and the proportion is increasing in education.
Furthermore, 70 to 85 percent of women will had their first child within ten quarters 
(2.5 years ) of their marriage and the average time between marriage and first birth has 
become shorter over time.
The most likely explanation for the reductions in duration and the increase 
in the number of women who have their first child before 21 quarters is improvements
in health status.
This is also consistent with the differences between education groups where more
educated women are healthier and therefore more likely to conceive.
There are two implications of this.
First, it reinforces the need for estimating the models separately for different
education levels.
Secondly, it confirms that the first spell length can serve as a suitable
proxy for fecundity and that Hindu women in India have their first birth very soon
after marriage, even among highly educated, urban women.

% \subsection{High Education Group}
% 
% If lower fertility is the driving force behind sex selection,
% women with the most education should be the earliest adopters and use it
% for lower parity children than those with less education.
% The predicted percentage boys for the second, third and fourth spells for 
% women with eight or more years of education show that this is, indeed, the case.
% As for the first spell, the results are presented using a representative woman
% with the same characteristics as above, except that the ages at the beginning of
% the spell are 22, 24 and 25 years for the second, third and fourth
% spell and that the first spell length is set equal to 16 months.
% 
% Figures \ref{fig:results_spell2_high_urban} and \ref{fig:results_spell2_high_rural}
% show the second spell predicted percentage boys and the survival curves by spell length 
% for urban and rural women.%
% \footnote{
% Recall, zero quarters is equivalent to nine months after the first birth.
% }
% For both figures, the top panel shows the results if the first child was a girl and the
% bottom panel the results if the first child was a boy.
% Not surprisingly, there is no evidence of sex selection for the 1972--1984 period;
% the predicted percentage boys match up closely with the natural.
% There is some evidence that spacing is shorter after the birth of a girl than
% after a boy, but the differences between the survival curves in the beginning of
% the spells are not large.
% The largest difference at five quarters is for urban women and that 
% is only around five percentage points.
% This is consistent with more educated women being more aware of the
% potential negative effects of close spacing.
% 
% 
% % SPELL 2 - URBAN
% 
% \begin{center}
% [Figure 4 about here.]
% \end{center}
% 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{First child a girl}
% % \setcounter{subfigure}{-2}
% % \subfloat[1972-1984 (N=2,791)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r4_pc}}\\
% % \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r4_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}} 
% % \setcounter{subfigure}{-1}
% % \subfloat[1985--1994 (N=5,011)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r4_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r4_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{0}
% % \subfloat[1995--2006 (N=3,827)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r4_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r4_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption*{First child a boy}
% % \setcounter{subfigure}{1}
% % \subfloat[1972-1984 (N=2,877)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r2_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r2_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % } 
% % \setcounter{subfigure}{2}
% % \subfloat[1985--1994 (N=5,421)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r2_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r2_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{3}
% % \subfloat[1995--2006 (N=4,015)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r2_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r2_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption{Predicted probability of having a boy and probability of
% % no birth yet from nine months after first birth for urban 
% % women with 8 or more years of education by quarter. 
% % Predictions based on age 22 at first birth.
% % Left column shows results prior to sex selection available, middle column before
% % sex selection illegal and right column after sex selection illegal.
% % N indicates the number of women in the relevant group in the underlying samples.
% % }
% % \label{fig:results_spell2_high_urban}
% % \end{figure}
% 
% 
% 
% % SPELL 2 - RURAL
% 
% \begin{center}
% [Figure 5 about here.]
% \end{center}
% 
% 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{First child a girl}
% % \setcounter{subfigure}{-2}
% % \subfloat[1972-1984 (N=1,183)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r3_pc}}\\
% % \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r3_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}} 
% % \setcounter{subfigure}{-1}
% % \subfloat[1985--1994 (N=2,967)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r3_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r3_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{0}
% % \subfloat[1995--2006 (N=2,740)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r3_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r3_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption*{First child a boy}
% % \setcounter{subfigure}{1}
% % \subfloat[1972-1984 (N=1,305)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r1_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r1_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % } 
% % \setcounter{subfigure}{2}
% % \subfloat[1985--1994 (N=3,213)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r1_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r1_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{3}
% % \subfloat[1995--2006 (N=2,945)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r1_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r1_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption{Predicted probability of having a boy and probability of
% % no birth yet from nine months after first birth for rural
% % women with 8 or more years of education by quarter. 
% % Predictions based on age 22 at first birth.
% % Left column shows results prior to sex selection available, middle column before
% % sex selection illegal and right column after sex selection illegal.
% % N indicates the number of women in the relevant group in the underlying samples.
% % }
% % \label{fig:results_spell2_high_rural}
% % \end{figure}
% 
% The first substantive evidence of sex selection is for urban women with one girl
% in the 1985--1994 period (Figure \ref{fig:results_spell2_high_urban}b).
% The percentage boys begins at the natural level but then increases to almost 60 
% percent after which it drops slightly.
% The use of sex-selective abortions becomes even more apparent for the 1995--2006
% period, where the percentage boys born to highly educated women with one girl 
% begins at just below 60 percent, increases to above 65 percent, followed by a decline to
% just over 55 percent.
% These results are especially remarkable given that the numbers are for all of India, not
% just the states with traditionally strong son preferences.
% 
% % [explain shape of curves]
% Of particularly interest is how the predicted percentage boys changes with spell length.
% The development in the percentage boys by quarter for the 1995--2006 period
% (Figure \ref{fig:results_spell2_high_urban}c), 
% is consistent with some women changing their 
% decision to use sex selection after going through one or more abortions.
% The results are for the median level of fecundity, and it is therefore unlikely 
% that the decline is due to lower fecundity for women who give birth later in the spell.%
% \footnote{
% It is, of course, possible that women find it harder to conceive during the
% second spell, but the survival curve would be very close to the survival 
% curve for the first period if the female fetuses were carried to term.
% }
% The 1985--1994 period (Figure \ref{fig:results_spell2_high_urban}b)
% shows the same fall at the end of the observed spell, but it also
% begins at the natural level, consistent with increasing access and 
% acceptance of sex selection over the early period.
% Women who began their second spell early in the 1985--1994 period were less likely
% to have access than women who began their second spell later in the period.
% % Furthermore, women who did not conceive early after a girl but
% % either decided to wait or had trouble getting pregnant were also more likely
% % to have sex selection techniques available to them.
% The 1985--1994 results can be thought of as the pattern that will
% prevail as sex selection is introduced, and the 1995--2006 results show
% the pattern with a mature and readily available technology.
% That sex selection for the 1995--2006 period is used earlier in the 
% spell compared to later also shows that not taking account of timing of births
% and the potential censoring of birth spell can lead to an upward bias in the
% final sex ratio of second born children.
% 
% % [compare to changes in rural areas for shape]
% For rural areas there is no sign of sex-selective abortion during the 1985--1994 period,
% but for the 1995--2006 period the percentage boys born to highly educated women who have
% a girl as their first child increases gradually from 55 percent to over 60 percent until
% dropping off to the normal percentage boys (Figure \ref{fig:results_spell2_high_rural}c).
% Hence, there appears to be easy access to prenatal sex determination even
% in rural areas.
% The difference between urban and rural women for the 1985--1994
% period is more likely to be the result of differences in fertility than in access
% to prenatal sex determinantion as shown below.
% As more and more rural women have only two children the use of sex selection
% goes up.
% The implication is that rural areas lag behind urban areas but not by 
% much and that the expected pattern for the next round of 
% the NFHS for rural educated women should closely mirror what is found 
% for the 1995--2006 period for urban women.
% 
% Women who had a boy as their first child do not appear to be using sex-selective abortions.
% This holds for urban and rural areas and for all time periods.
% In addition, women with a boy are less likely to have a second birth within the
% 21 quarters.
% This is especially prevalent for urban women, although rural women show the same pattern.
% For urban women with one boy, more than thirty percent are not predicted to have
% their second child within the 21 quarters covered here.
% This means that six years have passed since the birth of their first child and given
% the flatness of the survival function it it unlikely that many of these women will
% ever have a second child.
% Even for rural women around twenty percent of women will likely not have a second child.
% 
% 
% [THIS NEEDS TO BE REWRITTEN TO INCORPORATE R3'S QUESTION ABOUT PANELS (A) AND (D) AND
% THE HIGH SEX RATIO FOR THE 1972-1984 PERIOD]
% 
% The results for the third spell are shown in Figures \ref{fig:results_spell3_high_urban} 
% and \ref{fig:results_spell3_high_rural} for the representative 
% urban and rural women.
% The top panel shows the result if the first two children were girls, the middle
% panel if the couple had one boy and one girl, and the bottom panel if the
% first two children were boys.
% Again there is no evidence of sex-selective abortions before 1985, but
% clearly there is more volatility in the predicted percentage boys due
% to the lower sample sizes.
% 
% 
% % SPELL 3 - URBAN
% 
% \begin{center}
% [Figure 6 about here.]
% \end{center}
% 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{First two children girls}
% % \setcounter{subfigure}{-2}
% % \subfloat[1972-1984 (N=735)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r6_pc}}\\
% % \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r6_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}} 
% % \setcounter{subfigure}{-1}
% % \subfloat[1985--1994 (N=1,681)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r6_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r6_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{0}
% % \subfloat[1995--2006 (N=1,015)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r6_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r6_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption*{First two children one boy and one girl}
% % \setcounter{subfigure}{1}
% % \subfloat[1972-1984 (N=1,416)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r4_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r4_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % } 
% % \setcounter{subfigure}{2}
% % \subfloat[1985--1994 (N=3,506)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r4_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r4_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{3}
% % \subfloat[1995--2006 (N=2,367)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r4_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r4_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption{Predicted probability of having a boy and probability of
% % no birth yet from nine months after second birth for urban 
% % women with 8 or more years of education by quarter. 
% % Predictions based on age 24 at second birth.
% % Left column shows results prior to sex selection available, middle column before
% % sex selection illegal and right column after sex selection illegal.
% % N indicates the number of women in the relevant group in the underlying samples.
% % }
% % \label{fig:results_spell3_high_urban}
% % \end{figure}
% % 
% % 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{First two children boys}
% % \setcounter{subfigure}{4}
% % \subfloat[1972-1984 (N=789)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r2_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r2_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % } 
% % \setcounter{subfigure}{5}
% % \subfloat[1985--1994 (N=1,665)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r2_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r2_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{6}
% % \subfloat[1995--2006 (N=1,086)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r2_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r2_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % % \label{fig:results_spell3_high_urban}
% % \caption{(Continued) Predicted probability of having a boy and probability of
% % no birth yet from nine months after second birth for urban 
% % women with 8 or more years of education by quarter. 
% % Predictions based on age 24 at second birth.
% % Left column shows results prior to sex selection available, middle column before
% % sex selection illegal and right column after sex selection illegal.
% % N indicates the number of women in the relevant group in the underlying samples.
% % \ContinuedFloat
% % }
% % \addtocounter{figure}{1}
% % \end{figure}
% 
% % HAVE NO IDEA WHY THIS IS NEEDED BUT SOMETHING SEESM TO GO WRONG WITH THE COMBINATION
% % OF SUBFIGURE AND CONTINUEDFLOAT
% % 
% 
% % SPELL 3 - RURAL
% 
% \begin{center}
% [Figure 7 about here.]
% \end{center}
% 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{First two children girls}
% % \setcounter{subfigure}{-2}
% % \subfloat[1972-1984 (N=312)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r5_pc}}\\
% % \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r5_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}} 
% % \setcounter{subfigure}{-1}
% % \subfloat[1985--1994 (N=942)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r5_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r5_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{0}
% % \subfloat[1995--2006 (N=842)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r5_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r5_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption*{First two children one boy and one girl}
% % \setcounter{subfigure}{1}
% % \subfloat[1972-1984 (N=666)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r3_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r3_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % } 
% % \setcounter{subfigure}{2}
% % \subfloat[1985--1994 (N=1,909)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r3_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r3_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{3}
% % \subfloat[1995--2006 (N=1,763)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r3_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r3_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption{Predicted probability of having a boy and probability of
% % no birth yet from nine months after second birth for rural
% % women with 8 or more years of education by quarter. 
% % Predictions based on age 24 at second birth.
% % Left column shows results prior to sex selection available, middle column before
% % sex selection illegal and right column after sex selection illegal.
% % N indicates the number of women in the relevant group in the underlying samples.
% % }
% % \label{fig:results_spell3_high_rural}
% % \end{figure}
% % 
% % 
% % \begin{figure}[htpb]
% % % \TheContinuedFloat
% % \centering
% % \caption*{First two children boys}
% % \setcounter{subfigure}{4}
% % \subfloat[1972-1984 (N=339)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r1_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r1_s}} 
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % } 
% % \setcounter{subfigure}{5}
% % \subfloat[1985--1994 (N=899)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r1_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r1_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \setcounter{subfigure}{6}
% % \subfloat[1995--2006 (N=763)]{
% % \begin{minipage}{0.25\textwidth}
% % \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% % \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r1_pc}}\\
% % \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r1_s}}
% % \captionsetup[subfigure]{labelformat=parens}
% % \end{minipage}
% % }
% % \caption{(Continued) Predicted probability of having a boy and probability of
% % no birth yet from nine months after second birth for rural
% % women with 8 or more years of education by quarter. 
% % Predictions based on age 24 at second birth.
% % Left column shows results prior to sex selection available, middle column before
% % sex selection illegal and right column after sex selection illegal.
% % N indicates the number of women in the relevant group in the underlying samples.
% % \ContinuedFloat
% % }
% % % \label{fig:results_spell3_high_rural}
% % \addtocounter{figure}{1}
% % \end{figure}
% 
% % HAVE NO IDEA WHY THIS IS NEEDED BUT SOMETHING SEESM TO GO WRONG WITH THE COMBINATION
% % OF SUBFIGURE AND CONTINUEDFLOAT
% % \addtocounter{figure}{1}
% 
% 
% 
% In urban areas there is clear evidence of sex-selective abortions among women
% with two girls for both 1985--1994 and 1995--2006.
% For the first of these periods, births early in the spell are close to the natural sex ratio,
% but births after the fifth quarter are significantly above the natural level
% at around 65 percent boys.%
% \footnote{
% The sex ratio increases substantially for births after the 15th quarter, but 
% given the low number of births it is not clear how much one can learn from this increase.
% }
% Interestingly, there does not appear to be an increase in the use of sex-selective
% abortions in the 1995--2006 period compared to the 1985--1994 period.
% This can be explained by the increase in the use of sex-selective
% abortions for second births combined with lower desired fertility.
% Fewer women have two girls and given the apparent ready access to prenatal
% sex determination those who do are more likely to do so by choice.%
% \footnote{
% For urban women, 25 percent of the sample in the first period had two girls, 
% while only 22.7 percent of the third period sample had two girls.
% }
% Even then, the percentage boys born is between 60 and 65 percent.
% This pattern of sex selection for women with two girls is mirrored by rural women.
% As expected, given that most educated women do not want more than two children,
% there is no evidence for declining use of sex selection toward the end of the spell.
% 
% Women with either two boys or one boy and one girl as 
% their first two children do not appear to use sex selection.
% Relatively few of these women go on to have a third child
% and the proportion that does declines substantially over time.
% This is best illustrated by the urban women in the 1995--2006 period (Figures 
% \ref{fig:results_spell3_high_urban}c, f, i).
% Close to sixty percent of the women with two girls will have a third child,
% while only around thirty percent of those with at least one boy will
% have a third child.
% For comparison, during the 1972--1984 period eighty percent of the urban women 
% with two girls had a third child and almost sixty percent of those
% with at least one boy had a third child.
% The numbers are higher for rural women, but the pattern is the same.
% 
% 
% Relatively few women with eight or more years of education have three children and 
% even fewer have four.
% This means that caution should be used in interpreting the results and in the interest
% of space the graphs are not presented here.%
% \footnote{
% Graphs and underlying estimation results are available on request.
% }
% As for the other spells there is no evidence of sex selection for the first period.
% Again, families without boys are the main users of sex-selective abortions.
% For urban women with three girls the predicted percentage boys is between 60 and 75 for both
% the 1985--1994 and 1995--2006 time periods.
% The numbers for rural women with three girls are in line with those for urban women.
% It is worth noting that there are only 211 and 155 observed births for
% urban women for the two time periods, and only 167 and 178 births for the rural women.
% % What appear to be the only evidence against the ``funeral pyre'' hypothesis
% % comes from urban women with two girls and one boy for whom the predicted percentage
% % boys is above 65 percent for all quarters, but this result is based on only 292 births
% % and does not show up for other urban women or for rural women.
% Urban women with with two girls and one boy show a predicted percentage boys above 65 
% percent for all quarters, but this result is based on only 292 births.
% Finally, it is interesting that the probability of having a fourth child is
% higher for women with three boys than it is for women with two boys and one girl.
% This might be evidence of a slight gender balance preference, although 
% there is absolutely no evidence that women are aborting boys and the number
% of observed births to both groups is very small.
% 
% 
% 
% % SPELL 2 - URBAN
% % \begin{figure}
% % \centering
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r7_pc}}
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r8_pc}} \\ \setcounter{subfigure}{0}
% % \subfloat[Rural (N=4,663)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r7_s}}
% % \subfloat[Urban (N=7,940)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r8_s}} \\
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r3_pc}}
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r4_pc}} \\ \setcounter{subfigure}{2}
% % \subfloat[Rural (N=1,044)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r3_s}}
% % \subfloat[Urban (N=898)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r4_s}}
% % \caption{Predicted percent boys born and ``survival'' functions by spacing after first birth
% % and by areas of residence for women with 8 or more years of education with the first child a girl.
% % Top Panel shows Wants 2 or Fewer Children and Bottom Panel 3 or More. \protect\linebreak
% % N indicates Number of Women who Began a Spell.}
% % \label{fig:results_spell2_high_want}
% % \end{figure}
% 
% 
% % \begin{figure}[htpb]
% % \centering
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r7_pc}}
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r3_pc}}
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r8_pc}} 
% % \subfloat{\includegraphics[width=.25\textwidth]{want_spell2_high_r4_pc}} \\ \setcounter{subfigure}{0}
% % \subfloat[Rural (N=4,663)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r7_s}}
% % \subfloat[Rural (N=1,044)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r3_s}}
% % \subfloat[Urban (N=7,940)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r8_s}} 
% % \subfloat[Urban (N=898)]{\includegraphics[width=.25\textwidth]{want_spell2_high_r4_s}}
% % \caption{Predicted percent boys born and survival functions by spacing after first birth
% % and by areas of residence for women with 8 or more years of education with first child a girl.
% % Left shows wants 2 or fewer children and right 3 or more. 
% % N indicates number of women who began a spell.}
% % \label{fig:results_spell2_high_want}
% % \end{figure}
% 
% \subsection{Middle and Low Education Groups}
% 
% For women with one to seven years of education, a representative woman is 
% used with the first spell length set to 16 months and the ages at the 
% beginning of the spell 19, 21 and 24 for the second, third and fourth spells, respectively.
% There are two factors that make it difficult to establish whether sex selection
% is used among the middle education group.
% First, there are fewer observations for this group compared with the high
% education group.
% That problem becomes especially acute when looking at the higher spells
% divided by sex composition of previous children.
% Secondly, even for those spells with more information there is substantially
% more noise in the 1972--1984 results than for the high education group.%
% \footnote{
% The results are available on request.
% }
% % The results are presented in Appendix A.
% 
% For the second spell there is no evidence of sex-selective abortions for
% either urban or rural women and the proportion of women who have a second
% birth remains high across all three time periods.
% The same is the case for the third spell, although there are signs of falling
% fertility in that the proportion of women who have a third child is decreasing,
% especially among those with one or two boys.%
% \footnote{
% There are relatively few urban women with two girls in this education group 
% and although there might be some evidence of sex-selective abortions the
% results are too noisy to conclude anything with any degree of confidence.
% }
% Even for the fourth spell, women with three girls do not show statistically
% significant use of sex-selective abortions and this holds for both urban 
% and rural women.
% For both urban and rural women there has been a substantial decline in the
% number of women who have a fourth child and again this decline is concentrated
% among women with one or more boys.
% 
% 
% For women with no education, the starting ages are 18, 21 and 23 for 
% the second, third and fourth spells, respectively.
% Although there are not a large number of women in urban areas without education
% there are substantial numbers in the rural areas.%
% \footnote{
% The results are available on request.
% }
% % The results are presented in Appendix A.
% For urban women without education there is little evidence of sex selection
% for any of the spells.%
% \footnote{
% Urban women with three girls do have a higher percentage boys, but only 208 are in this group.
% }
% Although the percentage boys for rural women with one girl for the second
% spell is a very precisely estimated flat line for the 1985--1994 period,
% the percentage boys for the third period is between 55 and 60 from the
% seventh to the thirteenth quarter and this percentage is statistically 
% significant.\footnote{
% These quarters account for more than twenty percent of the births or
% slightly more than 700 births.
% }
% Two things makes this result puzzling.
% First, there is little evidence of women stopping
% childbearing once they have a son.
% Even among rural women with two sons almost ninety percent go on to
% have a third child and eighty percent of women with three boys
% have a fourth child.
% Hence, fertility remains very high for this group and therefore also
% the probability of having at least one son without the use of sex-selective 
% abortions.
% Secondly, one would expect more use of sex selection for women with
% two girls in the third spell or three girls in the fourth spell, but
% for both the predicted percentage boys is close to the natural rate.
%

\clearpage
\newpage

\section{Migration}

\setcounter{figure}{0}
\setcounter{table}{0}

\begin{table}[hp]
\centering
\footnotesize
\begin{threeparttable}
\caption{Migration by Beginning of Spell}
\label{tab:migrate}
\begin{tabular}{l  D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} } \toprule
                & \multicolumn{9}{c}{Years of Education} \\ \cmidrule(lr){2-10}
                & \multicolumn{3}{c}{None}                   & \multicolumn{3}{c}{1--7 Years}              & \multicolumn{3}{c}{8 or More Years} \\ \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(lr){8-10}
                & \mco{1972--} & \mco{1985--} & \mco{1995--} & \mco{1972--} & \mco{1985--} & \mco{1995--}  & \mco{1972--} & \mco{1985--} & \mco{1995--} \\
                & \mco{1984}   & \mco{1994}   & \mco{2006}   & \mco{1984}   & \mco{1994}   & \mco{2006}    & \mco{1984}   & \mco{1994}   & \mco{2006} \\
\midrule                
\multicolumn{10}{l}{\textbf{After marriage (\%)}} \\
Did not move    &       51.20&       59.60&       64.04&       50.24&       58.38&       63.68&       42.16&       51.07&       61.07\\
Within same type&       39.76&       32.69&       27.99&       34.85&       29.94&       25.95&       41.38&       34.69&       27.35\\
Rural to urban  &        7.20&        5.91&        5.87&       10.91&        8.18&        7.33&       11.91&        9.93&        7.75\\
Urban to rural  &        1.85&        1.80&        2.10&        4.00&        3.49&        3.04&        4.55&        4.30&        3.83\\
\addlinespace 
\multicolumn{10}{l}{\textbf{After first birth (\%)}} \\
Did not move    &       80.66&       86.63&       88.78&       75.16&       82.93&       88.32&       66.60&       76.85&       86.33\\
Within same type&       14.06&        9.23&        7.02&       16.11&       10.95&        7.12&       24.16&       16.21&        9.44\\
Rural to urban  &        4.25&        3.28&        3.23&        6.39&        4.33&        3.36&        6.63&        4.97&        3.04\\
Urban to rural  &        1.03&        0.85&        0.97&        2.34&        1.78&        1.21&        2.61&        1.96&        1.18\\
\addlinespace 
\multicolumn{10}{l}{\textbf{After second birth (\%)}} \\
Did not move    &       80.66&       85.40&       85.77&       75.17&       81.42&       85.08&       66.64&       74.58&       82.48\\
Within same type&       14.08&       10.12&        8.91&       16.15&       11.93&        9.33&       24.08&       17.59&       11.84\\
Rural to urban  &        4.22&        3.56&        4.02&        6.37&        4.72&        4.07&        6.72&        5.62&        4.09\\
Urban to rural  &        1.04&        0.93&        1.30&        2.32&        1.94&        1.51&        2.56&        2.20&        1.58\\
\addlinespace 
\multicolumn{10}{l}{\textbf{After third birth (\%)}} \\
Did not move    &       80.76&       84.22&       84.47&       75.19&       81.01&       82.24&       68.55&       75.25&       82.54\\
Within same type&       13.98&       10.89&        9.98&       15.88&       12.13&       11.51&       21.98&       16.62&       10.82\\
Rural to urban  &        4.25&        3.91&        4.28&        6.57&        4.93&        4.74&        7.02&        5.89&        4.98\\
Urban to rural  &        1.01&        0.98&        1.26&        2.35&        1.93&        1.51&        2.46&        2.24&        1.66\\
\bottomrule
\end{tabular}
\begin{tablenotes} 
\scriptsize
\item \hspace*{-0.5em} \textbf{Note.}
``Within same type'' indicates either a move from a rural area to another rural area,
or a move from an urban area to another urban area.
\end{tablenotes}
\end{threeparttable}
\end{table}

% \begin{table}[hp]
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Migration by beginning of spell for women with no education}
% \label{tab:migrate_low}
% \begin{tabular}{l  D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} } \toprule
%                 & \mco{1972--1984} & \mco{1985--1994} & \mco{1995--2006} \\
%                 & \multicolumn{3}{c}{After marriage (\%)} \\
% Did not move    &       51.20&       59.60&       64.04\\
% Within same type&       39.76&       32.69&       27.99\\
% Rural to urban  &        7.20&        5.91&        5.87\\
% Urban to rural  &        1.85&        1.80&        2.10\\
%                 & \multicolumn{3}{c}{After first birth (\%)} \\
% Did not move&       80.66&       86.63&       88.78\\
% Within same type&       14.06&        9.23&        7.02\\
% Rural to urban&        4.25&        3.28&        3.23\\
% Urban to rural&        1.03&        0.85&        0.97\\
%                 & \multicolumn{3}{c}{After second birth (\%)} \\
% Did not move&       80.66&       85.40&       85.77\\
% Within same type&       14.08&       10.12&        8.91\\
% Rural to urban&        4.22&        3.56&        4.02\\
% Urban to rural&        1.04&        0.93&        1.30\\
%                 & \multicolumn{3}{c}{After third birth (\%)} \\
% Did not move&       80.76&       84.22&       84.47\\
% Within same type&       13.98&       10.89&        9.98\\
% Rural to urban&        4.25&        3.91&        4.28\\
% Urban to rural&        1.01&        0.98&        1.26\\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} 
% \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.}
% ``Within same type'' indicates either a move from a rural area to another rural area,
% or a move from an urban area to another urban area.
% \end{tablenotes}
% \end{threeparttable}
% \end{table}
% 
% 
% \begin{table}
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Migration by beginning of spell for women with 1 through 7 years of education}
% \label{tab:migrate_med}
% \begin{tabular}{l  D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} } \toprule
%                 & \mco{1972--1984} & \mco{1985--1994} & \mco{1995--2006} \\
%                 & \multicolumn{3}{c}{After marriage (\%)} \\
% Did not move    &       50.24&       58.38&       63.68\\
% Within same type&       34.85&       29.94&       25.95\\
% Rural to urban  &       10.91&        8.18&        7.33\\
% Urban to rural  &        4.00&        3.49&        3.04\\
%                 & \multicolumn{3}{c}{After first birth (\%)} \\
% Did not move    &       75.16&       82.93&       88.32\\
% Within same type&       16.11&       10.95&        7.12\\
% Rural to urban  &        6.39&        4.33&        3.36\\
% Urban to rural  &        2.34&        1.78&        1.21\\
%                 & \multicolumn{3}{c}{After second birth (\%)} \\
% Did not move    &       75.17&       81.42&       85.08\\
% Within same type&       16.15&       11.93&        9.33\\
% Rural to urban  &        6.37&        4.72&        4.07\\
% Urban to rural  &        2.32&        1.94&        1.51\\
%                 & \multicolumn{3}{c}{After third birth (\%)} \\
% Did not move    &       75.19&       81.01&       82.24\\
% Within same type&       15.88&       12.13&       11.51\\
% Rural to urban  &        6.57&        4.93&        4.74\\
% Urban to rural  &        2.35&        1.93&        1.51\\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} 
% \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.}
% ``Within same type'' indicates either a move from a rural area to another rural area,
% or a move from an urban area to another urban area.
% \end{tablenotes}
% \end{threeparttable}
% \end{table}
% 
% 
% \begin{table}
% \centering
% \footnotesize
% \begin{threeparttable}
% \caption{Migration by beginning of spell for women with 8 or more years of education}
% \label{tab:migrate_high}
% \begin{tabular}{l  D{.}{.}{2,2} D{.}{.}{2,2} D{.}{.}{2,2} } \toprule
%                 & \mco{1972--1984} & \mco{1985--1994} & \mco{1995--2006} \\
%                 & \multicolumn{3}{c}{After marriage (\%)} \\
% Did not move    &       42.16&       51.07&       61.07\\
% Within same type&       41.38&       34.69&       27.35\\
% Rural to urban  &       11.91&        9.93&        7.75\\
% Urban to rural  &        4.55&        4.30&        3.83\\
%                 & \multicolumn{3}{c}{After first birth (\%)} \\
% Did not move    &       66.60&       76.85&       86.33\\
% Within same type&       24.16&       16.21&        9.44\\
% Rural to urban  &        6.63&        4.97&        3.04\\
% Urban to rural  &        2.61&        1.96&        1.18\\
%                 & \multicolumn{3}{c}{After second birth (\%)} \\
% Did not move    &       66.64&       74.58&       82.48\\
% Within same type&       24.08&       17.59&       11.84\\
% Rural to urban  &        6.72&        5.62&        4.09\\
% Urban to rural  &        2.56&        2.20&        1.58\\
%                 & \multicolumn{3}{c}{After third birth (\%)} \\
% Did not move    &       68.55&       75.25&       82.54\\
% Within same type&       21.98&       16.62&       10.82\\
% Rural to urban  &        7.02&        5.89&        4.98\\
% Urban to rural  &        2.46&        2.24&        1.66\\
% \bottomrule
% \end{tabular}
% \begin{tablenotes} 
% \scriptsize
% \item \hspace*{-0.5em} \textbf{Note.}
% ``Within same type'' indicates either a move from a rural area to another rural area,
% or a move from an urban area to another urban area.
% \end{tablenotes}
% \end{threeparttable}
% \end{table}

\clearpage
\newpage

\section{All Graphs using Main Sample}

\setcounter{figure}{0}
\setcounter{table}{0}

% Figures \ref{fig:results_spell2_low_urban} to \ref{fig:results_spell4_low_rural_cont}
% show the results for the group of women with no education.
% The starting ages are 18, 21 and 23 for the second, third and fourth spells,
% respectively.

% The results for women with one to seven years of education are presented
% in Figures \ref{fig:results_spell2_med_urban} to \ref{fig:results_spell4_med_rural_cont}.
% Again a representative woman is used with the first spell length set to
% 16 months and the ages at the beginning of the spell 19, 21 and 24
% for the second, third and fourth spells, respectively.

% [FIRST SPELL]

% \begin{figure}[htpb!]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=4,518)]{
% \begin{minipage}{0.28\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell1_g1_low_r2_pc}}\\
% \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell1_g1_low_r2_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}} 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=4,309)]{
% \begin{minipage}{0.28\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g2_low_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g2_low_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=1,762)]{
% \begin{minipage}{0.28\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g3_low_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g3_low_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=25,486)]{
% \begin{minipage}{0.28\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g1_low_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g1_low_r1_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=23,360)]{
% \begin{minipage}{0.28\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g2_low_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g2_low_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=7,922)]{
% \begin{minipage}{0.28\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g3_low_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g3_low_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with no education by quarter (3 month
% period). 
% Predictions based on age 16 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell1_low}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=3,666)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell1_g1_med_r2_pc}}\\
% \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell1_g1_med_r2_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}} 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=4,104)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g2_med_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g2_med_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=2,302)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g3_med_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g3_med_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=7,496)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g1_med_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g1_med_r1_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=9,000)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g2_med_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g2_med_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=5,208)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g3_med_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g3_med_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with 1 to 7 years of education by quarter (3 month
% period). 
% Predictions based on age 17 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell1_med}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=7,338)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell1_g1_high_r2_pc}}\\
% \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell1_g1_high_r2_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}} 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=11,045)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g2_high_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g2_high_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=9,116)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g3_high_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g3_high_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=3,933)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g1_high_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g1_high_r1_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=8,027)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g2_high_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g2_high_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=7,504)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell1_g3_high_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell1_g3_high_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with 8 or more years of education by quarter
% (3 month period). 
% Predictions based on age 20 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell1_high}
% \end{figure}
% 
% \clearpage
% \newpage

\subsection{Second Spell}

% SPELL 2 - URBAN - LOW
\begin{figure}[h]
\centering
\caption*{First child a girl}
\setcounter{subfigure}{-2}
\subfloat[1972-1984 (N=1,596)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_low_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985-1994 (N=2,232)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_low_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{0}
\subfloat[1995-2006 (N=952)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_low_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\caption*{First child a boy}
\setcounter{subfigure}{1}
\subfloat[1972-1984 (N=1,754)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_low_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_r2_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985-1994 (N=2,411)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_low_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995-2006 (N=930)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_low_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after first birth for urban 
women with no education by quarter (3 month period). 
Predictions based on age 18 at first birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell2_low_urban}
\end{figure}

% SPELL 2 - rural - LOW
\begin{figure}
\centering
\caption*{First child a girl}
\setcounter{subfigure}{-2}
\subfloat[1972-1984 (N=8,426)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_low_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985-1994 (N=11,481)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_low_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{0}
\subfloat[1995-2006 (N=3,686)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_low_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\caption*{First child a boy}
\setcounter{subfigure}{1}
\subfloat[1972-1984 (N=9,395)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_low_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_r1_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985-1994 (N=12,127)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_low_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995-2006 (N=3,860)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_low_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after first birth for rural 
women with no education by quarter (3 month period). 
Predictions based on age 18 at first birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell2_low_rural}
\end{figure}

% SPELL 2 - URBAN - MEDIUM
\begin{figure}
\centering
\caption*{First child a girl}
\setcounter{subfigure}{-2}
\subfloat[1972-1984 (N=1,342)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_med_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985-1994 (N=1,969)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_med_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{0}
\subfloat[1995-2006 (N=1,082)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_med_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\caption*{First child a boy}
\setcounter{subfigure}{1}
\subfloat[1972-1984 (N=1,461)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_med_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_r2_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985-1994 (N=2,182)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_med_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995-2006 (N=1,104)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_med_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after first birth for urban 
women with 1 to 7 years of education by quarter (3 month period). 
Predictions based on age 19 at first birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell2_med_urban}
\end{figure}

% SPELL 2 - rural - MEDIUM
\begin{figure}
\centering
\caption*{First child a girl}
\setcounter{subfigure}{-2}
\subfloat[1972-1984 (N=2,603)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_med_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985-1994 (N=4,095)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_med_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{0}
\subfloat[1995-2006 (N=2,159)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_med_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\caption*{First child a boy}
\setcounter{subfigure}{1}
\subfloat[1972-1984 (N=2,803)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_med_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_r1_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985-1994 (N=4,203)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_med_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995-2006 (N=2,197)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_med_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after first birth for rural 
women with 1 to 7 years of education by quarter (3 month period). 
Predictions based on age 19 at first birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell2_med_rural}
\end{figure}

% High education

\begin{figure}[htpb]
\centering
\caption*{First child a girl}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=2,689)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r4_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=4,869)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=3,774)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First child a boy}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=2,806)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r2_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=5,246)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=3,969)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after first birth for urban 
women with 8 or more years of education by quarter (3 month period). 
Predictions based on age 22 at first birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell2_high_urban}
\end{figure}

\begin{figure}[htpb]
\centering
\caption*{First child a girl}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=1,305)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r3_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=3,105)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=2,823)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First child a boy}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=1,404)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g1_high_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_r1_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=3,381)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g2_high_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=3,044)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell2_g3_high_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after first birth for rural
women with 8 or more years of education by quarter (3 month period). 
Predictions based on age 22 at first birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell2_high_rural}
\end{figure}

\clearpage
\newpage

\subsection{Third Spell}

% low education

\begin{figure}[htpb]
\centering
\caption*{First two children girls}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=493)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_low_r6_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_r6_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=996)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_low_r6_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_r6_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=475)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_low_r6_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_r6_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First two children one boy and one girl}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=1,083)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_low_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=2,141)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_low_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=950)]{
    \begin{minipage}{0.26\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_low_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for urban 
women with no of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell3_low_urban}
\end{figure}

\begin{figure}[htpb]
\centering
\ContinuedFloat
\caption*{First two children boys}
\setcounter{subfigure}{4}
\subfloat[1972--1984 (N=569)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_low_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_r2_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{5}
\subfloat[1985--1994 (N=1,124)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_low_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[1995--2006 (N=426)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_low_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
% \label{fig:results_spell3_low_urban}
\caption{(Continued) Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for urban 
women with no of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\end{figure}

\begin{figure}[htpb]
\centering
\caption*{First two children girls}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=2,599)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_low_r5_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_r5_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=5,062)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_low_r5_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_r5_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=1,879)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_low_r5_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_r5_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First two children one boy and one girl}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=5,382)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_low_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=10,473)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_low_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=3,779)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_low_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for rural
women with no of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell3_low_rural}
\end{figure}

\begin{figure}[htpb]
\ContinuedFloat
\centering
\caption*{First two children boys}
\setcounter{subfigure}{4}
\subfloat[1972--1984 (N=3,070)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_low_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_r1_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{5}
\subfloat[1985--1994 (N=5,339)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_low_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[1995--2006 (N=1,846)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_low_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{(Continued) Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for rural
women with no of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
% \label{fig:results_spell3_low_rural}
\end{figure}

% Medium education

\begin{figure}[htpb]
\centering
\caption*{First two children girls}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=432)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_med_r6_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_r6_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=825)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_med_r6_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_r6_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=425)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_med_r6_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_r6_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First two children one boy and one girl}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=883)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_med_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=1,673)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_med_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=848)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_med_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for urban 
women with 1 to 7 years of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell3_med_urban}
\end{figure}

\begin{figure}[htpb]
\centering
\ContinuedFloat
\caption*{First two children boys}
\setcounter{subfigure}{4}
\subfloat[1972--1984 (N=446)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_med_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_r2_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{5}
\subfloat[1985--1994 (N=848)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_med_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[1995--2006 (N=438)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_med_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
% \label{fig:results_spell3_med_urban}
\caption{(Continued) Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for urban 
women with 1 to 7 years of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\end{figure}

\begin{figure}[htpb]
\centering
\caption*{First two children girls}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=775)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_med_r5_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_r5_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=1,659)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_med_r5_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_r5_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=848)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_med_r5_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_r5_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First two children one boy and one girl}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=1,639)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_med_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=3,222)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_med_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=1,773)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_med_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for rural
women with 1 to 7 years of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell3_med_rural}
\end{figure}

\begin{figure}[htpb]
\ContinuedFloat
\centering
\caption*{First two children boys}
\setcounter{subfigure}{4}
\subfloat[1972--1984 (N=785)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_med_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_r1_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{5}
\subfloat[1985--1994 (N=1,548)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_med_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[1995--2006 (N=730)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_med_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{(Continued) Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for rural
women with 1 to 7 years of education by quarter (3 month period). 
Predictions based on age 21 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
% \label{fig:results_spell3_med_rural}
\end{figure}

% High education

\begin{figure}[htpb]
\centering
\caption*{First two children girls}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=709)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r6_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r6_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=1,654)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r6_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r6_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=1,000)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r6_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r6_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First two children one boy and one girl}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=1,409)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r4_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=3,460)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=2,357)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r4_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r4_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for urban 
women with 8 or more years of education by quarter (3 month period). 
Predictions based on age 24 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell3_high_urban}
\end{figure}

\begin{figure}[htpb]
\centering
\ContinuedFloat
\caption*{First two children boys}
\setcounter{subfigure}{4}
\subfloat[1972--1984 (N=780)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r2_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{5}
\subfloat[1985--1994 (N=1,660)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[1995--2006 (N=1,098)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r2_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r2_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
% \label{fig:results_spell3_high_urban}
\caption{(Continued) Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for urban 
women with 8 or more years of education by quarter (3 month period). 
Predictions based on age 24 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\end{figure}

\begin{figure}[htpb]
\centering
\caption*{First two children girls}
\setcounter{subfigure}{-2}
\subfloat[1972--1984 (N=348)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r5_pc}}\\
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r5_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-1}
\subfloat[1985--1994 (N=982)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r5_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r5_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{0}
\subfloat[1995--2006 (N=863)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r5_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r5_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{First two children one boy and one girl}
\setcounter{subfigure}{1}
\subfloat[1972--1984 (N=700)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r3_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{2}
\subfloat[1985--1994 (N=1,987)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{3}
\subfloat[1995--2006 (N=1,819)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r3_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r3_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for rural
women with 8 or more years of education by quarter (3 month period). 
Predictions based on age 24 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
\label{fig:results_spell3_high_rural}
\end{figure}

\begin{figure}[htpb]
\ContinuedFloat
\centering
\caption*{First two children boys}
\setcounter{subfigure}{4}
\subfloat[1972--1984 (N=370)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g1_high_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_r1_s}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{5}
\subfloat[1985--1994 (N=945)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g2_high_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[1995--2006 (N=790)]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell3_g3_high_r1_pc}}\\
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_r1_s}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{(Continued) Predicted probability of having a boy and probability of
no birth yet from nine months after second birth for rural
women with 8 or more years of education by quarter (3 month period). 
Predictions based on age 24 at second birth.
Left column shows results prior to sex selection available, middle column before
sex selection illegal and right column after sex selection illegal.
N indicates the number of women in the relevant group in the underlying samples.
}
% \label{fig:results_spell3_high_rural}
\end{figure}

\clearpage
\newpage

% \subsection{Fourth Spell}
% 
% % low education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First three children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=130)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r8_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r8_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=446)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r8_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r8_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=200)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r8_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r8_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children one boy and two girls}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=410)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=1,185)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=601)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell4_low_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \ContinuedFloat
% \caption*{First three children two boys and one girl}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=389)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=1,145)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=486)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children three boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=141)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=382)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=155)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% % \label{fig:results_spell4_low_urban}
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First three children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=752)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r7_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r7_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=1,954)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r7_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r7_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=888)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r7_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r7_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children one boy and two girls}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=2,065)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=5,990)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=2,456)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell4_low_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \ContinuedFloat
% \centering
% \caption*{First three children two boys and one girl}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=2,168)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=5,439)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=2,155)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=777)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=1,895)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=696)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% % \label{fig:results_spell4_low_rural}
% \end{figure}
% 
% 
% 
% % Medium education
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First three children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=110)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r8_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r8_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=282)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r8_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r8_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=144)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r8_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r8_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children one boy and two girls}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=299)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=765)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=372)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell4_med_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \ContinuedFloat
% \caption*{First three children two boys and one girl}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=274)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=600)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=275)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children three boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=77)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=230)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=90)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% % \label{fig:results_spell4_med_urban}
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First three children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=210)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r7_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r7_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=601)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r7_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r7_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=317)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r7_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r7_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children one boy and two girls}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=581)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=1,532)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=800)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell4_med_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \ContinuedFloat
% \centering
% \caption*{First three children two boys and one girl}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=492)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=1,197)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=578)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children three boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=158)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=357)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=192)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% % \label{fig:results_spell4_med_rural}
% \end{figure}
% 
% 
% 
% 
% 
% % High education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First three children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=131)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r8_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r8_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=355)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r8_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r8_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=191)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r8_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r8_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children one boy and two girls}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=353)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=1,062)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=590)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell4_high_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \ContinuedFloat
% \caption*{First three children two boys and one girl}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=309)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=681)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=344)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children three boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=87)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=233)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=110)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% % \label{fig:results_spell4_high_urban}
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First three children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=62)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r7_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r7_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=257)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r7_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r7_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=201)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r7_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r7_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children one boy and two girls}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=176)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=664)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=564)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_spell4_high_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \ContinuedFloat
% \centering
% \caption*{First three children two boys and one girl}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=173)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=515)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=367)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First three children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=56)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g1_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=166)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g2_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=92)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{spell4_g3_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% % \label{fig:results_spell4_high_rural}
% \end{figure}
% 
% 
% \clearpage

% 
% \section{Graphs using Women Married less than 19 Years}
% 
% \setcounter{figure}{0}
% \setcounter{table}{0}
% 
% 
% % Figures \ref{fig:results_spell2_low_urban} to \ref{fig:results_spell4_low_rural_cont}
% % show the results for the group of women with no education.
% % The starting ages are 18, 21 and 23 for the second, third and fourth spells,
% % respectively.
% 
% % The results for women with one to seven years of education are presented
% % in Figures \ref{fig:results_spell2_med_urban} to \ref{fig:results_spell4_med_rural_cont}.
% % Again a representative woman is used with the first spell length set to
% % 16 months and the ages at the beginning of the spell 19, 21 and 24
% % for the second, third and fourth spells, respectively.
% 
% 
% % [FIRST SPELL]
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=2,529)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g1_low_r2_pc}}\\
% \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g1_low_r2_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}} 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=3,831)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g2_low_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g2_low_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=1,762)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g3_low_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g3_low_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=16,293)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g1_low_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g1_low_r1_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=21,631)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g2_low_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g2_low_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=7,922)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g3_low_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g3_low_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with no education by quarter (3 month
% period). 
% Predictions based on age 16 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell1_low}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=2,022)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g1_med_r2_pc}}\\
% \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g1_med_r2_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}} 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=3,653)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g2_med_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g2_med_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=2,302)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g3_med_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g3_med_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=4,575)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g1_med_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g1_med_r1_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=8,356)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g2_med_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g2_med_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=5,208)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g3_med_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g3_med_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with 1 to 7 years of education by quarter (3 month
% period). 
% Predictions based on age 17 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell1_med}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=4,408)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g1_high_r2_pc}}\\
% \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g1_high_r2_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}} 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=10,118)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g2_high_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g2_high_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=9,116)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g3_high_r2_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g3_high_r2_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=2,668)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g1_high_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g1_high_r1_s}} 
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=7,585)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g2_high_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g2_high_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=7,504)]{
% \begin{minipage}{0.31\textwidth}
% \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
% \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell1_g3_high_r1_pc}}\\
% \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell1_g3_high_r1_s}}
% \captionsetup[subfigure]{labelformat=parens}
% \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with 8 or more years of education by quarter
% (3 month period). 
% Predictions based on age 20 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell1_high}
% \end{figure}
% 
% \clearpage
% 
% \subsection{Second Spell}
% 
% 
% % SPELL 2 - URBAN - LOW
% \begin{figure}[hp]
% \centering
% \caption*{First child a girl}
% \setcounter{subfigure}{-2}
% \subfloat[1972-1984 (N=861)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985-1994 (N=1,763)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[1995-2006 (N=946)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption*{First child a boy}
% \setcounter{subfigure}{1}
% \subfloat[1972-1984 (N=951)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985-1994 (N=1,929)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995-2006 (N=921)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for urban 
% women with no education by quarter (3 month period). 
% Predictions based on age 18 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell2_low_urban}
% \end{figure}
% 
% 
% % SPELL 2 - rural - LOW
% \begin{figure}
% \centering
% \caption*{First child a girl}
% \setcounter{subfigure}{-2}
% \subfloat[1972-1984 (N=5,103)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985-1994 (N=9,913)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[1995-2006 (N=3,653)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption*{First child a boy}
% \setcounter{subfigure}{1}
% \subfloat[1972-1984 (N=5,500)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_low_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985-1994 (N=10,437)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_low_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995-2006 (N=3,834)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_low_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for rural 
% women with no education by quarter (3 month period). 
% Predictions based on age 18 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell2_low_rural}
% \end{figure}
% 
% 
% % SPELL 2 - URBAN - MEDIUM
% \begin{figure}
% \centering
% \caption*{First child a girl}
% \setcounter{subfigure}{-2}
% \subfloat[1972-1984 (N=714)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985-1994 (N=1,578)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[1995-2006 (N=1,074)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption*{First child a boy}
% \setcounter{subfigure}{1}
% \subfloat[1972-1984 (N=791)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985-1994 (N=1,766)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995-2006 (N=1,096)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 19 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell2_med_urban}
% \end{figure}
% 
% 
% % SPELL 2 - rural - MEDIUM
% \begin{figure}
% \centering
% \caption*{First child a girl}
% \setcounter{subfigure}{-2}
% \subfloat[1972-1984 (N=1,464)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985-1994 (N=3,510)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[1995-2006 (N=2,151)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption*{First child a boy}
% \setcounter{subfigure}{1}
% \subfloat[1972-1984 (N=1,645)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_med_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985-1994 (N=3,671)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_med_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995-2006 (N=2,182)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_med_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for rural 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 19 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell2_med_rural}
% \end{figure}
% 
% % High education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First child a girl}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=1,576)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=4,105)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=3,762)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First child a boy}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=1,667)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=4,431)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=3,953)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 22 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell2_high_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First child a girl}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=847)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=2,788)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=2,816)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First child a boy}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=925)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g1_high_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=3,033)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g2_high_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=3,039)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell2_g3_high_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 22 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell2_high_rural}
% \end{figure}
% 
% \clearpage
% \subsection{Third Spell}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First two children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=252)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r6_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=690)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=458)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First two children one boy and one girl}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=564)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=1,497)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=915)]{
%     \begin{minipage}{0.26\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell3_low_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \ContinuedFloat
% \caption*{First two children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=293)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=783)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=410)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% % \label{fig:results_limit_spell3_low_urban}
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First two children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=1,417)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r5_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=4,060)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=1,827)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First two children one boy and one girl}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=2,803)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=8,265)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=3,661)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell3_low_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \ContinuedFloat
% \centering
% \caption*{First two children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=1,613)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_low_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=4,200)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_low_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=1,786)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_low_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% % \label{fig:results_limit_spell3_low_rural}
% \end{figure}
% 
% 
% 
% % Medium education
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First two children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=223)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r6_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=577)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=409)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First two children one boy and one girl}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=433)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=1,188)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=804)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell3_med_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \ContinuedFloat
% \caption*{First two children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=229)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=594)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=423)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% % \label{fig:results_limit_spell3_med_urban}
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First two children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=395)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r5_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=1,264)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=840)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First two children one boy and one girl}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=829)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=2,549)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=1,737)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell3_med_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \ContinuedFloat
% \centering
% \caption*{First two children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=419)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_med_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=1,216)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_med_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=716)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_med_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% % \label{fig:results_limit_spell3_med_rural}
% \end{figure}
% 
% 
% % High education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First two children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=390)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r6_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=1,196)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=981)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r6_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r6_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First two children one boy and one girl}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=778)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=2,543)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=2,309)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r4_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell3_high_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \ContinuedFloat
% \caption*{First two children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=442)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=1,230)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=1,068)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r2_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% % \label{fig:results_limit_spell3_high_urban}
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{First two children girls}
% \setcounter{subfigure}{-2}
% \subfloat[1972--1984 (N=200)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r5_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[1985--1994 (N=798)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{0}
% \subfloat[1995--2006 (N=857)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r5_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r5_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{First two children one boy and one girl}
% \setcounter{subfigure}{1}
% \subfloat[1972--1984 (N=412)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{2}
% \subfloat[1985--1994 (N=1,646)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{3}
% \subfloat[1995--2006 (N=1,805)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r3_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_limit_spell3_high_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \ContinuedFloat
% \centering
% \caption*{First two children boys}
% \setcounter{subfigure}{4}
% \subfloat[1972--1984 (N=209)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g1_high_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{5}
% \subfloat[1985--1994 (N=767)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g2_high_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[1995--2006 (N=782)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{limit_spell3_g3_high_r1_s}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{(Continued) Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% % \label{fig:results_limit_spell3_high_rural}
% \end{figure}
% 
% 
% 
% \clearpage
% 
% 
% \section{Graphs using Women Dropped Because of Recall Error}
% 
% \setcounter{figure}{0}
% \setcounter{table}{0}
% 
% 
% % [FIRST SPELL]
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1950--1980}
% \setcounter{subfigure}{-2}
% \subfloat[Urban (N=3,775)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell1_g1_low_r2_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell1_g1_low_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[Rural (N=19,635)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell1_g1_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell1_g1_low_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with no education by quarter (3 month
% period). 
% Predictions based on age 16 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell1_low}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1954--1980 }
% \setcounter{subfigure}{-2}
% \subfloat[Urban (N=2,864)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell1_g1_med_r2_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell1_g1_med_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[Rural (N=4,885)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell1_g1_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell1_g1_med_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with 1 to 7 years of education by quarter (3 month
% period). 
% Predictions based on age 17 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell1_med}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1955--1980}
% \setcounter{subfigure}{-2}
% \subfloat[Urban (N=3,157)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell1_g1_high_r2_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell1_g1_high_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[Rural (N=1,491)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell1_g1_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell1_g1_high_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from time of marriage for women with 8 or more years of education by quarter
% (3 month period). 
% Predictions based on age 20 at marriage.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell1_high}
% \end{figure}
% 
% \clearpage
% 
% \subsection{Second Spell}
% 
% 
% % SPELL 2 - URBAN - LOW
% \begin{figure}[hb]
% \centering
% \caption*{1954--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[First child a girl (N=1,760)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[First child a boy (N=2,050)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for urban 
% women with no education by quarter (3 month period). 
% Predictions based on age 18 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell2_low_urban}
% \end{figure}
% 
% 
% % SPELL 2 - rural - LOW
% \begin{figure}
% \centering
% \caption*{1954--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[First child a girl (N=8,594)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[First child a boy (N=10,123)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_low_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for rural 
% women with no education by quarter (3 month period). 
% Predictions based on age 18 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell2_low_rural}
% \end{figure}
% 
% 
% % SPELL 2 - URBAN - MEDIUM
% \begin{figure}
% \centering
% \caption*{1955--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[First child a girl (N=1,362)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[First child a boy (N=1,472)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 19 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell2_med_urban}
% \end{figure}
% 
% 
% % SPELL 2 - rural - MEDIUM
% \begin{figure}
% \centering
% \caption*{1955--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[First child a girl (N=2,254)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[First child a boy (N=2,368)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_med_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for rural 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 19 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell2_med_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1957--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[First child a girl (N=1,498)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r4_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[First child a boy (N=1,607)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 22 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell2_high_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1957--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[First child a girl (N=655)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r3_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[First child a boy (N=743)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell2_g1_high_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after first birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 22 at first birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell2_high_rural}
% \end{figure}
% 
% \clearpage
% 
% 
% \subsection{Third Spell}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1957--1984}
% \setcounter{subfigure}{-2}
% \subfloat[Two girls (N=763)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r6_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[One boy and one girl (N=1,743)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[Two boys (N=982)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell3_low_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1957--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[Two girls (N=3,796)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r5_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[One boy and one girl (N=8,390)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[1972--1984 (N=4,903)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_low_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with no of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell3_low_rural}
% \end{figure}
% 
% 
% % Medium education
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1958--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[Two girls (N=633)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r6_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[One boy and one girl (N=1,236)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[Two boys (N=716)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell3_med_urban}
% \end{figure}
% 
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1958--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[Two girls (N=995)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r5_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[One boy and one girl (N=2,037)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[Two boys (N=1,093)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_med_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 1 to 7 years of education by quarter (3 month period). 
% Predictions based on age 21 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell3_med_rural}
% \end{figure}
% 
% 
% % High education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1960--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[Two girls (N=660)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r6_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r6_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[One boy and one girl (N=1,375)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r4_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r4_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[Two boys (N=673)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r2_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r2_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for urban 
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell3_high_urban}
% \end{figure}
% 
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{1960--1984 }
% \setcounter{subfigure}{-2}
% \subfloat[Two girls (N=267)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r5_pc}}\\
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r5_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-1}
% \subfloat[One boy and one girl (N=573)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r3_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r3_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{0}
% \subfloat[Two boys (N=341)]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. boy (\%)]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r1_pc}}\\
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{dropped_spell3_g1_high_r1_s}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \caption{Predicted probability of having a boy and probability of
% no birth yet from nine months after second birth for rural
% women with 8 or more years of education by quarter (3 month period). 
% Predictions based on age 24 at second birth.
% Left column shows results prior to sex selection available, middle column before
% sex selection illegal and right column after sex selection illegal.
% N indicates the number of women in the relevant group in the underlying samples.
% }
% \label{fig:results_dropped_spell3_high_rural}
% \end{figure}

\clearpage

\bibliographystyle{aer}
% \bibliography{econ}
% \bibliography{library}
\bibliography{collection}

\addcontentsline{toc}{section}{References}

\end{document}