### 2019-02-06 paper/birth_spacing_lit.tex
Initial commit of possible lit and write-up

To read:

In birth_spacing:

In collection.bib
\citep{Rafalitnanana2000} 
\citep{Rahman1993}
\citep{Jayachandran2017}
\citep{Jayachandran2017a}
\citep{Lambert2016}

In collection.bib but maybe not that relevant
\citep{Basu2010} 
\citep{dasgupta95} 
\citep{Kugler2017}