### 2019-11-27 paper/conceptual_framework.tex
Edit and shorten conceptual framework
Below 2,000 words now. Still seems too long.

Second, the paper lacks a conceptual framework to organize (and
motivate) the analysis. Following on the comments of one reviewer, one
suggestion would be to focus on the role of women’s education, with a
more thorough description of the context of changing education in India
over the last half century and clearly lay out a framework for the
mechanisms through which birth spacing is affected by educational
expansion.

\section{Birth Spacing and Education: Mechanisms and Findings}

This section provides a conceptual framework for understanding what determines  
birth spacing.
I first discuss the main determinants of spacing examined in the prior research, and 
then turn to how son preference and sex selection fits into the decisions on spacing.
Since sex selection and education are closely linked, I pay special attention to how 
education affects spacing.

In the simplest possible dynamic model of fertility, the spacing decision is trivial.
Assume that parents derive benefits from the presence of children, rather than from 
the births themselves, that spacing of children by itself does impact the parental utility, 
and that the parents perfectly control when to have children. 
The optimal decision is then to have the children as early as possible 
\citep{Newman1988}.
The simplest model does not fit the observed spacing patterns, which suggests 
that one or more of the assumptions are wrong.

First, parents do not entirely control the timing of births, but rather decide on 
the levels of sexual activity and contraceptive effort, which, in turn, affect the 
likelihood of conceiving, while taking into account fecundity and their knowledge 
of and access to contraceptive technology.

Fecundity determinants, such as age and health of the mother, are, however, unlikely 
to be significant factors in how spacing changed over time or across education levels 
in India for the period covered.
Outside of direct famine conditions, women's health and nutritional status do not 
appear to affect spacing substantially \citep{Huffman1987,John1987,lindstrom99}.
The Maharashtra drought occurred in 1970--1973 and would, therefore, only impact the first
year or two of the data.
Furthermore, I only examine spacing up to parity fourth, and the age of the mother is,
therefore, unlikely to be a concern when it comes to fecundity, even though 
better-educated women are, on average, older when they marry.

Despite a general belief that better access to and knowledge of contraception 
should lead to longer average spacing, the empirical relationship between contraception 
use and birth spacing is ambiguous 
\citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Yeakey2009,Kim2010,Soest2018}.
This lack of evidence is consistent with the ambiguous theoretical effect of introducing 
modern contraceptives on birth spacing.
On the one hand, in the absence of high-efficacy contraceptive technologies, parents may try to 
avoid overshooting their desired number of children by spacing their children further apart
than they would otherwise do.
Increased reliability of the access to and effectiveness of contraceptives then lead to 
shorter birth intervals because parents are more confident they can effectively
stop childbearing when they want \citep{Keyfitz1971,Heckman1976}.
On the other hand, when contraceptives are ineffective, some birth intervals will
be shorter than ideal, in the way discussed below, through simple random variation.
With increased reliability of contraceptive efforts, parents can better avoid too short 
birth spacing, and average spacing should go up.

To further complicate the relationship between contraception and spacing, the ability 
to successfully use ``low efficacy'' contraceptive methods, such as the rhythm method, 
is increasing in education, while there is no difference across education levels 
when it comes to modern contraceptives such as pills or injectables \citep{Rosenzweig1989}.
Additionally, even in the absence of modern contraceptives, families appear able to 
time or at least postpone births when they have substantial enough incentive 
\citep{Jayachandran2011,Alam2018}.

Second, parents may care about the spacing between birth because of the 
potential effects of spacing on the children.%
\footnote{
In principle, the mother could also be affected, but the estimated impact on 
mothers' anthropometric status is mixed, and there appears to be no 
effect on mortality 
\citep{Ronsmans1998,Menken2003,Dewey2007,Conde-Agudelo2012}.
}
One theory is that child ``quality'' increases with spacing 
since parental attention per child increases
\citep{Zajonc1975,Zajonc1976,Razin1980}.
In this case, birth spacing should increase in family income, while the 
effect of maternal education is ambiguous since higher education, on the 
one hand, makes the mother's time spent out of the labor market more 
costly and on the other hand increases how productive the mother is 
in childrearing.
The evidence on spacing's effect on child quality measures such as IQ 
and education is, however, mixed for developed countries and non-existing 
for developing countries
\citep{Powell1993,Pettersson-Lidbom2009,Buckles2012,Barclay2017}.

For one crucial aspect of child quality---health and mortality---shorter spacing 
does lead to worse outcomes, especially for very short intervals of 24 months 
or less, although this effect decreases with maternal education
\citep{Whitworth2002,Conde-Agudelo2006,Conde-Agudelo2012,Molitoris2019}.%
\footnote{
The underlying mortality risk could also affect birth spacing, but
there is little empirical evidence for this 
\citep{Newman1983,Newman1988,Bhalotra2008}.
}
If parents realize short birth spacing increases their children's mortality risk, 
they likely prefer to space births further apart.
However, this effect should diminish with education level since women with 
secondary or tertiary education experience relatively low mortality risk for 
their children, even with very short spacing is associated.

Third, parents incur both direct and time costs when they have children, 
and economic factors can, therefore, affect birth spacing \citep{Hotz1997,Schultz1997}.
I divide the economic factors into the effects of price and income shocks,
the expected income trajectory of the husband and wife,
and economies of scale considerations.

The expected effect of economies of scale in childrearing, which, for example, 
can arise because taking care of two young children does not take twice as much 
time as taking care of one child, is to shorten intervals between births
\citep{Vijverberg1982,Espenshade1984}.
To the extent that children require the mother to do less market work, parents 
can lower the cost of children by reducing birth spacing.
In this case, better-educated women will both have lower fertility and shorter
spacing than less-educated women because of their higher opportunity cost of time 
\citep{Ross1974,Newman1981,Newman1984}.
However, if the mother mainly works at home or as household income increases, there 
is less incentive to space children for this reason.

In the absence of perfect capital markets, theory predicts that the steeper 
the parents' lifetime income profile, the longer the spacing between births, 
which suggests that with economic development we should see a general increase 
in spacing across education groups and that this increase in spacing should be 
stronger with increasing education levels
\citep{Heckman1976,Wolpin1984,Newman1988}.
The idea is that although parents would prefer to have their children as early 
as possible to enjoy them for longer, there is also a direct cost to children.
Early on income is low, which means that the marginal utility of consumption is 
higher, making parents less willing to give up consumption now to have a child, 
compare to later when they know that their income will be higher 
\citep{Newman1984,Happel1984}.

The income profile argument assumes that parents know their future income, 
but income and price shocks may also affect spacing. 
A temporary, higher cost of women's time or a lower household income, will 
temporarily decrease fertility and, thereby, increase birth spacing 
\citep{Moffitt1984,Hotz1988,Portner2001,Alam2018}.
Increasing education likely reduces the risk of income shocks, but the 
effect on other types of shocks that might impact the spacing is unclear.

Finally, the discussion has so far assumed that children are identical---except 
for any differences that arise from spacing---but there is substantial evidence 
that the sex composition of previous children affects birth spacing, with the
absence of boys significantly decreasing the duration to next birth
\citep{Haughton1995,Haughton1996,Rahman1993,Bhalotra2008,Kumar2016,Soest2018}.
There is also evidence of differential stopping behavior, with an additional child 
more likely the fewer sons a family has 
\citep{repetto72,Das1987,Arnold1997,arnold98,clark00,Basu2010,Barcellos2014}.

It is not, however, immediately apparent that we should expect shorter spacing in 
the absence of boys, if short spacing lowers the survival probability of the 
next child, as discussed above.
There are two potential reasons why parents might still try to conceive sooner 
in the absence of boys.%
\footnote{
The significant differences in spacing by sex composition indicate that
parents control the timing of birth, at least to some extent.
In Matlab, the effect of son preference on birth spacing is stronger in 
areas with better access to family planning \citep{Rahman1993}.
}
One possibility is that parents do not realize that short spacing is potentially
detrimental to the next child's health, resulting in high mortality for both boys 
and girls born after short spacing across all households whether or not they have 
a son preference.%
\footnote{
We may even see a higher likelihood of dying for sons than girls, given that boys 
have naturally higher mortality.
}

The other possibility is that parents realize that short spacing has adverse 
health effects, but their son-preference is so strong that they are willing to 
divert substantial resources to the next child if it is a boy.
Assuming that these investments can compensate for the adverse effects,
we should see significant differences in mortality across sex for short-spacing 
children in the absence of boys.
If a son is born, his mortality risk should be substantially lower than expected 
for the spacing duration, and we should expect worse outcomes for the other 
children (all girls).
If a girl is born, there are two possible outcomes.
If parents treat all children/girls the same, we should see a slight worsening in 
health outcomes because fewer resources are available per child; 
if they do not, the last girl will have a much lower likelihood of surviving 
because the parents' incentive is to invest in their prior children rather than 
the new child.

As mentioned above, the adverse effects of shorter spacing are strongly mitigated 
by higher education, which, in theory, allows better-educated women to have shorter
spacing in the absence of a boy without substantial increasing mortality risk.
Differences in mortality outcomes across boys and girls are still possible,
but should not be as large as for less-educated women.

With the introduction of sex selection, parents can avoid the birth of girls, 
although at the cost of the procedure and having to wait longer until 
the next child if they abort the fetus.
The question is, who is likely to use sex selection, when, and why?

Theory suggests that the lower the desired fertility, the 
higher the likelihood of sex selection \citep{Portner2015b}.
Simple probabilities drive this prediction:
in the absence of sex selection, 
48.8 percent will not have a son when only having one child, 
23.8 percent for two children, 11.6 percent for three children, 
5.7 percent for four children, 2.8 percent for five children, 
and 1.4 percent for six children.
Furthermore, for a given level of desired fertility, the higher
the parity, the higher is the likelihood of using sex selection.

The higher the cost of prenatal sex determination and abortion,
the less use we should see.
Similarly, any potential adverse effects on the mother and offspring 
from repeated abortions would also tend to reduce the use of sex 
selection---provided that parents are aware of these costs.
Finally, the more abortions between births, the less the opportunity 
there is to take advantage of any economies of scale in childrearing.

The use of sex selection in India increases with education and is higher 
in urban areas than in rural
\citep{das_gupta97,retherford03b,jha06,Guilmoto2009a,Bongaarts2013,Portner2015b,
Jayachandran2017}.
Women with more education generally have lower desired fertility, either 
because of the higher opportunity cost of time or because of other changes 
in the preferred number of children.%
\footnote{
It is also possible that the effect is driven by a stronger son preference, although 
higher education women have shown declining stated son preference \citep{bhat03,pande07}.
}
Better-educated women also tend to live in households with higher 
income, which better enables them to access prenatal sex determination 
and lowers the relative costs of using sex selection.

Given the increase in birth spacing with each abortion, parents may try to 
conceive even sooner to compensate for the potentially longer spacing and, 
thereby, still take advantage of economies of scale in childrearing.
Furthermore, if health outcomes for short spacing births have improved over 
time---as is likely---we should see an even stronger push toward shorter 
spacing for women using sex selection.

In summary, birth spacing is affected by female education through 
multiple pathways, with the main one of interest here the use of 
sex selection. 
But, to fully understand how sex selection affects spacing, we also 
need to understand the other changes that affect birth spacing over time.
The most likely factors are the changes in household income and 
wealth and the improvements in health as India has developed.
I, therefore, now look at how female education and labor force 
participation has changed over time in India.