### 2020-01-29 paper/infant_mortality.tex
Add figures and initial discussion of results

\section{Infant Mortality}

The prior literature show that shorter birth spacing, especially very short spacing,
is associated with higher risk of infant mortality.
Hence, it is possible that the increasing use of sex-selective abortions could
directly reduce infant mortality by lengthening the duration from the prior birth.

There are a number of potential issues with understanding to what extent this
has happened.

First, the sex of the next child is no longer random, and boys, in the absence of
differential treatment, have higher infant mortality than girls.

Second, it may be that what matters is not the time between births, but rather 
between pregnancies, in which case multiple abortions followed quickly by new 
pregnancies would lead to shorter real spells of recovery.
The implication is that long spells might have a higher mortality risk since
they are now more likely with multiple abortions.

Third, as I show above the use of sex selection is more prevalent the higher the
education level of the mother.
Higher maternal education is, however, also associated with lower mortality risk
to begin with.
On top of this, with economic growth and better health care, infant mortality
may have fallen across all education groups and areas in the period covered.

I address three questions.

First, does the increased use of sex selection lead to higher mortality risk?

Second, has the negative effects of short spacing changed over time?

Third, how do the effects of short spacing vary by education level in India?

There are a number of problems with estimating the relationship between 
sex-selective abortions and infant mortality.

First, there are potential selection problem.
For example, women who face higher mortality risk of their children might forego 
the use of sex selection to ensure that they reach their desired number of 
children.
If these women also know to avoid very short birth spacing, the precense of 
these women in the sample would inflate the mortality of the "middle" birth
spacing, relative to the very short or very long birth spacing.

[our JDE paper]

An additional selection with the opposite effect is that women who have
difficulties conceiving or carrying a preganancy to term may also have higher 
mortality risk for their offspring, leading to a supriuos correlation
between very long birth spacing and mortality \citep{Kozuki2013}. 
If better health care and higher income means that more of these women are
now able to carry their pregnancy to term we can end up with more births
after long spells that have higher mortality risk but not because of the
increased use of sex selection.

[Our Demography paper shows a different type of selection effect, where the
use of sex selection biases the effect estimates of other variables that
vary together with the use of sex selection.]

Given the data and that the focus of this paper is mainly on the birth
spacing, a more in-depth treatment of these issue is the topic of future work.
Part of the problem here is also that the method used previously, such
as family fixed effects does not work well when the number of births is
very low as it is for higher education women who use sex 
selection, especially since the spacing from marriage to first birth
does not provide any information \citep{Kozuki2013,Molitoris2019}.
However, the fixed effects results and linear probability do not deviate
substantially.

Second, the actual number of abortions are unobserved.
The results above do show, however, how the use of sex selection vary
by education and sex composition of prior children.
I, therefore, estimate infant mortality risk taking into account the
sex composition of the prior children, as well as the index child.

Starting with the sample used for estimating birth spacing, I select, by
parity, children born more than 12 months before the survey month.
The dependent variable is whether the child died within the first twelve
months of life (infant mortality).
As above, I restrict the analysis to parity two through four.

The main set of explanatory variable consists of dummies for birth spacing,
divided into less than 15 months from when conception could realistically
could happen (same definition as above) and based on the finding that 
spacing less than 24 months between birth is associated with higher
infant mortality \citep{Molitoris2019}, and then in 12 months intervals
until TK, where all births after and until 96 months are included.%
\footnote{
\citet{Molitoris2019} has used quartic functions, but the concern is
that with the relatively small number sample here the shape
of the function would be strongly influenced by a small number of
observations.
Using dummies provides a more flexible specification and the standard
error will directly reflect the higher degree of noise for the
spells periods with fewer observations.
}
In addition, I use dummies for sex of the index child (the child whose 
mortality is estimated) and the sex composition of the prior children.
Birth spacing dummies, sex of index child, and sex composition dummies
are all interacted.
The other explanatory variables are the same as above and estimations
are done separately by education level.

I estimate probability of infant mortality using a logit model.
Results are presented as the predicted probability of dying within the
first year at the possible combinations of index child sex, sex composition
of prior children, and the birth spacing dummied.
All other variables are set to their average values.
[TK ehh, did I also do urban/rural?]

Figures \ref{fig:mortality_low} to \ref{fig:mortality_highest} show the 
estimated probabilities.
To ease legibility all confidence intervals are suppressed
(graphs with confidence intervals are available on request).

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_low_dummies} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p1_low_dummies}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_low_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_low_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_low_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_low_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_low}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_med_dummies} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p1_med_dummies}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_med_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_med_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_med_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_med_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_med}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_high_dummies} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p1_high_dummies}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_high_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_high_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_high_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_high_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_high}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_highest_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p2_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_highest_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p3_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_highest_dummies} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{mortality_s4_p4_highest_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_highest}
\end{figure}

For women with twelve or more years of education the results become very noisy
because of the low number of births, especially for parity three and four, and
the low mortality rate.
As above, I do not show results for the 1972--84 period because there were too
few women with this level of education for this period.

Based on the second parity births, women with twelve or more years of education
have the lowest infant mortality risk. 
Even for very short birth intervals the mortality risk is less than five percent,
and fall to less than half that for longer intervals.
Given the low mortality there is, not surprisingly, much difference between
the three periods covered.
Furthermore, there is no evidence that longer intervals are associated with
significantly higher mortality, especially for the groups most likely to have
used sex selection (women with a girl as their first child and a boy as their
second with a long interval).

Women with eight to eleven years of education have marginally higher mortality
risk than women with the most education, and also show slightly more evidence
of increased mortality with very short spacing.
Mortality risk have decreased slightly between periods, although the changes
are relatively minor.
As for women with the most education, there are not any statistically significant 
differences or increases in mortality with longer spacing.
Because of the low mortality risk and the relatively small number of births,
the estimates for the third and fourth parity are noisy.

Infant mortality risk for women with one to seven years of education have
clearly decreased over time.
This is also the first group that show increasing mortality risk with longer
spacing, although this effect mostly disappears for the later periods.
In other words, it does not appear to be the result of sex selection, but
rather arise from the selection bias discussed above where women with long
spells are those who find it difficult to either conceived and/or carry a 
pregnancy to term and also have higher mortality risk for their offspring.

Finally, women with no education have experienced substantial improvements
in infant mortality risk over time.
This decrease is especially pronounced for shorter intervals.
For the longer intervals, there is surprisingly little difference across
the education groups for the later periods.