### 2020-02-10 paper/sex_selection_spacing-ver3.tex
Delete old conclusion

\documentclass[12pt,letterpaper]{article}

\usepackage{amsmath}
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}
\usepackage[title]{appendix}
\usepackage[margin=1.0in]{geometry}
\usepackage[figuresleft]{rotating}
\usepackage[longnamesfirst]{natbib}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}
\usepackage{setspace}
\usepackage[justification=centering]{caption}
\usepackage[font=scriptsize]{subfig}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{adjustbox}
\usepackage{xfrac}
\usepackage{placeins}

\usepackage{endfloat}

% \bibpunct{(}{)}{;}{a}{}{,}
\newcommand{\mco}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\mct}[1]{\multicolumn{2}{c}{#1}}
\newcommand{\mcth}[1]{\multicolumn{3}{c}{#1}}
\newcommand{\X}{$\times$ }
\newcommand{\hs}{\hspace{15pt}}

% Attempt to squeeze more floats in
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

%------------------------------------------------------------------------

\title{Birth Spacing in the Presence of Son Preference and Sex-Selective Abortions:
India's Experience over Four Decades%
\protect\thanks{%
I am grateful to Andrew Foster and Darryl Holman for discussions about the method.
I owe thanks to Shelly Lundberg, Daniel Rees, David Ribar, 
Hendrik Wolff, seminar participants at University of Copenhagen, University of Michigan, 
University of Washington, University of Aarhus, the Fourth 
Annual Conference on Population, Reproductive Health, 
and Economic Development, and the Economic Demography Workshop for helpful 
suggestions and comments.
I would also like to thank Nalina Varanasi for research assistance.
Support for development of the method from the University of Washington Royalty 
Research Fund and the Development Research Group of the World Bank is gratefully 
acknowledged.
The views and findings expressed here are those of the author and
should not be attributed to the World Bank or any of its member countries.
Partial support for this research came from a Eunice Kennedy Shriver National
Institute of Child Health and Human Development research infrastructure grant,
5R24HD042828, to the Center for Studies in Demography and Ecology at the
University of Washington.
}
}

\author{Claus C P\"ortner\\
    Department of Economics\\
    Albers School of Business and Economics\\
    Seattle University, P.O. Box 222000\\
    Seattle, WA 98122\\
    \href{mailto:claus@clausportner.com}{\texttt{claus@clausportner.com}}\\
    \href{http://www.clausportner.com}{\texttt{www.clausportner.com}}\\
    \& \\
    Center for Studies in Demography and Ecology \\
    University of Washington\\ \vspace{2cm}
    }

\date{January 2019}

%------------------------------------------------------------------------

\begin{document}
\graphicspath{{../figures/}}
\DeclareGraphicsExtensions{.eps,.jpg,.pdf,.mps,.png}

\setcounter{page}{-1}
\maketitle
\thispagestyle{empty}

% \setcounter{page}{0}

\newpage
\thispagestyle{empty}
\doublespacing

\begin{abstract}

% Demography abstract
\noindent

\noindent JEL: J1, O12, I1
\noindent Keywords: India, prenatal sex determination, censoring, competing risk, non-proportional hazard
\end{abstract}

\newpage

%------------------------------------------------------------------------

\section{Introduction\label{sec:intro}}

India has changed significantly over the last four decades:
the economy has grown substantially,
education levels have increased for both males and females,
and the total fertility rate has fallen to 2.2 
\citep{Bosworth2008,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

India has, however, also experienced a rapidly expanding 
access to prenatal sex determination.
Combined with continued strong son preference, the result has been a dramatic 
increase in the males-to-females ratio at birth
\citep{das_gupta97,Arnold2002,retherford03b,Guilmoto2012,
Portner2015b,Jayachandran2017}.

What the prior research on India has not appreciated is that since each abortion 
increases the interval between births by six to twelve months, the growing use of 
sex-selective abortions may substantially increase birth spacing.%
\footnote{
The increase consists of three parts.
First, starting from the abortion, the uterus needs at 
least two menstrual cycles to recover;  otherwise, the likelihood 
of spontaneous abortion increases substantially \citep{zhou00b}.
The second part is the waiting time to conception, which is between 
one and six months \citep{Wang2003}.
Finally, sex determination tests are reliable only from three months 
of gestation onwards.
}
Furthermore, the combination of increasing female education, 
lower fertility, higher household income, and the low and declining female 
labor force participation are also likely to impact birth spacing.
Hence, the changes in birth spacing may be much more substantial in India 
than what we have observed in other countries.

Understanding how birth spacing changes is of interest for two main reasons, in 
addition to a better general understanding how fertility decisions are made.

First, if birth intervals increased substantially over time, India's fertility may 
be higher than generally accepted based on the total fertility rate (TFR).
Increasing birth spacing is a tempo effect, which, by postponing births, makes the TFR a 
downward biased estimate of cohort fertility 
\citep{Hotz1997,Bongaarts1999,Ni-Bhrolchain2011}.

Second, longer birth spacing is associated with lower mortality and 
morbidity risk \citep{Conde-Agudelo2012,Molitoris2019}.
In India, the duration to the next birth has traditionally been shorter with fewer sons, 
which likely contributed to the higher mortality and morbidity risk for girls
\citep{Whitworth2002,Bhalotra2008,Maitra2008,Jayachandran2011,Jayachandran2017a}.
Longer spacing between births then potentially lowers mortality, although there may
be counteracting effects if the longer spacing arises because of multiple abortions.
There has, so far, been no attempt to examine how mortality has responded to the
increase in sex selection.

% Finally, birth spacing has long served as a measure of son preference, with strong
% son preference associated with shorter birth intervals, the fewer sons a family has 
% \citep{ben-porath76b,Leung1988}.
% Now, longer birth intervals in the absence of sons may instead arise because couples 
% with a strong preference for sons use sex selection, possibly to the point where we 
% observe a reversal of the traditional spacing pattern.
% Despite a large literature on son preference and birth spacing, no prior research has,
% however, examined how sex selection affects spacing.

I examine how birth spacing in India has changed over time and across 
groups with the introduction of sex selection.
Using data from the four National Family and Health Surveys (NFHS), I apply
a competing risk hazard model to the birth histories of Hindu women, covering
the period 1972 to 2016.
The competing risk hazard model directly incorporates the effects of 
sex-selective abortions on birth intervals
\emph{and} 
the likelihood of a son, 
but also works in cases without prenatal sex selection.
The empirical model allows me to predict fertility behavior taking into
account both the likelihood of parity progression and having a child of a specific sex.

% [Hypotheses / questions]
There are four main set of results.

First, birth intervals have increased for all education groups over the four decades, 
with larger increases, the higher the parity and the higher the education level.
Women who are most likely to use sex selection---well-educated women with no
sons---have seen the most substantial increases in birth intervals.
[TK check this] As a result, well-educated women now show a reversal of the traditional spacing 
pattern, with the longest intervals for women with no sons.
Those least likely to use sex selection, women with less education in rural areas, 
still follow the traditional pattern of short spacing when they have girls and 
little evidence of the use of sex selection.
[TK add sex ratios and parity progression discussions]

Second, there is little difference over time and across groups in the likelihood of 
having a very short birth interval given the sex composition of prior children, 
especially for the lower parities.
Hence, the increases in average spacing come predominantly from the longer birth intervals
getting even longer over time, driven to a large degree by a substantial use of sex 
selection.
The exception to this result is that the better-educated women with \emph{very high} 
predicted use of sex selection appear to conceive later exactly because the next child is 
almost certainly a boy.

Third, the fertility rate substantially overestimated how fast cohort fertility fell 
in the 1990s and early 2000s, when spacing began to increase as the result of secular 
changes and increasing use of sex selection.
Although predicted cohort fertility and the fertility rate has lately begun to converge,
the predicted cohort fertility is still ten to twenty percent higher than the
fertility rate.
Furthermore, even with the convergence, predicted cohort fertility is still at or above 
replacement level, except for the best-educated urban women.

Finally, infant mortality risk has decline substantially over time for all groups,
but fastest for the lower education groups, who are now close to the level of women with 
the most education.
Mortality risk is, however, still inversely related to education level, especially for 
very short birth intervals.
There is no evidence that the increasing use of sex selection is associated with higher 
mortality risk.

%------------------------------------------------------------------------------------

\section{Female Education and Labor Force Participation in India}

To set the stage for the subsequent analyses, I first describe how female education
and labor force participation has changed.
Both are potentially important factors in the birth spacing decision, as I discuss
in the next section.

Female educational attainment has increased substantially over time.
Figure \ref{fig:education_over_time} shows the distribution of schooling by birth 
cohort for urban and rural women, 
twenty years or older, whether married or not, based on the four rounds of the NFHS.
The education groups are no education, one to seven years, and eight to eleven years,
and twelve years and above.

\begin{figure}[htpb]
\centering
\subfloat[Rural]{
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth]{educ_over_time_rural}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth]{educ_over_time_urban} 
    \end{minipage}
}
\caption{Distribution of education by cohort for women twenty years or older at survey}
\label{fig:education_over_time}
\end{figure}

For rural areas, the percentage of women with no education has gone from around 90 percent
for the 1930s cohorts to less than 20 percent for the 1990s cohorts. 
The proportion of women with one to seven years of education has remained remarkably
constant at around ten percent.
The difference is made up by the women with eight or more years of education, who have
gone from almost zero for the 1930 cohort to more than sixty percent for the 1990s cohorts,
with about half in the eight to eleven years group and 
the other half in the twelve years or more group.

Female education is higher in urban areas than in rural areas.
Around sixty percent of urban women born in the 1930s had no education, 25 percent had
between one and seven years, about ten percent had eight or more years of education, and 
only five percent has twelve years or more.
The proportions with no education and one to seven years have both declined to just below 
ten percent for the latest cohort.
Although the proportion of women with eight to eleven years of education has increased
to about twenty, most of the increase in urban female education has come from the 
twelve plus group, which now account for more than half of all urban women.

Even as the level of female education has increased, the female labor force participation 
in both urban and rural areas has stagnated or decreased
\citep{Klasen2015,Fletcher2017,Afridi2018,Bhargava2018,Chatterjee2018,Bhargava2019}.
A decline in female labor force participation at the beginning of development is
consistent with the hypothesis of a U-shaped labor force participation as a country
develops \citep{Goldin1994}.
India's female labor force participation is, however, lower than most other
countries and more in line with countries in the Middle East and North Africa, and
does not yet show any signs of increasing
\citep{Klasen2015,Chatterjee2018}.

The NFHS data show a U-shaped relationship between education and working for married
women, with the highest percent working for women with either no education 
or with twelve or more years and the lowest for women with eight to eleven years of 
education.%
\footnote{
See Appendix Figures \ref{fig:work_by_survey} through \ref{fig:work_family_by_survey}. 
}
Women are more likely to report working if they live in rural than urban areas 
and the older they are.
For married women there is relatively little change in the percentage currently working 
over time,
but all education and age groups have become substantially more likely to work for
a family member.

The increases in female educational attainment imply that access to education has 
expanded beyond the higher castes. 
One possible effect of the rapid expansion in female education, and the 
associated change in the composition of better-educated women, is that the behavior
of the better-educated would change. 
However, ``Sanskritization'' implies that as lower castes females gain access to 
education and their husbands' income increases, the women adopt higher caste 
norms such as stronger son preference and a retraction from the formal labor 
market \citep{Srinivas1956,Chen1995,Abraham2013,Chatterjee2018}.
The low and declining female labor force participation is a first indication that this
process still operates.

\section{Birth Spacing: Mechanisms and Prior Findings}

To understand the potential implications for birth spacing of the substantial changes in 
income, female education, and female labor force participation that India has experienced, 
this section discuss relevant theories and prior findings on birth spacing.

The standard economic argument for why increasing female education has predominantly
been associated with shorter spacing is that parents incur time costs when they have 
children \citep{Hotz1997,schultz97}.
Specifically, if having children require the mother to reduce her market work, parents 
can lower the cost of children by shortening birth spacing to 
take advantage of economies of scale in childrearing \citep{Vijverberg1982}.

The low and declining female labor force participation, especially for younger women, 
suggests, however, that families face little incentive to space children more closely 
together for economic reasons.
This may happen if household income has increased so substantially that the income effect 
dominates any substitution effect.
Two findings speak for this effect.
First, although real wages for both men and women have almost doubled between 1987 and 
2011, the mean male wage is still close to 70 percent higher than the female wage 
\citep{Klasen2015,Bhargava2018}.
Second, women’s labor supply appear to be more negatively elastic to their husbands' 
wages than it is positively elastic to own wages \citep{Bhargava2018}.

The combination of rising incomes and continued son preference may lead to even longer 
spacing than the standard income effect alone predicts.
As women's education increases, their productivity in the production of offspring human 
capital also increases.
With relatively more boys born because of increased access to sex-selective 
abortions and an increasing potential income for (male) offspring, demand for 
better-educated women can increase, even if they do not participate in the labor market 
\citep{Behrman1999}.
If more and ``better'' parental attention per child results in higher child ``quality'' 
we should expect longer birth intervals 
\citep{Zajonc1975,Zajonc1976,Razin1980}.
The evidence on spacing's effect on child quality measures such as IQ 
and education is, however, mixed for developed countries and non-existing 
for developing countries
\citep{Powell1993,Pettersson-Lidbom2009,Buckles2012,Barclay2017}.
The exception is one crucial aspect of child quality---health and mortality---where
longer spacing does lead to better outcomes, although this relationship weakens with 
maternal education \citep{Whitworth2002,Conde-Agudelo2012,Molitoris2019}.

% While one might expect that increasing income and female education would also lengthen
% average birth spacing through better access and knowledge of contraception, the empirical 
% relationship between contraception use and birth spacing is ambiguous 
% \citep{Tulasidhar1993,Whitworth2002,Bhalotra2008,Yeakey2009,Kim2010,Soest2018}.
% Furthermore, the ability to successfully use ``low efficacy'' contraceptive methods, 
% such as the rhythm method, is increasing in education, while there is no difference 
% across education levels when it comes to modern contraceptives such as pills or 
% injectables \citep{Rosenzweig1989}.

The introduction of sex selection allows parents to avoid the birth of girls,
but increases the expected duration to the next birth.
Theory suggests that sex selection increases with lower desired fertility, and, for a 
given desired number of children, the higher the parity \citep{Portner2015b}.
The increased use of sex selection in India with education and in urban
areas is consistent with a lower desired fertility for both groups
\citep{das_gupta97,retherford03b,Guilmoto2009a,Portner2015b,Jayachandran2017}.
Better-educated women also tend to live in households with higher income, which better 
enables them to access prenatal sex determination and lowers the relative costs of using 
sex selection and have long birth intervals.

We may even observe a ``delay'' effect, where parents conceive later because they know 
that short spacing is detrimental to the next child's health, and that the next child is 
more likely to be a boy because of access to sex selection.
Working in the opposite direction is that women with more education can, in principle, 
space their children closer together without substantially increasing child mortality 
risk.
Given the apparent lack of pressure to return to the labor force,
this effect is, however, likely to be small.

In summary, with substantial increases in husbands' income and a declining female labor 
force participation, I expect a push toward longer birth spacing over time, independent
of education levels, based on the income effects and the effects of spacing
on child outcomes.
As desired fertility decreases with increasing education, I, furthermore, expect 
birth spacing to increase the most among the better-educated because of the larger
increase in household income income---even with declining female labor participation---and 
their use of sex selection.
Even with the large increase in the number of better-educated women, "Sanskritization" 
imply that the changing composition does not substantially change this group's use of 
sex selection.

%------------------------------------------------------------------------------------

\section{Estimation Strategy\label{sec:strategy}}

% Requirements for method and the intuition behind each:  
%  - Account for censoring (hazard model)
%  - Work both with and without sex selection (competing risk part)
%  - Allow for changes in use of sex selection within spell (flexible specification of baseline hazard ?)
%  - Capture that the shape of the hazard function differ across groups
%    (non-proportionality); needs a better description, maybe using example
%    of shorter spacing and/or use of sex selection. Differences in parity
%    progression likelihood.

The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit, the birth of a child.%
\footnote{
See \citet{Sheps1970} and \citet{Newman1984} for early discussions of why 
hazard models are the preferred way to deal with the censoring of birth
intervals.
}
There are two problems with the standard approach in this setting.
First, and most importantly, the introduction of sex selection means that the sex of the 
next child is no longer necessarily random and that parents' choices will impact the 
spacing to the birth of a girl or a boy differently.
I, therefore, use a competing risk set-up, which can capture both the non-randomness of
the birth outcome and the differential spacing.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

Second, it is unlikely---even in the absence of prenatal sex determination---that 
characteristics, such as the sex composition of previous births, have the same effects 
throughout the entire spell as is assumed by the proportional hazard model.
The proportionality assumption is especially problematic for higher-order spells where 
there are substantial differences across groups both in the likelihood of progressing to 
the next birth and how soon couples want their next child if they are going to have one.

The introduction of prenatal sex determination exacerbates any bias from the 
proportionality assumption since sex-selective abortion use varies across groups,
which affects birth spacing, and because a household's use of sex selection may vary 
within a spell, which means that the effects of covariates vary within the spell as well.

I, therefore, use a non-proportional hazard specification, which allows the shape of the 
hazard functions to vary across groups.
The use of a non-proportional specification also mitigates any potential effects 
of unobserved heterogeneity when used in conjunction with a flexible baseline hazard 
\citep{Dolton1995}.

The model is a discrete time, non-proportional, competing risk hazard model with two exit 
states: either a boy or a girl is born.
The unit of analysis is a spell, the period from nine months after one birth to the next.
For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$, and 
the spell continues until time $t_i$, when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: the birth of a boy, $j=1$, or the birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is 
\begin{equation}
 h_{ijt} = \frac{\exp(D_j(t) + \alpha_{jt}'\mathbf{Z}_{it} + \beta_j'\mathbf{X}_{i})} 
 {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{it} + \beta_l'\mathbf{X}_{i})} \: \: \; \; \;  j = 1,2
 \label{eq:hazard}
\end{equation}
where the explanatory variable vectors, $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$, capture 
individual, household, and community characteristics discussed below,
and $D_{j}(t)$ is the piece-wise linear baseline hazard for outcome $j$, captured
by dummies and the associated coefficients,
\begin{equation}
D_j(t) = \gamma_{j1} D_1 + \gamma_{j2} D_2 + \ldots + \gamma_{jT} D_T,
\end{equation}
with $D_m = 1$ if $t=m$ and zero otherwise.
This approach to modeling the baseline hazard is flexible and does not place 
overly strong restrictions on the baseline hazard.

The explanatory variables in $\mathbf{Z}$, and the interactions between them, 
constitute the non-proportional part of the model, which means that they are
interacted with the baseline hazard:
\begin{equation}
 \mathbf{Z}_{it} = D_j(t) \times (\mathbf{Z}_1 + Z_2 + \mathbf{Z}_1 \times Z_2),
\end{equation}
where $D_j(t)$ is the piece-wise linear baseline hazard, $\mathbf{Z}_1$ captures sex 
composition of previous children, if any, and $Z_2$ captures area of residence.
The remaining explanatory variables, $\mathbf{X}$, enter proportionally,
but to further minimize any potential bias from assuming proportionality, estimations 
are done separately for different levels of mothers' education and different 
periods.%
\footnote{
I discuss the choice of which variables to use for non-proportionality in more detail
in the ``Explanatory Variables'' section.
}

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, transforming the data, so each observation is an interval---here equal
to three months---the model can be estimated using a standard multinomial logit model.
%
% \footnote{
% The multinomial model assumes that alternative exit states are stochastically independent
% (Independence of Irrelevant Alternatives or IIA), which rules out any individual-specific 
% unmeasured or unobservable factors that affect both the hazard of having a girl and the 
% hazard of having a boy.
% I, therefore, include a proxy for fecundity discussed in Section \ref{sec:data}.
% Also, the multivariate probit model can be used as an alternative because the IIA is not 
% imposed \citep{han90}.
% The results are substantially identical between these two models and
% available upon request.
% }
In the reorganized data the outcome variable is 0 if the woman does not have a child in a 
given interval (the base outcome), 1 if she gives birth to a son in that interval, and 2 
if she gives birth to a daughter in that interval.

% The distribution of spacing is captured by the survival curve, which shows the probability 
% of not having had a birth yet by spell duration, for a given set of explanatory variables.
% The survival curve at time $t$ is 
% \begin{equation}
% \label{eq:survival}
% S_{t} 
% = 
% \prod_{d=1}^t
% \left(
% \frac{ 1 }
% {1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{ld}'\mathbf{Z}_{kd} + \beta_l'\mathbf{X}_{k})}
% \right).
% \end{equation}
% 
% Because the probability of ever having a next birth varies across groups, a direct 
% comparison of standard survival curves tells us little about how the spread of sex 
% selection affects birth spacing across groups.
% I, therefore, condition on the predicted likelihood of parity progression when examining 
% birth spacing measures, such as the average duration to a birth.
% The reliability of this approach depends on whether the spell length covered is 
% sufficiently long that few women are likely to give birth after the spell cut-off.
% I discuss the choice of spell length below.%
% \footnote{
% It is important to note that the approach is not the same as merely calculating 
% the birth spacing measures for women who already have a given 
% parity child in the survey because that number does not take into account
% the censoring of spells that will eventually lead to a birth.
% }
% In addition, I present graphs of survival curves conditional on parity progression,
% which therefore begin at 100\% and end at 0\%.

Interpretation of the model coefficients is challenging \citep{thomas96}.
It is, however, possible to calculate the predicted probabilities of 
having a boy, $b$, and of having a girl, $g$, in period $t$, conditional on 
a set of explanatory variables and not having had a child before that period.
% , as
% \begin{align}
% P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt}, t ) 
% & =  
% \frac{ \exp(D_j(t) + \alpha_{1t}' \mathbf{Z}_{kt} + \beta_1' \mathbf{X}_{k} )}
% {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt} ' \mathbf{Z}_{kt} + \beta_l ' \mathbf{X}_{k})}
% \label{eq:probability_boy} \\
% P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t ) 
% & =  
% \frac{ \exp(D_j(t) + \alpha_{2t}'\mathbf{Z}_{kt} + \beta_2'\mathbf{X}_{k} )}
% {1 + \sum_{l=2}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{kt} + \beta_l'\mathbf{X}_{k})}
% \label{eq:probability_girl}
% \end{align}
It is then straightforward to calculate the estimated percentage of children born that 
are boys, $\hat{Y}$, at each $t$:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100.
\label{eq:probability_son}
\end{equation}
Combining the percentage boys and the likelihood of exiting the spell 
across all $t$ gives the predicted percent boys born over the entire spell.

\section{Data\label{sec:data}}

The data come from the four rounds of the National Family Health Survey
collected in 1992--1993, 1998--1999, 2005--2006, and 2015--2016.%
\footnote{
A delay in the survey for Tripura means that NFHS-2 has some observation 
collected in 2000.
}
The surveys are large: NFHS-1 covered 89,777 ever-married women 
aged 13--49 from 88,562 households;
NFHS-2 covered 90,303 ever-married women aged 15--49 from 92,486 households;
NFHS-3 covered 124,385 never-married and ever-married women aged 
15--49 from 109,041 households;
and 
NFHS-4 covered 699,686 never-married and ever-married women aged
15--49 from 601,509 households surveyed.

I exclude visitors to the household, as well as women who fall into any of the following 
categories: never married; no gauna yet; married more than once; divorced; 
not living with husband; inconsistent age at marriage; or education information missing.  
The same goes for women who had at least one multiple births, reported having a birth 
before age 12, had a birth before marriage, or an interval between births of less than 
nine months. 
Finally, I restrict the sample to Hindus, who constitute about 80\% of India’s population.%
\footnote{
If the use of sex selection differ between Hindu and other groups assuming
that the baseline hazard is the same would lead to bias. The other
groups are each so small that it is not possible to estimate different
baseline hazards for each group, and are so different that combining
them into one group would not make sense.
}

In addition to the large number of women surveyed and the long period covered, a
benefit of the NFHS is that surveys enumerators pay careful attention to spacing between 
births and probe for ``missed'' births.
Nevertheless, systematic recall error, where the likelihood of reporting a deceased 
child depends on the sex of the child, is still a potentially substantial problem and
will bias both spacing and sex ratios.
As I show in the online Appendix, recall error is heavily dependent on how long ago a 
woman was married, and I, therefore, drop women married 22 years or more.
The final sample consists of 
\input{../tables/num_women.tex}
% [number] women, with [number] parity one through four births.

Direct information on the use of sex selection is not available, so I compare different 
periods as a proxy, based on the changes in access and legality of prenatal sex 
determination in India.
Abortion has been legal in India since 1971.
Reports of sex determination appeared around 1982--83, and the number of clinics 
quickly increased \citep{Sudha1999,bhat06,Grover2006}.
In 1994, the Prenatal Diagnostic Techniques (PNDT) Act made determining and communicating 
the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
There is little evidence that the ban significantly affected sex ratios \citep{Das-Gupta2016}.
}
Although the use of sex selection increased even after 1994, some researchers argue that 
we have passed a turning point in its use \citep{Das_Gupta2009,Diamond-Smith2015}.
I, therefore, use four periods: 1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The allocation of spells into periods is determined by when conception, and therefore 
decisions on sex selection, can begin. 
This allocation mechanism means that some spells cover two periods, which may bias 
downward the differences between the periods.

I focus here on the second through fourth spells, that is, on the intervals from the first 
birth until at most the fourth birth.
I exclude the spell from marriage to the first birth because of data issues.
In NFHS-1 and NFHS-2 the only information on marriage is the age at marriage, and 
all first spell durations are therefore imputed.
In NFHS-3 about a third of respondent did not provide complete information on the
timing of marriage and had the length of their first spell imputed using the age at 
marriage.
The imputation results in large numbers of women with spells that are less than nine 
months and makes a comparison of spell duration over time subject to an 
excessive amount of noise.

The spells all begin nine months after the previous birth, which is the earliest we should 
expect to observe a new birth.
A spell continues until either a child is born or censoring occurs.
Censoring can happen for three reasons:
the survey takes place;
sterilization of the woman or her husband;
or too few births are observed for the method to work.
For all spells censoring is set at 96 months (eight years) after a birth can occur.
With these cut-offs, less than one percent of observed births occur after
the spell cut-off.

% %
% \footnote{
% \input{../tables/num_missed.tex}
% }

\subsection{Explanatory Variables}

I divide the explanatory variables into two groups.
The first group consists of characteristics that the prior literature finds affect 
spacing choice and the use of sex selection:
mother's education, sex composition of previous children, and area of residence.
The second group of variables consists of those expected to have an approximately 
proportional effect on the hazard.
These include the age of the mother at the beginning of the spell, whether the household 
owns any land, and whether the household belongs to a scheduled tribe or caste.

Women with different education levels have different hazard profiles
\citep{Whitworth2002,Bhalotra2008,Kim2010}.
Furthermore, as discussed above, the use of sex selection increases with education, 
either because of the associated lower fertility or because higher income enables them 
to access and afford prenatal sex determination
I divide women into four groups based on education attainment: no
education, one to seven years of education, eight to eleven years of education,
and twelve and more years of education.
The models are estimated separately for each education level to reduce the potential 
problem of including other variables as proportional.

The sex composition of previous children affects both the timing of births and the use of 
sex-selective abortions 
\citep{retherford03b,Bhalotra2008,Portner2015b,Soest2018}.
I capture sex composition with dummy variables for the
possible combinations for the specific spell, ignoring the ordering of births.
Finally, urban women have both lower fertility and, presumably, better access to
sex selection, both of which lead to higher use of sex-selective abortions than in 
rural areas \citep{retherford03b,Portner2015b}.
Area of residence is a dummy variable for the household living in an urban area.

% %
% \footnote{
% Fathers' education has two opposite predicted effects: the associated higher income
% should increase fertility and therefore lower the pressure to use sex selection, but
% the higher income also makes the use of sex selection cheaper.
% In practice, fathers' education had little effect on the hazards and the use of 
% sex-selective abortions, and I, therefore, do not include it.
% }

Appendix Tables \ref{tab:des_stat1} and \ref{tab:des_stat2} presents descriptive 
statistics for the spells by education level and period.
The level of censoring increases with parity and time, as we should expect with later 
childbearing, falling fertility, and the hypothesized increases in birth intervals from 
sex-selective abortions.

\section{How Birth Spacing has Changed\label{sec:results}}

% [How to read tables]
I estimate the model for each spell/education/period subsample, the results of which I use 
to predict average birth spacing, sex ratio, and the probability of having a birth in that 
spell.
Figures \ref{fig:bootstrap_low} through \ref{fig:bootstrap_highest}
show these predicted outcomes for the four education levels separated by the area of 
residence.%
\footnote{
For legibility, the graphs do not show standard errors.
The graphed values, with standard errors, are shown in Appendix Tables 
\ref{tab:avg_sex_ratio_low} through \ref{tab:avg_sex_ratio_highest}.
Furthermore, I do not show the results for the fourth spell the best-educated women since 
the low number of births, especially with one or more boys, make the spacing and sex ratio 
results very noisy.
Results are available upon request.
}

The predicted parity progression probability is the likelihood of a birth by 96 months 
after the beginning of the spell.
To find the expected average duration I first calculate, for each woman, the probability 
of giving birth in each period $t$, which I use as weights to calculated her expected 
duration. 
I then average the individual expected durations across women using the parity progression
probabilities as weights.

The predicted sex ratio captures the percent boys to women in the sample when 
childbearing for that spell is over.
A woman's predicted sex ratio is the weighted average of the predicted percentage boys 
over each $t$ in the spell, calculated using equation (\ref{eq:probability_son}), using
the probability of giving birth at time $t$ as weights.%
\footnote{
If T=2 with the estimated percent boys 54\% and 66\% and the likelihoods of giving 
birth 20\% and 40\%, the percent boys is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$.
}
The predicted sex ratio shown is the weighted average of the individual predicted sex 
ratios, using the parity progression probabilities as weights.
For comparison, the graphs also show the natural sex ratio, approximately 51.2\% 
\citep{ben-porath76b,jacobsen99,Portner2015b}.

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_low_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_low_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_low_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_low_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_low_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_low_rural_all}} 
% \caption{low education}
% \label{fig:bootstrap_low}
% \end{figure}

% \captionsetup[subfigure]{position=top,captionskip=-1pt,farskip=-0.5pt}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_low_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_low_urban_all} 
%     \end{minipage}
% }
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_low_urban_all}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_low_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_low_rural_all} 
%     \end{minipage}
% } 
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_low_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average spell length, sex ratio, and probability of 
% a next birth for women with no education}
% \label{fig:bootstrap_low}
% \end{figure}

\begin{sidewaysfigure}
\captionsetup[subfigure]{captionskip=-1pt,farskip=-0.5pt}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_low_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_low_urban_all} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_low_urban_all}
    \end{minipage}
}
\caption{Estimated average spell length, sex ratio, and probability of 
a next birth for women with no education}
\label{fig:bootstrap_low}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_low_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_low_rural_all} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.31\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_low_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average spell length, sex ratio, and probability of 
a next birth for women with no education}
\label{fig:bootstrap_low}
\end{sidewaysfigure}

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_med_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_med_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_med_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_med_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_med_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_med_rural_all}} 
% \caption{med education}
% \label{fig:bootstrap_med}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_med_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_med_urban_all} 
%     \end{minipage}
% }
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_med_urban_all}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_med_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_med_rural_all} 
%     \end{minipage}
% } 
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_med_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average spell length, sex ratio, and probability of 
% a next birth for women with one to seven years of education}
% \label{fig:bootstrap_med}
% \end{figure}

\begin{sidewaysfigure}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_med_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_med_urban_all} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_med_urban_all}
    \end{minipage}
}
\caption{Estimated average spell length, sex ratio, and probability of 
a next birth for women with one to seven years of education}
\label{fig:bootstrap_med}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_med_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_med_rural_all} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_med_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average spell length, sex ratio, and probability of 
a next birth for women with one to seven years of education}
\label{fig:bootstrap_med}
\end{sidewaysfigure}

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_rural_all}} 
% \caption{high education}
% \label{fig:bootstrap_high}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_high_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_high_urban_all} 
%     \end{minipage}
% }
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_high_urban_all}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_high_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_high_rural_all} 
%     \end{minipage}
% } 
% \subfloat[Fourth Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell4_high_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average spell length, sex ratio, and probability of 
% a next birth for women with eight to eleven years of education}
% \label{fig:bootstrap_high}
% \end{figure}

\begin{sidewaysfigure}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_high_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_high_urban_all} 
    \end{minipage}
}
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_high_urban_all}
    \end{minipage}
}
\caption{Estimated average spell length, sex ratio, and probability of 
a next birth for women with eight to eleven years of education}
\label{fig:bootstrap_high}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_high_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_high_rural_all} 
    \end{minipage}
} 
\subfloat[Fourth Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell4_high_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average spell length, sex ratio, and probability of 
a next birth for women with eight to eleven years of education}
\label{fig:bootstrap_high}
\end{sidewaysfigure}

% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_urban_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_urban_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_urban_all}} 
% \\
% \caption*{Rural}
% \subfloat[Second Spell]{\includegraphics[width=0.27\textwidth]{bs_spell2_highest_rural_all}} 
% \subfloat[Third Spell]{\includegraphics[width=0.27\textwidth]{bs_spell3_highest_rural_all}} 
% \subfloat[Fourth Spell]{\includegraphics[width=0.27\textwidth]{bs_spell4_highest_rural_all}} 
% \caption{high education}
% \label{fig:bootstrap_high}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{Urban}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_highest_urban_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_highest_urban_all} 
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{Rural}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell2_highest_rural_all}
%     \end{minipage}
% } 
% \subfloat[Third Spell]{
%     \begin{minipage}{0.29\textwidth}
%         \includegraphics[width=\textwidth]{bs_spell3_highest_rural_all} 
%     \end{minipage}
% } 
% \caption{Estimated average spell length, sex ratio, and probability of 
% a next birth for women with twelve or more years of education}
% \label{fig:bootstrap_highest}
% \end{figure}

\begin{sidewaysfigure}
\centering
\rotatebox[origin=c]{90}{\small{Urban}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_highest_urban_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_highest_urban_all} 
    \end{minipage}
}
\caption{Estimated average spell length, sex ratio, and probability of 
a next birth for women with twelve or more years of education}
\label{fig:bootstrap_highest}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\ContinuedFloat
\rotatebox[origin=c]{90}{\small{Rural}}
\subfloat[Second Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell2_highest_rural_all}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.29\textwidth}
        \includegraphics[width=\textwidth]{bs_spell3_highest_rural_all} 
    \end{minipage}
} 
\caption{(Continued) Estimated average spell length, sex ratio, and probability of 
a next birth for women with twelve or more years of education}
\label{fig:bootstrap_highest}
\end{sidewaysfigure}

Average birth intervals have increased over time across all groups, and are, for a given 
parity, longer the more education a woman.
The increases in birth spacing are larger the higher the parity
and, within parities, the higher the education.

In the latest period, the lowest average birth interval is for women with no 
education in the second spell at just shy of 25 months, which is an increase of a
couple of months from the first period.
The longest average interval was 42 months for the third spell for women with twelve or 
more years of education who had no boys, which is an increase of approximately a year.
For comparison, the median birth intervals reported for the NFHS have barely
moved over time, staying at approximately 22 months calculated with the starting point 
nine months after the prior birth used here.%
\footnote{
See
\citet[][p.\ 110--112]{International-Institute-for-Population-Sciences-IIPS1995},
\citet[][p.\ 98--103]{International-Institute-for-Population-Sciences-IIPS2000},
\citet[][p.\ 88--91]{International-Institute-for-Population-Sciences-IIPS2007}, and
\citet[][p.\ 81--82]{International-Institute-for-Population-Sciences-IIPS2017}.
Median estimates are in
Appendix Tables \ref{tab:p25_p50_p75_low} through \ref{tab:p25_p50_p75_highest}.
}
This lack of change underscores the importance of accounting for censoring of birth spells 
when trying to understand how birth spacing is changing over time.

Parity progression probabilities are lower the higher the education and are
falling across all groups over time, but the extent of the decline depends crucially on 
the sex composition of prior children and the parity.
As expected with a strong son preference, women with no sons still are the most likely to 
have a next birth, and this group has seen often substantially smaller declines in
the likelihood of having another birth than women with one or more sons.%
\footnote{
There is, however, also some evidence that parents prefer a mixed sex composition: for 
the fourth spell women with two boys \emph{and} a girl
consistently have the lowest parity progression probability.
}
For urban women with eight to eleven years of education, for example, the second spell
probability of a next child declined by eight percentage points if the first child was a 
girl, but 15 percentage points if it was a boy.
Even more extreme is the fourth spell where the likelihood of a birth declined by six 
percentage points to 75 percent if the woman had three children, but by almost 40 
percentage points to 25 percent if she had two boys and a girl.

To what extent has the introduction of sex selection changed birth spacing across sex 
compositions?
There are three broad groups in the data.

First, women who have only girls but do not use sex selection had, and continue to have,
the shortest spacing, as shown, for example, by rural women with no education.
For this group spacing is shortest with only girls across all spell, and the difference
between sex compositions have even widen over time.

Second, even with sex selection, and the associated increases in expected spacing, the
intervals for the other sex compositions have increased even more and the relatively 
pattern, therefore, remains more or less the same.
This is the case, for example, for the second spell for women with eight to eleven
and twelve or more years of education.

Finally, women with the most unequal sex ratios show the longest spacing in the 
absence of boys, and this spacing is longer than for the other possible sex 
compositions.
The most extreme example is that with a sex ratio for the best-educated urban women  
approaching 80 percent, the 42 months birth intervals with two girls is seven months 
longer than with at least one boy.
Even urban women with eight to eleven years of education show the longest intervals with 
only girls in the last period.

\subsection{Distribution of Birth Spacing}

Average spacing is a convenient way to understand the overall changes in behavior but may 
hide crucial differences in the distribution of spacing.
For example, if sex selection drives the changes in birth spacing, most of the increase in 
spacing will come from people who have two or more pregnancies rather than one and,
therefore, from changes in the longer intervals,
while a general increase in spacing would show up as an increase in spacing across the 
board.

Figures \ref{fig:combined_p25_p75_urban} and \ref{fig:combined_p25_p75_rural} 
show how the predicted 25th and 75th percentile birth intervals has changed over time 
by spell and education.
For each woman in a given spell/period combination, I calculate the time point at 
which there is a given percent chance that she will have given birth, relative to the
probability that she will eventually give birth in that spell.
The reported statistics is the average of the percentile duration
across all women in a given sample using the individual predicted
probabilities of having had a birth by the end of the spell as weights.%
\footnote{
The Appendix shows conditional survival curves for all groups and 
Appendix Tables \ref{tab:p25_p50_p75_low} through \ref{tab:p25_p50_p75_highest} 
show 25th, 50th, and 75th percentiles durations with bootstrapped standard errors.
}

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_low_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_low_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_low_urban}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_med_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_med_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_med_urban}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_high_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_high_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_high_urban}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_highest_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_highest_urban}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
25th and 75th Percentile Birth Intervals for Urban Women by Spell and Education
}
\label{fig:combined_p25_p75_urban}
\end{figure}

\begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup[subfigure]{position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_low_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_low_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_low_rural}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_med_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_med_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_med_rural}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_high_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_high_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell4_high_rural}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell2_highest_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{combined_p25_p75_spell3_highest_rural}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
25th and 75th Percentile Birth Intervals for Rural Women by Spell and Education
}
\label{fig:combined_p25_p75_rural}
\end{figure}

For both women with no education and those with one to seven years of education,
almost all of the increases in the average spacing discussed above come from increases in 
the long end of the spacing distribution rather than from a general increase.
Hence, spacing is still very short for many women, with 25 percent or more having
their next birth within two years of the prior birth, even in the last period.
Furthermore, the increases in the 75th percentile birth intervals are relatively small
for the second spell and slightly larger for the third spell at around five months.
Most of the divergence across sex compositions happen for the fourth spell, where 
the largest increases are for women with two or three sons.
However, even for women with no or one son there has been increases in the birth intervals 
of a half to a full year consistent with some use of sex selection for this group.

The 25th percentile birth intervals for women with eight to eleven years of education also 
changed relatively little over time and is, as for the less-educated women, close to
around two years since the prior birth.
The result is that most of the change in average spacing is driven by increases in the
longer spacing end of the distribution.
For example, the 75th percentile spacing length for women with two girls in the third 
spell increased by more than 15 months, mainly as the result of more widespread use of sex 
selection as the sex ratios above indicate.

The most educated urban women show an interesting pattern for the third spell, where there 
is essentially no change in either the 25th or 75th percentile birth intervals in the
presence of at least one boy.
The 75th percentile birth interval, however, increases by more than 20 months,
corresponding to the massive increase in the sex ratio to almost 80 precent in the 
latest period.
Hence, the increases in birth spacing and the reversal in spacing pattern by sex
composition over time is mainly driven by increases in the length of the longer intervals, 
consistent with a substantial use of sex selection.

Although most of the increases come from among the longer intervals, there is some 
evidence for the ``delay'' effect discussed above, where parents conceive later
exactly because they know that the next child is likely to be a boy since they plan
to use sex selection.
The best example is for the best-educated urban women in the third spell, where, in the 
last two periods, the spacing to the third child is significant longer if the two prior 
children were girls than if they already had one or two boys.
These two periods are the periods where the predicted sex ratios are 70 and 80 percent,
respectively.
There is also a significantly longer spacing with only girls for women with eight to
eleven years of education in the last period for the third and fourth spell, although
the difference is substantially smaller.

\section{What has Happened to Fertility?\label{sec:fertility}}

A potential effect of the substantial increases in birth spacing is that TFR may 
underestimate cohort fertility, at least temporarily.
I, therefore, compare predicted fertility based on a variation of the TFR and the 
hazard model by urban/rural and education group in Table \ref{tab:fertility}.

\input{../tables/fertility.tex}

To make the fertility rate comparable to the hazard model, it uses only births up to 
parity four, and is, therefore, not directly comparable to those in the NFHS reports.
Otherwise, I follow the procedure used in the NFHS reports and calculate age specific 
fertility rates for five-year age groups using the number of births in months 1 to 36 
before the survey month (births in the survey month are not included) and the number
of women in each age group.%
\footnote{
See \citet{Croft2018} for more detail.
% Because NFHS-1 and NFHS-2 did not interview all women about fertility, the number of women 
% is based on the household rosters, assuming that never-married women have had no children.
% Because of the low number of births to women aged 45 to 49 this age group was not
% used for the hazard model estimations and, therefore, is also excluded for the fertility
% rate.
% The marital total fertility rate lead to unrealistically high numbers and is, therefore,
% not presented \citep{Hoem2011,Laplante2015}.
}

Because NFHS-1 was collected after the introduction of sex selection it is not possible 
to calculate a fertility rate in exactly the same manner for a period 
before sex selection was widely available.
Instead, I calculate the fertility rates for women who was between 15 and 39 years of
age five years prior to the survey month, again using the number of births three years 
before that.
Given the relatively low number of births to women 40 to 45 years of age, this approach 
provides the best estimate of the fertility rate at the time when sex selection still was 
not widespread.
This fertility rate is shown as ``1987--88''.

To predict cohort fertility based on the hazard models above, I estimate the probability 
of progression to the next parity for each spell.
Since the parity progression depends on the sex composition for prior children for parity 
two and above, I estimate the progression probability for each possible sex composition
and weigh those probabilities with the likelihood that they will occur based on the
prior spells.

I include the spell from marriage to first birth, despite the problems capturing the 
exact timing of marriages for the first two rounds of the survey, since
the estimated progression probabilities should not be affected by this problem.
To ensure that the predictions depend solely on sample composition and the estimate, I
use the age at marriage for each woman and predict the likelihood of progressing to
each parity, assuming three years increases in age between each parity.
Shorter assumed duration between parities leads to slightly higher predicted 
fertility.

The survey rounds do not coincide directly with the periods I use for the hazard model.
For the NFHS-4, I use the model results for the 2005--2016 period, 
for NFHS-3 the 1995--2004 period, 
for NFHS-2 the 1985--1994 period, 
and for NFHS-1 the 1972--1984 period.
Because of the small number of women who had the highest level of education in the
1972--1984 period, I do not present a predicted fertility for this case.

In almost all cases the predicted cohort fertility based on the hazard model is higher 
than the four-parity fertility rate.
Only for women with no education in the first period are there almost no difference 
between the two fertility measures; a situation where fertility is high and spacing
very short and likely unchanged over a long period before.

Consistent with a larger bias in the fertility rate when age of marriage and the length of 
birth intervals are increasing, the absolute bias is least in the first and the last 
period and highest in the middle two periods.
Hence, the fertility rate declined too fast between the mid-1980s and the end of the 
century, and only recently, as the rate of increase for the birth intervals have slowed, 
are the predicted cohort fertility and the fertility beginning to converge again.
Even with the convergence, however, the predicted cohort fertility for 2005--16 is 
above the 1992--93 fertility rate for every group except urban women with no education.
Furthermore, for the last period, the predicted cohort fertility is stil between ten
and 20 percent higher than the fertility rate.

The bias in the fertility rate is also evident in the fact that the fertility rate is 
increasing over time for some groups.
For example, urban women with eight to eleven years of education showed 1.84, 1.81, and
1.87 over the last three surveys, and for rural women with one to seven years of
education fertility rate goes from 2.25 for NFHS-3 to 2.37 for NFHS-4.
These increases are likely the result of a stabilization of the age of first birth and
the spacing between births.%
\footnote{
Increases in TFR over time, even in the face of declining cohort fertility is not 
unusual, as shown by Columbia in the early 1990s \citep{Bongaarts1999}.
}

Despite the declines in the predicted cohort fertility it is still mostly above the 
generally accepted replacement level of 2.1.
Only for urban women with twelve or more years of education is the predicted cohort
fertility below 2.1, and even then it is still more than 0.3 children higher than the 
fertility rate estimate of 1.5.
Furthermore, because only the first four births are used, the predicted cohort fertility
numbers are likely too low, especially for the lower education groups.

\section{Mortality and the Changing Birth Spacing\label{sec:mortality}}

The prior literature show that shorter birth spacing, especially very short spacing,
is associated with higher risk of infant mortality.
Hence, the longer spacing, both from the increasing use of sex-selective abortions and
from secular changes, may directly reduce infant mortality.
To the extent that the longer birth intervals arise from multiple abortions
the short duration between \emph{pregnancies} could, however, increase mortality risk.
This section addresses how infant mortality has changed and whether there is evidence
of a negative impact of sex selective abortions.

Starting with the sample used for estimating birth spacing, I select children born 
more than 12 months before the survey month.
I restrict the analyses to parities two and three because of the small number
of births and deaths for parity four.
Furthermore, I do not show results for women with twelve or more years of
education for the 1972--84 period because of the low number of women.

The dependent variable is whether the child died within the first twelve
months of life.
The main set of explanatory variable consists of dummies for birth spacing
covering 12 months intervals, starting from nine months after the prior birth 
as above, until 48 months, which includes all births until 96 months.
In addition, I use dummies for sex of the index child and the sex composition 
of the prior children.
The birth spacing dummies, the sex of index child, and the sex composition dummies
are all interacted.
Since the actual number of abortions are unobserved, the interactions between
sex composition of prior children and the sex of the index child serve
as proxies for the use of sex selection.
The other explanatory variables are the same as above and estimations
are done separately by education level and parity.

I estimate the probability of infant mortality using a logit model.
Figures \ref{fig:mortality_low_med} and \ref{fig:mortality_high_highest} present 
the predicted probability of the second child dying within the first year at the 
possible combinations of index child sex, sex composition of prior children, and 
the birth spacing dummied.%
\footnote{
The online Appendix show the corresponding graphs for the third child.
}
All other variables are set to their average values.
To ease legibility all confidence intervals are suppressed.

An important caveat is that the estimations do not 
address potential selection problems.
For example, if women who have difficulties conceiving or carrying a pregnancy 
to term also have higher mortality risk for their offspring, a spurious correlation 
between long birth spacing and mortality may arise \citep{Kozuki2013}. 
Unfortunately, methods used previously to address selection, such
as family fixed effects, do not work well when the number of births is
low as it is for higher education women \citep{Kozuki2013,Molitoris2019}.
However, in prior research the fixed effects results and linear probability do not 
deviate substantially.

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\setcounter{subfigure}{-1}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \captionsetup[subfigure]{font=footnotesize,labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[No Eduction]{\includegraphics[width=\textwidth]{mortality_s2_p1_low_dummies}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{-0} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \captionsetup[subfigure]{font=footnotesize,labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[One to Seven Years of Education]{\includegraphics[width=\textwidth]{mortality_s2_p1_med_dummies}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_low_dummies}
    \end{minipage}
} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_low_dummies}
    \end{minipage}
} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_low_dummies}
    \end{minipage}
} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_med_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Second Child's Infant Mortality 
for Women with No Education
and Women with One to Seven Years of Education}
\label{fig:mortality_low_med}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\setcounter{subfigure}{-1}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \captionsetup[subfigure]{font=footnotesize,labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Eight to Eleven Years of Education]{\includegraphics[width=\textwidth]{mortality_s2_p1_high_dummies}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\begin{minipage}[t]{0.42\textwidth}\vspace{-2.68cm}\centering\footnotesize{Twelve or More Years of Education}\end{minipage}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_high_dummies}
    \end{minipage}
} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_highest_dummies}
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_high_dummies}
    \end{minipage}
} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_high_dummies}
    \end{minipage}
} 
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_highest_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Second Child's Infant Mortality 
for Women with Eight to Eleven of Education
and Women with Twelve or More Years of Education}
\label{fig:mortality_high_highest}
\end{figure}

There has been substantial convergence in mortality risk across groups over time.
For the 12 months and above intervals, there is now little difference across 
the education groups, with even the no education group showing an infant mortality risk 
below five percent.

Very short birth intervals still exhibit a higher mortality risk, although the 
effect declines with education level.
For the best-educated women the mortality risk is three to four percent,
whereas women with no education still show a risk that is close to ten percent.

Despite the prior findings of differential mortality by sex, there is little evidence 
that girls have substantially higher mortality risk.
There is some weak evidence that a boy born after a girl have lower mortality
risks in the earliest periods.
This difference disappears, however, with the general decline in mortality risk.

If sex selective abortions lead to higher mortality risk, boys born after a girl 
(the solid lines)
should have an increased risk with longer spacing for the two highest education groups 
in the last two periods.
There are, however, no apparent consistent differences between these groups and the other 
potential combinations.

% line 3661 in an_infant_mortality.log
The raw numbers for the group with the most uneven sex ratio---women with twelve or more
years of education who have no boys in the third spell in the last period---indicates
that even with high use of sex selection there is no impact on child mortality.
A total of 1,004 women in this group had a third child, of which 685 are boys.
Of these 685 boys, only 6 died within the first year of life and half of those
were in born the 0--23 months interval, and none in 48 months or more interval.

\section{Conclusion\label{sec:conclusion}}

% What has happened to spacing and why

One effect of the major changes that India has experienced over the last four decades is
that spacing between births have increased for everybody.
The increases in spacing are larger the higher the parity and the higher the mother's 
education and are larger in urban than in rural areas.

The most substantial increases in birth intervals arise, however, with heavy use of sex 
selection, as, for example, urban women with twelve or more years of education and two 
girls.
We, therefore, now in some cases observe a reversal of the traditional spacing
pattern where strong son preference was associated with shorter birth intervals the fewer 
sons a family has \citep{ben-porath76b,Leung1988}.
This reversal comes about because couples with a strong preference for sons now use sex 
selection, not because son preference has changed over time.

Despite the longer average birth intervals, a large number of women still have very
short birth intervals.
Hence, the changes in birth spacing is driven by the longer intervals becoming longer
over time, both for women who do not use sex selection, but especially for those who
do and end up with multiple abortions.
The exception to the small change in short spacing is the delay in conception for
women with very high sex selection use, presumably because this group knows that the
next child will be a boy.

% fertility
Fertility has declined for all groups, but the likelihood of having an additional child
still depends strongly on the number of sons, with women with no sons having the highest 
probabilities.

That those women most likely to use sex selection are also the most likely to have a
next birth, combined with the secular increases in spacing, makes the total fertility 
rate a more biased measure of cohort fertility.
This bias was most prominent early in the spread of sex selection, where the fertility
rate was up to one child lower than the predicted cohort fertility, but is still
present with the predicted cohort fertility ten to twenty percent higher than 
the fertility rate.

Tempo effects are studied extensively in the literature,%
\footnote{
See, for example, \citet{Bongaarts1999}.
}
but there are, however, to my knowledge, no other cases where there has been as 
substantial an increase in birth intervals and associated bias in fertility rates 
as for India.
It is, therefore, conceivable that we might see increases in TFR as birth spacing 
stabilizes or even declines again if interventions against sex selection are successful.

The only group who have a predicted cohort fertility below replacement are the 
best-educated urban women, with a predicted fertility of 1.8. 
Note that the predicted cohort fertility estimates come with the caveats that they are
based on married Hindu women and the model only includes births up to parity four.
However, almost everybody marry and Hindu constitute 80 percent of the population,
and since only the first fourth parities are used, it is possible that these 
estimates are lower-bound, especially for the less educated women.

% Mortality
There has been a substantial reduction in infant mortality over time with the size of
the reductions larger the less educated the mother.
This convergence means that there is now little difference in mortality risk across
education groups if the birth took place more than 21 months from the prior birth.
Short birth spacing is still more dangerous for the child, with the least increase for
the best-educated women.
Despite the potential that extensive use of sex selection might negatively affect the 
next child's survival, there is no evidence that children born to women who are the most
likely to use sex selection have worse outcomes. 
The rapid reduction in infant mortality risk, combined with the longer spacing,
which further reduces mortality, and higher than previously realized fertility, 
implies that population growth in India is, and will be, higher than expected.

% Future research and policy

The results here lead to two important questions that future research should address.

First, how does the changing birth spacing affect prior children?
We know that girls' health outcomes improve in the presence of sex selection 
\citep{Lin2014,Hu2015}.
An extensive use of sex selection implies is that only-girls families are much less 
likely to have a very short birth interval, which may reduce sibling competition for 
resources.
Hence, better health outcomes for girls could be an unintended side-effect 
of the longer spacing that arises from sex-selective abortions, rather than because the 
smaller number of girls makes them more valued as is often assumed.
A more positive reading would be that parents are more likely to view prior girls as
wanted since they know that they will eventually have a son.
In either case, an effective ban on sex selection could, at least temporarily, worsen
health outcomes for girls.

Second, is there an interaction between female labor force participation and 
sex selection?
Women may be staying out of the labor market exactly because the easier access to sex 
selection means that they are now more likely to have a boy and because the expected 
spacing between births has increased.
The higher return to male offspring's human capital combined with access to sex
selection may also explain another puzzle, why is the labor force participation so
low when fertility is low and education has increased.
As shown here, part of the answer may be that the number of children is, in fact, not as 
low as previously thought, both because of the increased spacing and that more 
children survive.

Better job opportunities for women may be a direct way to affect sex selection for two 
reasons.
First, it makes it more expensive to be out of the labor market for long periods.
Second, it would moderate the differential in potential earnings between husband and
wife and make it more attractive to invest in female compared to male offspring's human 
capital.
This approach may, however, be a double-edged sword.
If better job opportunities further lower desired fertility the result may be even
higher use of sex selection, everything else equal.
Hence, understanding the trade-off between long-term benefits from improvements in 
women's labor force participation and short-term costs from potential increases in sex 
selection is of paramount importance.

\clearpage

\onehalfspacing
\bibliographystyle{aer}
\bibliography{sex_selection_spacing}

\addcontentsline{toc}{section}{References}

\clearpage
\newpage

\appendix

% CHANGING NUMBERING OF FIGURES AND TABLES FOR APPENDIX
\renewcommand\thefigure{\thesection.\arabic{figure}}    
\renewcommand\thetable{\thesection.\arabic{table}}

\section*{Appendices for Online Publication}

These appendices are intended for online publication.
They provide the descriptive statistics, additional
estimated duration tables, and graphs for all 
education groups and spells.

\clearpage
\newpage

\section{Characteristics of Women's Work Experiences}

\setcounter{figure}{0}
\setcounter{table}{0}

Figure \ref{fig:work_by_survey} shows the percent of married women who are 
currently working at the time of the survey by age group and education level.
No other labor force participation question is consistently available across all 
four surveys. 
Because the question refers to currently working, the percentages are lower in previous
studies.

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_20} 
    \end{minipage}
}
\caption{Percentage of married women who were working at the time of the 
survey by age group and area of residence}
\label{fig:work_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_20} 
    \end{minipage}
}
\caption{Percentage of women paid cash or cash and in-kind of those women who were 
working at the time of the survey by age group and area of residence}
\label{fig:work_cash_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_20} 
    \end{minipage}
}
\caption{Percentage women who worked for a family member of those 
working at the time of the survey, by age group and area of residence}
\label{fig:work_family_by_survey}
\end{figure}

\clearpage
\newpage

\section{Recall Error and the Sex Ratio}

\setcounter{figure}{0}
\setcounter{table}{0}

The reliability of the results depends on the correctness of the birth histories
provided by the respondents.
A significant concern here is underreporting of child mortality, especially a systematic
recall error where respondents' likelihood of reporting a deceased child depends on the 
sex of that child. 
This Appendix section assesses the degree of recall error across the surveys and discusses 
methods to address it.

NFHS enumerators probe for any missed births, although the method depends on the survey.
NFHS-1 probe for each calendar birth interval that is four or more years.
NFHS-2 asked for stillbirths, spontaneous and induced abortions and also probed 
for each calendar birth interval four or more years.
NFHS-3 and NFHS-4 did not directly use birth intervals, but asked whether there were any 
other live births between (name of previous birth) and (name), including any children who 
died after birth, and asked for births before the birth listed as first birth and
after the last birth listed as the last birth.

Probing catches many initially missed births, but systematic recall error based on son
preference may still be a problem.
First, son preference leads to significantly higher mortality for girls than boys.
Secondly, son preference makes it more likely that parents will remember deceased boys 
than deceased girls.
Finally, in the absence of sex-selective abortions, parents with a preference for sons may
have the next birth sooner if the last child was a girl than if it was a boy.
If this girl subsequently dies, she is more likely to be missed if probing for missed 
births is only done for long intervals as in NFHS-1 and NFHS-2.

I use two approaches to examine the degree of recall error.
The first approach is to test whether the observed sex ratio is significantly different
from the natural sex ratio.
The natural sex ratio is approximately 105 boys to 100 girls or
51.2\% \citep{ben-porath76b,jacobsen99,Portner2015b}.
Prenatal sex determination techniques did not become widely available until the mid-1980s, 
so any significant deviation from the natural sex ratio before that time is likely the 
result of recall error.
The second approach is to compare births that took place during the same period but
where captured in different surveys.
Recall error is likely to increase with time, so births and deaths that took place earlier 
are more likely to be subject to recall error than more recent events.

Table \ref{tab:recallBirthBO1} shows the sex ratios of children recorded as first-born by 
year of birth, together with tests for whether the observed sex ratio is significantly 
higher than the natural sex ratio and whether more recent surveys have a higher sex ratio 
for the cohort than earlier surveys for the same period births.
Births are combined into five-year cohorts to achieve sufficient power.

\input{../tables/recallBirthBO1.tex}

% \input{../tables/recallBirthBO2.tex}

The ``first-born'' sex ratios illustrate the systematic recall error problem well.
In all four surveys around 55 percent of children reported as first-born are boys
for the first cohort of births observed.
Given that these cohorts cover from 1960-1964 to 1980-1984, which is before sex selection 
techniques became available in India, the most likely explanation for the skewed sex ratio 
is that some children listed as first-borns were not, in fact, the first children born in 
their families.
Instead, for a substantial proportion of families, their first-born was a girl who died 
and went unreported when enumerators asked about birth history.

As expected, the difference between the observed sex ratio and the natural sex ratio is 
less pronounced the closer to the survey date the cohort is.
The observed sex ratio for children born just before the NFHS-1 survey and listed as 
first-born is 0.517, which is not statistically significantly different from the
natural sex ratio.
The same general pattern holds for the other three surveys, with cohorts further away
from the survey date more likely to have a sex ratio skewed male.

% Second births show a pattern very similar to that for first births.
% An interesting difference is that there is evidence of a U-shaped relationship between
% time and sex ratio for second births.
% Cohorts furthest away from the survey year show the highest sex ratios, but declines to the
% natural sex ratio in the mid-1980s, and the sex ratio is then significantly higher again 
% for more recent births.
% This is in line with the results in the main paper that show that sex selective abortions 
% take place on lower parity births as desired fertility declined.

Finally, across surveys, the same cohort tends to show a higher sex ratio the more recent 
the survey (births in the cohort took place earlier relative to the survey date).
Despite this, few cohorts show significantly different sex ratios across surveys, most 
likely because of a lack of power.
The exception is that comparisons involving NFHS-4 are mostly statistically significant
since the number of surveyed households in NFHS-4 were much larger than in prior surveys.

The problem with the above approach is that the year of birth is affected by recall error; 
a second born child listed as first-born is born later than the real first born child.
Year of marriage should, however, be affected neither by parental recall error 
nor the use of sex-selective abortions.
Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2}, therefore, shows sex 
ratios of children recorded as first-born and second-born by year of parents' marriage, 
together with tests for whether the observed sex ratio is significantly higher than the 
natural sex ratio and whether more recent surveys show a higher sex ratio for the cohort 
than earlier surveys.
The basic recall error pattern remains, with women married longer ago more
likely to report that their first-born is a boy.
Similarly, comparing women married in the same five-year period across surveys shows
that women married longer ago are more likely to report having a son.

\input{../tables/recallMarriageBO1.tex}

\input{../tables/recallMarriageBO2.tex}

The relationship between the length of marriage and recall error can also be seen in 
Figures \ref{fig:sex_ratio_recall_rounds_bo1} and \ref{fig:sex_ratio_recall_rounds_bo2}, 
which show the observed sex ratio for children reported as first born as a function of 
the duration of marriage at the time of the survey.
The solid line is the sex ratio of children reported as first-born by the number of years 
between the survey and marriage, while the dashed lines indicate the 95 percent confidence 
interval and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes I group years into twos.
In line with the results from Tables \ref{tab:recallMarriageBO1} and 
\ref{tab:recallMarriageBO2}, the observed ratio of boys is increasingly above the expected 
value the longer ago the parents were married.

\begin{figure}
\centering
\subfloat[NFHS-1]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_1}}
\subfloat[NFHS-2]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_2}} \\
\subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_3}} 
\subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_4}} 
\caption{Ratio of Boys for ``First'' Births by Survey Round}
\label{fig:sex_ratio_recall_rounds_bo1}
\end{figure}

\begin{figure}
\centering
\subfloat[NFHS-1]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_1_bo2}}
\subfloat[NFHS-2]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_2_bo2}} \\
\subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_3_bo2}} 
\subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_4_bo2}} 
\caption{Ratio of Boys for ``Second'' Births by Survey Round}
\label{fig:sex_ratio_recall_rounds_bo2}
\end{figure}

The increasingly unequal sex ratio with increasing marriage duration suggests that
a solution to the recall error problem is to drop observations for 
women who were married ``too far'' from the survey year.
The main problem is establishing what the best cut-off point should be, with the
trade-off between retaining enough observations and the correctness of the information.
As Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2} show, there are 
differences in recall error across the three surveys and between the two birth
orders, although this may be the result of differences in the number of observations 
across surveys.
Furthermore, the recall error pattern is not entirely consistent across observed birth 
orders.
Since most of the surveys start showing significantly biased sex ratio from around 22
years of marriage on, I drop all observations where the marriage took place 22 years
or more.

\clearpage
\newpage

\section{Descriptive Statistics}
\setcounter{figure}{0}
\setcounter{table}{0}

% Descriptive statistics tables
\input{../tables/des_stat.tex}

\clearpage
\newpage

\section{Duration Results Tables}

\setcounter{figure}{0}
\setcounter{table}{0}

The first set of tables, Tables \ref{tab:avg_sex_ratio_low}, \ref{tab:avg_sex_ratio_med}, 
\ref{tab:avg_sex_ratio_high}, and \ref{tab:avg_sex_ratio_highest}, show predicted average 
birth intervals, sex ratios, and 
probabilities of having a birth by decade, spell, and sex composition for the three 
education levels separated by the area of residence, together with standard errors for all 
three outcomes.
The standard errors are based on bootstrapping for all three measures,
where the model is repeatedly estimated using resampling with replacement.

I also show whether durations for sex composition other than only girls are statistically 
significantly different from the duration with only girls based on bootstrapped 
differences. 
The cleanest test is comparing durations after only boys with durations after
only girls, but the number of births to women with only sons becomes small 
in the later periods.
Hence, it is possible to have substantial differences in spacing that are
not statistically significant because of low power, especially for the third 
and fourth spell.

Each predicted percent of boys is tested against the natural percentage of
boys using the bootstrapped standard errors.
The natural sex ratio is approximately 105 boys to 100 girls or
51.2\% \citep{ben-porath76b,jacobsen99,Portner2015b}.
The predicted percentage boys may differ from the natural rate because of 
natural variation, any remaining recall error not corrected for, or 
sex selection.

\input{../tables/bootstrap_duration_avg_sex_ratio_low_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_med_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_high_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_highest_all.tex}

\input{../tables/bootstrap_duration_p25_p75_low_all.tex}

\input{../tables/bootstrap_duration_p25_p75_med_all.tex}

\input{../tables/bootstrap_duration_p25_p75_high_all.tex}

\input{../tables/bootstrap_duration_p25_p75_highest_all.tex}

\clearpage

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_low_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_low_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_low_urban}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_med_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_med_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_med_urban}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_high_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_high_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_high_urban}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_highest_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_highest_urban}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
Changes in 25th Percentile Birth Intervals for Urban Women by Spell and Education
}
\label{fig:p25_urban}
\end{figure}

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_low_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_low_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_low_rural}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_med_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_med_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_med_rural}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_high_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_high_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_high_rural}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_highest_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_highest_rural}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
Changes in 25th Percentile Birth Intervals for Rural Women by Spell and Education
}
\label{fig:p25_rural}
\end{figure}

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_low_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_low_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_low_urban}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_med_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_med_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_med_urban}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_high_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_high_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_high_urban}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_highest_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_highest_urban}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
Changes in 50th Percentile Birth Intervals for Urban Women by Spell and Education
}
\label{fig:p50_urban}
\end{figure}

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_low_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_low_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_low_rural}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_med_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_med_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_med_rural}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_high_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_high_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_high_rural}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_highest_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_highest_rural}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
Changes in 50th Percentile Birth Intervals for Rural Women by Spell and Education
}
\label{fig:p50_rural}
\end{figure}

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_low_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_low_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_low_urban}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_med_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_med_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_med_urban}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_high_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_high_urban}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_high_urban}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_highest_urban}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_highest_urban}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
Changes in Average Birth Intervals for Urban Women by Spell and Education
}
\label{fig:avg_urban}
\end{figure}

\begin{figure}[htpb]
\captionsetup[subfigure]{labelformat=empty,position=top}
\captionsetup{font=footnotesize,skip=2pt}
\centering
\caption*{No education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_low_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_low_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_low_rural}}
\\
\caption*{One to Seven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_med_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_med_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_med_rural}}
\\
\caption*{Eight to Eleven Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_high_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_high_rural}}
\subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_high_rural}}
\\
\caption*{Twelve or More Years of Education}
\subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_highest_rural}}
\subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_highest_rural}}
\begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
\captionsetup{font=normalsize}
\caption{
Changes in Average Birth Intervals for Rural Women by Spell and Education
}
\label{fig:avg_rural}
\end{figure}

\clearpage
\newpage

\section{Survival Curves Conditional on Parity Progression for All Education and Spell Groups}

\setcounter{figure}{0}
\setcounter{table}{0}

% [FIRST SPELL]

% \subsection{First Spell}
% 
% \input{../figures/appendix_spell1_low.tex}
% 
% \input{../figures/appendix_spell1_med.tex}
% 
% \input{../figures/appendix_spell1_high.tex}
% 
% \begin{figure}[htpb]
% \centering
% \caption*{No Education}
% \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_low_urban_pps}} 
% \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_low_rural_pps}} \\
% \caption*{1-7 Years of Education}
% \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_med_urban_pps}} 
% \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_med_rural_pps}} \\
% \caption*{8 or more Years of Education}
% \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_highest_urban_pps}} 
% \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_highest_rural_pps}} 
% \caption{Survival curves conditional on progression to first birth; start point is month of marriage}
% \label{fig:results_spell1_pps}
% \end{figure}
% 
% 
% \clearpage
% \newpage
% 
\subsection{Second Spell}

% \input{../figures/appendix_spell2_low.tex}

% PARITY PROGRESSION SURVIVAL CURVES - Low

% Low education
\vspace*{-0.75cm}
\begin{figure}[hp!]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2005]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with no education by month beginning 9 months after prior birth.
}
\label{fig:results_spell2_low_pps}
\end{figure}

% SPELL 2 - URBAN - MEDIUM

% \input{../figures/appendix_spell2_med.tex}

% PARITY PROGRESSION SURVIVAL - MEDIUM

% Low education

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.32\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with 1-7 years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell2_med_pps}
\end{figure}

% High education

% \input{../figures/appendix_spell2_high.tex}

% PARITY PROGRESSION SURVIVAL - High education

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with eight to eleven years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell2_high_pps}
\end{figure}

% Highest education

% PARITY PROGRESSION SURVIVAL - Highest education

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_highest_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_highest_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_highest_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_highest_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with twelve or more years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell2_highest_pps}
\end{figure}

\clearpage
\newpage

\subsection{Third Spell}

% low education
% \input{../figures/appendix_spell3_low.tex}

% PARITY PROGRESSION SURVIVAL - LOW
\vspace*{-0.75cm}
\begin{figure}[hp!]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with no education by month beginning 9 months after prior birth.
}
\label{fig:results_spell3_low_pps}
\end{figure}

% Medium education

% \input{../figures/appendix_spell3_med.tex}

% PARITY PROGRESSION SURVIVAL - MEDIUM

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with 1 to 7 years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell3_med_pps}
\end{figure}

% High education

% \input{../figures/appendix_spell3_high.tex}

% PARITY PROGRESSION SURVIVAL - high

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with eight to eleven years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell3_high_pps}
\end{figure}

% Highest education

% PARITY PROGRESSION SURVIVAL - highest

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_highest_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_highest_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_highest_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_highest_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_highest_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_highest_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_highest_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_highest_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with twelve or more years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell3_highest_pps}
\end{figure}

\clearpage
\newpage

\subsection{Fourth Spell}

% low education

% \input{../figures/appendix_spell4_low.tex}

% PARITY PROGRESSION SURVIVAL - LOW
\vspace*{-0.75cm}
\begin{figure}[hp!]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_low_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_low_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with no education by month beginning 9 months after prior birth.
}
\label{fig:results_spell4_low_pps}
\end{figure}

% Medium education

% \input{../figures/appendix_spell4_med.tex}

% PARITY PROGRESSION SURVIVAL - MEDIUM

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_med_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_med_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with 1 to 7 years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell4_med_pps}
\end{figure}

% High education

% \input{../figures/appendix_spell4_high.tex}

% PARITY PROGRESSION SURVIVAL - high

\begin{figure}[htpb]
\centering
\caption*{Urban}
\setcounter{subfigure}{-1}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_urban_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{-0}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{1}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{2}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_high_urban_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption*{Rural}
\setcounter{subfigure}{3}
\subfloat[1972--1984]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_rural_pps}} 
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} 
\setcounter{subfigure}{4}
\subfloat[1985--1994]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
} \\
\setcounter{subfigure}{5}
\subfloat[1995--2004]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\setcounter{subfigure}{6}
\subfloat[2005--2016]{
    \begin{minipage}{0.31\textwidth}
        \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
        \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_high_rural_pps}}
        \captionsetup[subfigure]{labelformat=parens}
    \end{minipage}
}
\caption{Survival curves conditional on parity progression
for women with eight to eleven years of education by month beginning 9 months after prior birth.
}
\label{fig:results_spell4_high_pps}
\end{figure}

\clearpage
\newpage

\section{Infant Mortality Graphs}

\setcounter{figure}{0}
\setcounter{table}{0}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_low_dummies} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_low_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_low}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_med_dummies} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_med_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with one to
seven years of education}
\label{fig:mortality_med}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_high_dummies} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_high_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with eight
to eleven years of education}
\label{fig:mortality_high}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_highest_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with twelve or
more years of education}
\label{fig:mortality_highest}
\end{figure}

\end{document}