### 2020-07-30 paper/sex_selection_spacing-ver4.tex
Major editing of birth interval section(s) plus other edits
- remove delay effect discussion
- comment out conditional survival curves
- Reorganize birth interval discussion
- Remove average discussion of Estimation Strategy section

\documentclass[12pt,letterpaper]{article}

\usepackage{amsmath}
\usepackage{fontspec}
\setromanfont[Ligatures=TeX]{TeX Gyre Pagella}
\usepackage{unicode-math}
\setmathfont{TeX Gyre Pagella Math}
\usepackage[title]{appendix}
\usepackage[margin=1.0in]{geometry}
\usepackage[figuresleft]{rotating}
\usepackage[longnamesfirst]{natbib}
\usepackage{dcolumn}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[flushleft]{threeparttable}
\usepackage{setspace}
\usepackage[justification=centering,skip=2pt]{caption}
\usepackage[font=scriptsize]{subfig}
\usepackage[xetex,colorlinks=true,linkcolor=black,citecolor=black,urlcolor=black]{hyperref}
\usepackage{adjustbox}
\usepackage{xfrac}
\usepackage{placeins}

% For word counting uncomment line below and raggedright below
% http://countwordsfree.com/#
% \usepackage[nolists]{endfloat}

% \bibpunct{(}{)}{;}{a}{}{,}
\newcommand{\mco}[1]{\multicolumn{1}{c}{#1}}
\newcommand{\mct}[1]{\multicolumn{2}{c}{#1}}
\newcommand{\mcth}[1]{\multicolumn{3}{c}{#1}}
\newcommand{\X}{$\times$ }
\newcommand{\hs}{\hspace{15pt}}

% Attempt to squeeze more floats in
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.1}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

%------------------------------------------------------------------------

\title{Birth Spacing in the Presence of Son Preference and Sex-Selective Abortions:
India's Experience over Four Decades%
\protect\thanks{%
I am grateful to Andrew Foster and Darryl Holman for discussions about the method.
I owe thanks to Shelly Lundberg, Daniel Rees, David Ribar, 
Hendrik Wolff, seminar participants at University of Copenhagen, University of Michigan, 
University of Washington, University of Aarhus, the Fourth 
Annual Conference on Population, Reproductive Health, 
and Economic Development, and the Economic Demography Workshop for helpful 
suggestions and comments.
I would also like to thank Nalina Varanasi for research assistance.
Support for development of the method from the University of Washington Royalty 
Research Fund and the Development Research Group of the World Bank is gratefully 
acknowledged.
The views and findings expressed here are those of the author and
should not be attributed to the World Bank or any of its member countries.
Partial support for this research came from a Eunice Kennedy Shriver National
Institute of Child Health and Human Development research infrastructure grant,
5R24HD042828, to the Center for Studies in Demography and Ecology at the
University of Washington.
}
}

\author{Claus C P\"ortner\\
    Department of Economics\\
    Albers School of Business and Economics\\
    Seattle University, P.O. Box 222000\\
    Seattle, WA 98122\\
    \href{mailto:cportner@seattleu.edu}{\texttt{cportner@seattleu.edu}}\\
    \href{mailto:claus@clausportner.com}{\texttt{claus@clausportner.com}}\\
    \href{http://www.clausportner.com}{\texttt{www.clausportner.com}}\\
    \& \\
    Center for Studies in Demography and Ecology \\
    University of Washington\\ \vspace{2cm}
    }

\date{July 2020}

%------------------------------------------------------------------------

\begin{document}
\graphicspath{{../figures/}}
\DeclareGraphicsExtensions{.eps,.jpg,.pdf,.mps,.png}

\setcounter{page}{-1}
\maketitle
\thispagestyle{empty}

% \setcounter{page}{0}

\newpage
\thispagestyle{empty}
\doublespacing

\begin{abstract}

% Demography abstract
\noindent 
Using four rounds of India's National Family and Health Surveys and a
competing-risk hazard model, I show that Hindu women's average birth intervals
increased over the last four decades for all education groups. The most
significant increases are among the women most likely to use sex
selection. Despite the rise in average intervals, the likelihood of very
short spacing did not change substantially. Hence, the increases come
predominantly from the longer birth intervals getting even longer. As a
result of the longer spacing, fertility rates significantly
overestimated how fast cohort fertility fell. Although cohort fertility
and the fertility rate have started to converge, the cohort fertility is
still substantially higher than the fertility rate. Furthermore, cohort
fertility is still at or above replacement level for all but the
best-educated urban women. Finally, infant mortality risk has declined
substantially over time for all groups, but fastest for the lower
education groups, who are now close to the level of women with the most
education. Short birth spacing is still associated with higher
mortality, although the difference is small for the best-educated women.
There is no evidence that the increasing use of sex selection is
associated with higher infant mortality risk.

\noindent JEL: J1, O12, I1
\noindent Keywords: India, prenatal sex determination, censoring, competing risk, non-proportional hazard
\end{abstract}

\newpage

%------------------------------------------------------------------------

\section{Introduction\label{sec:intro}}

% For word counting remove comment on the line below
% \raggedright

% Background
India experienced many positive changes over the last four decades:
the economy grew substantially,
schooling increased for males and females,
and the total fertility rate fell to 2.2 
\citep{Bosworth2008,Dharmalingam2014,
International-Institute-for-Population-Sciences-IIPS2017}.

However, India also witnessed the advent of sex selection---the selective abortion of 
female fetuses based on prenatal sex determination---with the introduction of ultrasound 
in the mid-eighties.
Combined with continued strong son preference, the result was a dramatic increase in the 
male-to-female ratio at birth
\citep{das_gupta97,Arnold2002,retherford03b,Guilmoto2012,Portner2015b,Jayachandran2017}.

% Questions addressed and why valuable
The main question here is: how did birth spacing respond to these changes, especially 
the spread of sex selection? 
The motivation is twofold.
First, prior research has failed to appreciate that each abortion increases the
interval between births by six to twelve months.%
\footnote{
The increase consists of three parts. 
First, after an abortion, the uterus needs at least two menstrual cycles to recover, 
or the likelihood of spontaneous abortion increases substantially \citep{zhou00b}. 
Second, the waiting time to conception is one to six months. 
Finally, sex determination tests are reliable only from three months of gestation. 
}
Hence, the growing use of sex-selective abortions may substantially increase birth spacing, 
but we know little about how much. 
Second, more female education, higher household income, and low and declining female labor 
force participation all likely impact birth spacing. 
The combined changes in birth spacing may, therefore, outpace what we have observed in 
other countries.

Studying birth spacing contributes to our understanding of fertility decisions, but, equally 
important, birth spacing affects the reliability of our fertility measures and may 
impact mortality. 
To understand these effects, I address two additional questions.

First, did changes in birth spacing bias the fertility estimates for India? 
With longer spacing, mothers will be older at each parity, and this tempo effect makes the 
total fertility rate a downward biased estimate of cohort fertility
\citep{Hotz1997,Bongaarts1999,Ni-Bhrolchain2011}. 
Hence, if birth intervals increased substantially, India's fertility may be higher 
than generally accepted.

Second, what is the relationship between infant mortality and the changes in birth spacing 
and sex selection?
In India, birth intervals have traditionally been shorter with fewer sons, contributing 
to the higher mortality risk for girls
\citep{Whitworth2002,Bhalotra2008,Maitra2008,Jayachandran2011,Jayachandran2017a}.
Hence, longer birth spacing, whether from sex selection or secular changes, may reduce 
mortality \citep{Conde-Agudelo2012,Molitoris2019}.
However, a counteracting effect is possible if longer birth intervals arise from multiple 
abortions since the short duration between \emph{pregnancies} could increase mortality. 
Nevertheless, there has been no attempt to examine how infant mortality changed with 
the increase in sex selection.

% How the study progresses
To investigate how birth spacing has changed, I use a competing risk hazard model with 
two exit states: the birth of a girl or the birth of a boy. 
I apply the model to the birth histories of Hindu women covering 1972 to 2016 using 
data from the four National Family and Health Surveys (NFHS).

The primary outcomes I examine are average birth intervals, the sex ratio of births, 
and the likelihood of giving birth. 
Because sex selection might also affect the distribution of birth intervals, I furthermore
examine the 25th and 75th percentiles of birth intervals. 
I estimate the model across four periods to capture the changing access and 
legality of sex selection. 
The key variables are maternal education, the sex composition of previous 
children, and area of residence.

The empirical model allows me to predict cohort fertility. 
To examine whether tempo effects bias our standard fertility measures, I compare 
the predicted cohort fertility with fertility calculated from age-specific fertility 
rates.

I use the same data to study how infant mortality changed with birth spacing and 
the increasing use of sex selection. 
The key explanatory variables remain the same, except for the addition of birth spacing 
and the sex of the index child.

% Main results
There are four main results.

First, average birth intervals lengthen over the four decades, more so the higher the 
parity and the more educated the woman.
Women most likely to use sex selection---well-educated with no sons---had the most 
substantial lengthening of birth intervals and the most biased sex ratios. 
As a result, we now see cases that reverse the traditional spacing pattern, with some 
women with no sons having \emph{longer} birth intervals than those with sons.
Those least likely to use sex selection---less-educated in rural areas---still follow 
the traditional pattern of short spacing when they have girls and little evidence of sex 
selection.

Second, the likelihood of a very short birth interval changed little. 
Hence, the increased average spacing comes mainly from longer birth intervals getting even 
longer, primarily because of sex selection. 
The exception is that those better-educated women with a very high predicted use of sex 
selection conceive later and are less likely to have a very short birth interval, precisely 
because their next child is almost certainly a boy.

Third, the fertility rate substantially overestimated how fast cohort fertility fell in the 
1990s and early 2000s as spacing began to increase. 
Although the two have lately been converging, the predicted cohort fertility is still ten to 
twenty percent higher than the fertility rate. 
Furthermore, predicted cohort fertility is still at or above replacement level for all but 
the best-educated urban women.

Finally, infant mortality has declined substantially over time for all groups, but fastest 
for the less educated, who are now close to the level of the best-educated women. 
However, mortality is still inversely related to education level, especially for very short 
birth intervals. 
There is no evidence that the increasing use of sex selection is associated with higher mortality.

%------------------------------------------------------------------------------------

% Combine descriptive and theory sections into one

\section{Background and Conceptual Framework}

To set the stage for the subsequent analyses and provide a conceptual framework for 
understanding birth spacing, I first discuss female education and labor force 
participation in India and relevant theories on birth spacing.

% [Female education]

Female education is a crucial explanatory variable here for three reasons.
First, higher female education is associated with lower fertility and increased use of 
sex selection
\citep{das_gupta97,dreze01,bhat03,retherford03b,Guilmoto2009a,Portner2015b,Jayachandran2017}.
Second, female labor force participation in India, as in other developing countries, 
first decreases and then increases with education
\citep{Klasen2015,Fletcher2017,Afridi2018,Bhargava2018,Chatterjee2018,Bhargava2019}.
Finally, child mortality decreases with maternal education
\citep{Rosenzweig1982a,Whitworth2002,Maitra2008}.

% [TK where to put son preference discussion?]
% %
% \footnote{
% It is also possible that the effect is driven by a stronger son preference, although 
% higher education women have shown declining son preference \citep{bhat03,pande07}.
% }

I divide eduction into four groups: no education, one to seven years, eight to eleven 
years, and twelve and more years.
The latter two correspond to having completed primary and completed secondary school, 
respectively.%
\footnote{
Although there are variations by state, elementary education in India consists of a 
primary school covering grades one through five and an upper primary---or middle 
school---covering grades six through eight. 
Similarly, secondary education covers grades nine and tenth for ``secondary education'' 
and eleven and twelve for ``upper secondary.''
}
To ensure that the results are comparable with the prior literature on fertility
and mortality in India, I follow the NFHS reports, except that I combine the 
less than five years and five to sevens years of schooling completed and the eight to nine 
and ten to eleven years of schooling completed.
There is a trade-off between having groups large enough for the analyses below and having 
groups detailed enough to capture differences in behavior across education levels, such as 
the U-shaped relationship between education and labor force participation.

% [what has happened to education]
Female education has increased substantially over time in rural and urban areas but 
is still substantially higher in urban than in rural areas.
Figure \ref{fig:education_over_time} shows the distribution of schooling by birth 
cohort for urban and rural women, 
twenty years or older, whether married or not, based on the four rounds of the NFHS.
In rural areas, women with no education have gone from 90 percent for the 1930s cohorts to 
less than 20 percent for the 1990s cohorts.
Women with eight or more years of education have gone from almost zero for the 1930 cohort 
to more than sixty percent for the 1990s cohorts, with about half in the eight to eleven 
group and the other half in the twelve plus group.
In urban areas, the proportions with no education or one to seven years have declined 
to just below ten percent each.
Most of the increase in urban female education came from the twelve plus group, which 
now account for more than half of all urban women for the lastest cohorts.

\begin{figure}[htpb]
\centering
\subfloat[Rural]{
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth]{educ_over_time_rural}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.49\textwidth}
        \includegraphics[width=\textwidth]{educ_over_time_urban} 
    \end{minipage}
}
\caption{Distribution of education by cohort for women twenty years or older at survey}
\label{fig:education_over_time}
\end{figure}

% [female labor force participation]

The standard economic argument for spacing becoming shorter with increasing female education 
is that parents incur time costs when they have children \citep{Hotz1997,schultz97}.
Specifically, if having children require the mother to reduce her market work, parents 
can lower the cost of children by shortening birth spacing to 
take advantage of economies of scale in childrearing \citep{Vijverberg1982}.

However, even as the level of female education has increased, the female labor force 
participation in both urban and rural areas has stagnated or decreased
\citep{Klasen2015,Fletcher2017,Afridi2018,Bhargava2018,Chatterjee2018,Bhargava2019}.
India's female labor force participation is now lower than most other countries and does 
not yet show any signs of increasing \citep{Klasen2015,Chatterjee2018}.
In line with previous research, the NFHS data show a U-shaped relationship between 
education and working for married women, with the highest percent working for women with 
either no education or twelve or more years and the lowest for women with eight to 
eleven years of education.%
\footnote{
See Appendix Figures \ref{fig:work_by_survey} through \ref{fig:work_family_by_survey}. 
}

The low and declining female labor force participation, especially for younger women, 
suggests that families face little incentive to space children more closely 
together for economic reasons.
One explanation is that household income has increased so substantially that the income 
effect dominates any substitution effect.
Two findings speak for this effect.
First, although real wages for both men and women have almost doubled between 1987 and 
2011, the mean male wage is still close to 70 percent higher than the female wage 
\citep{Klasen2015,Bhargava2018}.
Second, women’s labor supply appears to be more negatively elastic to husbands' 
wages than positively elastic to own wages \citep{Bhargava2018}.

% [sex selection]

The introduction of sex selection allows parents to avoid giving birth to girls but 
increases the expected duration to the next birth.
Theory suggests that sex selection increases with lower desired fertility and with
higher parity for a given desired number of children \citep{Portner2015b}.
Sex selection is more widespread among better-educated than less-educated and among urban 
than rural women, which is consistent with lower desired fertility increasing the
use of sex selection 
\citep{das_gupta97,retherford03b,Guilmoto2009a,Portner2015b,Jayachandran2017}.
Better-educated and urban women also tend to live in households with higher income, which 
better enables them to access prenatal sex determination and lowers the relative costs of 
using sex selection and have long birth intervals.

The combination of rising incomes and continued son preference may lead to even longer 
spacing than expected from the income effect alone.
As women's education increases, their productivity in the production of offspring human 
capital also increases.
With relatively more boys born because of increased access to sex-selective 
abortions and the increasing income potential for (male) offspring, demand for 
better-educated women can increase, even if they do not participate in the labor market 
\citep{Behrman1999}.

If more and ``better'' parental attention per child results in higher child ``quality,'' 
we should expect longer birth intervals \citep{Zajonc1975,Zajonc1976,Razin1980}.
However, the evidence on spacing's effect on child quality measures such as IQ 
and education is mixed for developed countries and non-existing for developing countries
\citep{Powell1993,Pettersson-Lidbom2009,Buckles2012,Barclay2017}.
The exception is one crucial aspect of child quality---health and mortality---where
longer spacing does lead to better outcomes, although this relationship weakens with 
maternal education \citep{Whitworth2002,Conde-Agudelo2012,Molitoris2019}.

% If parents realize that longer spacing leads to better child outcomes, we may even observe 
% later conception, especially when, with sex selection, the next child is likely to be a boy.
% Working in the opposite direction of this "delay" effect is that women with more education 
% can, in principle, space their children closer together without substantially increasing 
% child mortality risk. 
% However, given the apparent lack of pressure to return to the labor force, this effect is 
% likely small.

% Sanskritization

The increases in female educational attainment imply that access to education has 
expanded beyond the higher castes. 
One possible effect of the associated change in the composition of better-educated women 
is that this group's behavior would change.
However, ``Sanskritization'' implies that as lower castes females gain access to 
education and their husbands' income increases, they adopt higher caste 
norms such as stronger son preference and a retraction from the formal labor 
market \citep{Srinivas1956,Chen1995,Abraham2013,Chatterjee2018}.
The low and declining female labor force participation is an indication that this
process still operates.

% Summary / hypotheses

In summary, with substantial increases in husbands' income and a declining female labor 
force participation, I expect a push toward longer birth spacing over time, independent
of education levels, based on the income effects and the effects of spacing
on child outcomes.
Furthermore, I expect birth spacing to increase the most among the better-educated 
because their household income increases the most---even with declining female labor 
participation---and their use of sex selection.
Even with the substantial increase in the number of better-educated women, 
"Sanskritization" implies that the changing composition does not substantially change 
this group's use of sex selection.

%------------------------------------------------------------------------------------

\section{Estimation Strategy\label{sec:strategy}}

% Requirements for method and the intuition behind each:  
%  - Account for censoring (hazard model)
%  - Work both with and without sex selection (competing risk part)
%  - Allow for changes in use of sex selection within spell (flexible specification of baseline hazard ?)
%  - Capture that the shape of the hazard function differ across groups
%    (non-proportionality); needs a better description, maybe using example
%    of shorter spacing and/or use of sex selection. Differences in parity
%    progression likelihood.

The standard approach in the birth spacing literature is to use proportional hazard
models with a single exit---the birth of a child.%
\footnote{
See \citet{Sheps1970} and \citet{Newman1984} for early discussions of why 
hazard models are the preferred way to deal with the censoring of birth
intervals.
}
There are two problems with the standard approach in this setting.

First and foremost, the use of sex selection means that the sex of the next child is 
no longer random and that the spacing to the birth of a boy will differ from the 
spacing to a girl. 
Therefore, I use a competing risk set-up that captures both the non-randomness of 
the birth outcome and the differential spacing.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

Second, even without sex selection, it is unlikely that characteristics, such as 
the sex composition of previous births, have the same effects throughout the entire 
birth interval. 
The proportional hazard model requires that the hazard for any individual is a 
fixed proportion of the hazard for any other individual. 
That the effects are not constant violates that assumption, and the results from a 
proportional hazard model would, therefore, be biased. 
The proportionality assumption is especially problematic for higher-order birth 
intervals since there are substantial differences across groups in the likelihood 
of progressing to the next birth and how soon couples want their next child if they 
are going to have one \citep{Whitworth2002,Bhalotra2008,Kim2010}.

The introduction of prenatal sex determination exacerbates any bias from
the proportionality assumption for two reasons. 
First, different groups have different levels of sex-selective abortion use and, thereby, 
birth spacing. 
Second, within a birth interval, a household's use of sex selection may vary, and that
means that the effects of covariates vary as well.

Therefore, I use a non-proportional hazard specification that allows the shape of the 
hazard functions to vary across groups.
The use of a non-proportional specification also mitigates any potential effects 
of unobserved heterogeneity when used in conjunction with a flexible baseline hazard 
\citep{Dolton1995}.

The model is a discrete-time, non-proportional, competing risk hazard model with two 
exit states: either a boy or a girl is born. 
The unit of analysis is a spell---the period from one parity birth to the following 
birth or censoring. 
For estimation purposes, the spells begin nine months after the previous birth since 
this is the earliest we should expect to observe a new birth. 
Censoring can happen for three reasons: the survey takes place, sterilization of the 
woman or her husband, or imposed because of too few births for the method to work.

For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$, and 
the spell continues until time $t_i$ when either birth or censoring of the spell 
occurs.
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
There are two exit states: the birth of a boy, $j=1$, or the birth of a girl, $j=2$.

The discrete-time hazard rate $h_{ijt}$ is 
\begin{equation}
 h_{ijt} = \frac{\exp(D_j(t) + \alpha_{jt}'\mathbf{Z}_{it} + \beta_j'\mathbf{X}_{i})} 
 {1 + \sum_{l=1}^2 \exp(D_j(t) + \alpha_{lt}'\mathbf{Z}_{it} + \beta_l'\mathbf{X}_{i})} \: \: \; \; \;  j = 1,2.
 \label{eq:hazard}
\end{equation}
$D_{j}(t)$ is the piece-wise constant baseline hazard for outcome $j$, captured
by dummies and the associated coefficients,
\begin{equation}
D_j(t) = \gamma_{j1} D_1 + \gamma_{j2} D_2 + \ldots + \gamma_{jT} D_T,
\end{equation}
with $D_m = 1$ if $t=m$ and zero otherwise.
This approach to modeling the baseline hazard is flexible and does not restrict the 
baseline hazard unnecessarily.
$\mathbf{Z}$ is the non-proportional part, which includes the interactions between 
$D_j(t)$ and a set of explanatory variables and the interactions of those.
The remaining explanatory variables, $\mathbf{X}$, enter proportionally.

% To further minimize any potential bias from assuming proportionality, estimations 
% are done separately for different levels of mothers' education and different periods.
% I discuss the choice of which variables to use for non-proportionality below.

Equation (\ref{eq:hazard}) is equivalent to the logistic hazard model and has the same 
likelihood function as the multinomial logit model \citep{allison82,jenkins95}.
Hence, splitting spells into smaller intervals---here equal to three months---and 
treating them as observations, I can estimate the model using a standard multinomial 
Logit model. 
In the reorganized data, the outcome variable is 0 if the woman does not have a child 
in a given interval (the base outcome), 1 if she gives birth to a son in that interval, 
and 2 if she gives birth to a daughter in that interval.

I use the model to predict birth interval measures, parity progression probabilities, 
and the sex ratio rather than present coefficients since the interpretation 
of competing risk model coefficients is challenging \citep{thomas96}.
The predicted parity progression probability is the likelihood of giving birth by the
imposed censoring based on standard survival curve calculations averaged across all
women in a given sample.

For birth intervals, I estimate a set of percentile birth intervals.
To find a given percentile birth interval, I first calculate for each woman when there is 
a given percent chance that she will have given birth, conditional on the probability of 
giving birth in the spell.
I then add nine months to account for the start of the spell. 
For example, with an 80 percent chance of giving birth by the end of the spell, the median birth 
interval is the predicted number of months before a woman passes the 40 percent mark on her 
survival curve. 
The reported statistic is the average of a given percentile interval across all women in 
a given sample using the individual progression probabilities as weights.

The predicted sex ratio is the weighted average of individual predicted sex ratios, using 
parity progression probabilities as weights.
For an individual woman, the estimated percent of births that are boys at $t$, $\hat{Y_t}$, is:  
\begin{equation}
\hat{Y}_t 
= 
\frac{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )}
{ P(b_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t) + P(g_{t} | \mathbf{X}_{k}, \mathbf{Z}_{kt},t )} 
\times 100,
\label{eq:probability_son}
\end{equation}
where $b$ and $g$ are the predicted probabilities of having a boy and a girl, respectively,
conditional on the explanatory variables and not having had a child before $t$.
Weighting the percentages boys with the likelihood of exiting the spell with a birth across all 
$t$ gives the predicted percent of boys over the entire spell for an individual.%
\footnote{
Imagine $T=2$. 
If 54 and 66 percent of births are boys and the likelihood of giving birth 20 and 40 percent, 
then the predicted sex ratio is $\frac{54*0.2+66*0.4}{0.2+0.4} = 62$ percent boys. 
}

\section{Data\label{sec:data}}

The data come from the four rounds of the NFHS
collected in 1992--1993, 1998--1999, 2005--2006, and 2015--2016.
The surveys are large: 89,777, 90,303, 124,385, and 699,686 women, respectively. 
NFHS-1 and NFHS-2 surveyed only ever-married women, while the two later surveys included 
never-married.

I focus on the three birth intervals starting from the first birth and ending with the 
fourth birth.
I exclude the interval from marriage to the first birth because many are imputed and the
higher-order intervals because few women had five or more births, especially among the 
better-educated.

I restrict the sample to Hindus for two reasons.
First, Hindus are the majority population group, about 80 percent of India’s population.
Second, the prior literature shows that son preference and use of sex selection vary 
substantially between Hindus and the second largest group, Muslims. 
Combining them and assuming that the baseline hazard is the same would lead to biased 
results.
Because of space constraints and the relatively small number of observations once split 
by education and periods, I do not provide separate results for Muslims or any of the 
remaining groups, such as Sikhs, Jains, and Christians.

Finally, I exclude visitors and women in any of the following categories: never married; 
no gauna yet; married more than once; divorced; not living with husband; inconsistent age 
at marriage; or education information missing.  
The same goes for women who had at least one multiple births, reported giving birth 
before age 12, had a birth before marriage, or an interval between births of less than 
nine months.

In addition to a large number of women surveyed and the long period covered, a significant 
benefit of the NFHS over other surveys is that enumerators pay careful attention to the 
spacing between births and probe for ``missed'' births.
For India, the main concern is underreporting of deceased children, especially a systematic 
recall error where respondents' likelihood of reporting the birth of a deceased child depends 
on the sex of that child.
Unreported deceased children inflate the birth intervals and, with declining mortality, 
make changes over time appear too small.
I, therefore, provide a detailed analysis of systematic recall error in the Online Appendix.
Recall error depends heavily on how long ago a woman was married, and I, consequently, 
drop women married 22 years or more.%
\footnote{
Recall error is likely behind the designation of the first two rounds of NFHS as ``moderate 
quality'' in an analysis of the quality of birth histories in DHS surveys and its impact 
on fertility estimates \citep{Schoumaker2014}.
}

To ensure that there are enough births for the method to work, I censor each spell at 96 
months (eight years) after a woman can first give birth, which is equivalent to 105 months 
after the birth of the prior child.
Less than one percent of observed births occur after the cut-off.
The final sample consists of 
\input{../tables/num_women.tex}
% [number] women, with [number] parity one through four births.

% %
% \footnote{
% \input{../tables/num_missed.tex}
% }

Direct information on the use of sex selection is not available, so I compare periods 
based on the changes in access and legality of prenatal sex determination in India.
Abortion has been legal in India since 1971.
Reports of sex determination appeared around 1982--83, and the number of clinics 
quickly increased \citep{Sudha1999,bhat06,Grover2006}.
In 1994, the Prenatal Diagnostic Techniques Act made determining and communicating 
the sex of a fetus illegal.%
\footnote{
% Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
There is little evidence that the ban significantly affected sex ratios \citep{Das-Gupta2016}.
}
Finally, although sex selection increased even after 1994, we may have passed a 
turning point in its use in the mid-2000s \citep{Das_Gupta2009,Kumar2012,Bongaarts2013,Diamond-Smith2015}.

I use four periods: 1972--1984, 1985--1994, 1995--2004, and 2005--2016.
The first covers before sex selection became available and the second from when sex 
selection became available until the Prenatal Diagnostic Techniques Act.
I have split the period from 1995 until 2016 into two to examine if there is support for 
the prior literature's hypothesized reversal in child sex ratios and son 
preference in India.

The allocation of spells into periods is determined by when conception, and, therefore, 
decisions on sex selection, can begin. 
Hence, some spells cover two periods, which may bias downward the differences between the 
periods.
Most sterilizations take place right after a woman gives birth, and,
therefore, do not show up in the samples used. 
Furthermore, sterilization depends strongly on the sex composition of prior children
with lower probabilities, the fewer boys. 
The effect is to bias downward the differences in parity progression probabilities.

I divide the explanatory variables into two groups.
The first group consists of characteristics that the prior literature finds affect 
the spacing choice and the use of sex selection:
mother's education, sex composition of previous children, and area of residence.%
\footnote{
With sex selection, the composition of prior children is, in principle, endogenous.
It is beyond the scope of this paper to develop a method for dealing with this issue.
}
The second group of variables consists of those expected to have an approximately 
proportional effect on the hazard.
These include the mother's age when the spell begins, the household's
land ownership, and whether it belongs to a scheduled tribe or caste.
Appendix Tables \ref{tab:des_stat1} and \ref{tab:des_stat2} present descriptive statistics.

I estimate a separate model for each birth interval, education group, and period 
combination to minimize any potential bias from including proportional variables.
I capture sex composition with dummy variables for the
possible combinations, ignoring the ordering of births.
Area of residence is a dummy variable for living in an urban area.

\section{How Birth Spacing Changed\label{sec:results}}

% Explain question and what I am doing
The first question I address is how birth spacing responded to the significant changes, 
especially the spread of sex selection, that India experienced. 
I estimate the competing risk hazard model for each spell, education, and period 
combination and then predict 25th, 50th, and 75th percentiles birth intervals in 
months, the sex ratio, and the probability of having a birth in that spell.
Sex selection increases all three percentile birth intervals, but the effects of abortions 
are easiest to observe in the 75th percentile birth interval, 
and it is also the changes in the longest birth intervals that will likely the largest 
impact on fertility measures.
The 25th percentile captures changes in the short spacing, which is the most likely to 
have an adverse impact on health.

Figures \ref{fig:spacing_low_rural} through \ref{fig:spacing_highest_urban}
show the predicted outcomes for the four education levels by area of residence.
The underlying values with bootstrapped standard errors are available in the online 
appendix.
The sex ratio graphs also show the natural sex ratio, approximately 51.2 percent 
\citep{ben-porath76b,jacobsen99,Portner2015b}.
Results for urban women without education, rural women with twelve or more years of 
education, and the fourth spell for women with twelve or more years of education are
presented in the online appendix because they are based on relatively few observations.

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_low_rural}
\caption{combined graphs - low rural}
\label{fig:spacing_low_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_low_urban}
\caption{combined graphs - low urban}
\label{fig:spacing_low_urban}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_med_rural}
\caption{combined graphs - med rural}
\label{fig:spacing_med_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_med_urban}
\caption{combined graphs - med urban}
\label{fig:spacing_med_urban}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_high_rural}
\caption{combined graphs - high rural}
\label{fig:spacing_high_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_high_urban}
\caption{combined graphs - high urban}
\label{fig:spacing_high_urban}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_highest_rural}
\caption{combined graphs - highest rural}
\label{fig:spacing_highest_rural}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_highest_urban}
\caption{combined graphs - highest urban}
\label{fig:spacing_highest_urban}
\end{figure}

The results show three broad trends.
% [Birth intervals]
First, 
over the four decades, birth intervals have lengthen, whether measured by the
25th, 50th, or 75th percentile.
These increases are larger the higher the education of the mother, the higher the parity, 
and the higher the percentile measure.

% [Parity progression probability]
Second, 
the likelihood of a next birth has fallen over time, and the declines are
larger the higher the education, the higher the parity, and with the presence of at
least one son. 
Within a given spell and period, parity progressions are lower in urban than in rural areas, 
if at least one son is present, and the more educated the mother.

% [Sex ratios]
Finally, 
the sex ratios of next births has become more male dominated for women with no
sons, while there are no clear trends for the other sex compositions.
The increases in the predicted percent boys are larger the higher the education and the 
higher the parity.
Furthermore, for a given spell and period, sex ratios in the absence of a son are higher 
in urban than in rural areas, the higher the education of the mother, and the higher
the parity.
There is some evidence of unnaturally high percentage boys born even when women have one 
son, although most are clearly lower than in the absence of a son and the failing fertility
make more of those estimate noisy.

% [25th percentile birth intervals]
Despite the increases in birth intervals and the decline in fertility, a remarkably high 
proportion of birth intervals are still very short.
For all but the most educated, more than 25 percent of women have their second and their
third child within 24 months of their previous child.
Furthermore, even with the larger increases in birth intervals with higher parity, the 
25th percentile birth intervals for the fourth spell is around 24 months for women with 
less than eight years of education.

% [median birth intervals]
The median birth intervals have increased more than the 25th percentile birth intervals
but are still also relatively short.
Although most median birth intervals have increased by three to six months over the four 
decades, the majority of median birth intervals are still at 36 months or below, 
with the shortest only 29 months.
For comparison, fast changing countries have median birth intervals that increase around 
3.5 months per decade with declining fertility \citep{Rutstein2011,Casterline2016}.
Furthermore, the NFHS reports show median closed birth intervals that have barely
moved over time, staying at approximately 31 months, underscoring the importance of 
accounting for censoring of birth spells when trying to understand how birth spacing 
changes over time.

An important exception to the relatively short median intervals come from the better 
educated women, especially in the absence of sons.
The longest median interval was 48 months for the third spell for urban women with twelve 
or more years of education who had no boys, an increase of almost 15 months over
four decades.
Similary, urban women with eight to eleven years of education and no sons had median
birth intervals of around 40 months, an increase of about 10 months.
Finally, the best educated women now have a median second birth interval of around 42
months whether their first child was a boy or a girl.

% [75th percentile birth intervals]
The most substantial differences across women come from the longer birth intervals, 
captured by the 75th percentile birth intervals.
For the two lowest education groups the increases in the 75th percentile birth 
intervals are relatively small for the second spell and slightly more substantial for the 
third spell at around five months.
However, the fourth spell show significant increases, especially for women with two
or three sons, with increases of fifteen to twenty months.
The fourth spell is also where most of the divergence across sex compositions happen,
but even for women with no or one son, the 75th percentile birth intervals have
increased by a half to a full year.
For women with eight to eleven years of education, the 75th percentile spacing length for 
women with two girls in the third spell increased by more than 15 months, mainly as the 
result of more widespread use of sex selection as the sex ratios above indicate.

% [Exceptions to the results]
The best-educated women are an especially interesting case that run somewhat counter
to the results above.
First, despite a rapidly declining probability of having a third child when they already 
have a son, there is little change in the associated birth intervals and all are lower for 
the third spell than for the second.
Second, the massive increase in the sex ratio to almost 80 percent boys for women with two 
girls is associated with a more than 20 month increase in the 75th percentile birth 
interval.
The third spell is also the case with the clearest reversal in the spacing pattern by sex 
composition; the birth intervals with two girls are consistently longer than with one
or two sons no matter the percentile used.

What do these results tell us about the role of son preference and the extent to which 
the introduction of sex selection changed birth spacing across sex compositions?

The clearest illustration of how the combination of son preference and the introduction of 
sex selection impacted birth spacing comes from the third spell of the best-educated urban 
women.
Women with two girls have a predicted sex ratio for their third child approaching 80 
percent boys and a 75th percentile birth interval of almost 70 months, about 13 months
longer than if they had at least one son.
Even more striking is the change for this group when sex selection was first introduced.
The 75th percentile birth interval with two girls increased from 48 months to 
64 months with two girls, while the other sex compositions showed a 
slightly decrease from around 55 months to 54 months.
Theses changes may even be an underestimate of the changes in birth spacing sine
this particular group appear to have had access to sex selection even before it became
widespread as shown by the unequal sex ratio for the 1972--1984 period for women with
two girls.

Even for the women the least likely to use sex selection, such as rural women with no 
education, son preference shows up clearly. 
For this group, the shortest spacing---across all spells---is when they have only girls. 
Notably, the difference in birth intervals across sex compositions has grown over 
time as spacing when sons are present increased.
Despite the strong son preference in spacing and sex ratios, there is also some indication 
that parents prefer a mix of boys and girls;
women with two boys \emph{and} a girl consistently have the lowest parity progression 
probability in the fourth spell across education levels.

Did the predictions of a turn-around in the use of sex selection come true?
There is no clear evidence for or against a reversal in the use of sex selection, with 
some showing increases in sex ratios between the last two periods, others little change, 
and some a decline.
The best-educated women is again a good illustration.
The percent boys born to women with two girls continued to increase over the last two
periods, but the likelihood of a third birth declined and when the first child was a girl, 
the sex ratio for the second birth dropped slightly as did the probability of a second birth. 
There is, however, also cases where there is clearly no abatement in the increasing use of 
sex selection. 
The sex ratios in the absence of girls continued to increase for rural women with
one to seven years of education and the likelihood of an additional birth remained high.

% [Interpretation]
How do these changes in birth intervals tie in with the changes in India over the four
decades?
% [TK economic growth? -> higher return to education for males -> lower fertility and
% higher investment in boys' education -- has anybody looked at this?]
First, the parity progression probabilities have been falling over time across all 
education levels, especially for women with at least one son.
Falling fertility is associated with increases in longer spacing, although it is still an 
unresolved question why \citep{Casterline2016}.
What is especially interesting is that the results for the most educated women show that 
the rapidly declining probability of a third birth for women with one or two sons did not 
lengthen the birth intervals at any of the percentiles.
One possibility is that the most educated women have both better access to modern
contraceptive and are better at using traditional contraceptive methods 
\citep{Rosenzweig1989}.

Second, there is evidence that in situations where sex selection does not play a role 
in birth intervals, women with the smallest change in birth intervals are also those most 
likely to work.
For example, the increases in average third spell birth intervals for urban women with
one boy and one girl over the four decades were 2.7, 3.4, 5.8, and 1.8 months from lowest 
to highest education groups.%
\footnote{
For women with two sons, the numbers were 4.3, 6.3, 7.0, and 3.0, respectively.
See the online appendix tables for the average birth intervals.
}
This means that the group with the lowest labor force participation, women with eight to 
eleven years of education, was also the group with the largest increases in average 
spacing, likely driven by the substantial improvement in household income that came
from economic growth.

Finally, although birth spacing has increased over time with improving economic 
conditions and falling fertility, most of the increases in birth spacing come from the
increased use of sex selection. 
The most substantial increase over the four decades is a full year for the best-educated 
women with two girls, which is the group with the most biased sex ratio.

\section{What Happened to Fertility?\label{sec:fertility}}

The tempo effect from longer birth intervals means that the total fertility rate may 
underestimate cohort fertility. 
The next question I address is, therefore, to what extent did the changes bias the 
fertility estimates for India? 
To this end, I compare fertility based on a variation of the total fertility rate with 
predicted cohort fertility from the hazard model results.

\input{../tables/fertility.tex}

Table \ref{tab:fertility} shows the two fertility measures by urban/rural and education.
Because the hazard model predictions only use births up to parity four, I calculate
fertility rates using only the same births.
Hence, my fertility rates are not directly comparable to those in the NFHS reports,
but otherwise, I follow the procedure used in the NFHS reports \citep{Croft2018}.
I calculate age-specific fertility rates for five-year age groups using the number of 
births in months 1 to 36 before the survey month and the number of women in each age 
group.

%
% \footnote{
% See \citet{Croft2018} for more detail.
% Because NFHS-1 and NFHS-2 did not interview all women about fertility, the number of women 
% is based on the household rosters, assuming that never-married women have had no children.
% Because of the low number of births to women aged 45 to 49 this age group was not
% used for the hazard model estimations and, therefore, is also excluded for the fertility
% rate.
% The marital total fertility rate lead to unrealistically high numbers and is, therefore,
% not presented \citep{Hoem2011,Laplante2015}.
% }

Because NFHS-1 was after the introduction of sex selection, it is not possible 
to calculate a fertility rate in precisely the same manner for a period 
before sex selection was widely available.
Instead, I calculate the fertility rates for women who were between 15 and 39 years of
age five years before the survey month, again using the number of births three years 
before that, shown as ``1987--88''.
Given the relatively low number of births to women 40 to 45 years of age, this approach 
provides the best estimate of the fertility rate at the time when sex selection still was 
not widespread.

To predict cohort fertility based on the hazard models above, I estimate the probability 
of progression to the next parity for each spell.
Since the parity progression depends on the sex composition for prior children for parity 
two and above, I estimate the progression probability for each possible sex composition
and weigh those probabilities with the likelihood that they will occur based on the
prior spells.

Sterilizations are not incorporated into the hazard model because most occur
immediately after giving birth. 
To compensate, I estimate the likelihood of sterilization using a Logit model, predict 
the probability of not getting sterilized, and then use that to scale down the parity
progression probability when predicting cohort fertility.

I include the spell from marriage to first birth, despite the problems
capturing the exact timing of marriages since the estimated progression
probabilities should not be affected by this problem.
To ensure that the predictions depend solely on sample composition and the estimate, I
use the age at marriage for each woman and predict the likelihood of progressing to
each parity, assuming three years increases in age between each parity.
Shorter assumed increases in age between births leads to slightly higher predicted 
fertility.

The survey rounds do not coincide directly with the periods used for the hazard model.
I, therefore, use the model results for 1972--1984, 1985--1994, 1995--2004,
and 2005--2016 for rounds 1 through 4 of the NFHS, respectively.

% Because of the small number of women who had the highest level of education in the
% 1972--1984 period, I do not predict fertility for this case.

The predicted cohort fertility based on the hazard model is higher than the four-parity 
fertility rate in almost all cases.
Only for women with no education in the first period is there almost no difference 
between the two fertility measures; a situation where fertility is high, spacing
very short and likely unchanged for an extended period before.

Consistent with a more substantial bias in the fertility rate when the age of marriage 
and the length of birth intervals are increasing, the absolute bias is least in the first and the last 
period and highest in the middle two periods.
Hence, the fertility rate declined too fast from the mid-1980s to the century's end.
Only recently, as the rate of increase for the birth intervals have
slowed, are the predicted cohort fertility and the fertility beginning
to converge again.
Even with the convergence, however, the predicted cohort fertility for 2005--16 is 
above the 1992--93 fertility rate for every group except urban women with no education.
Furthermore, for the last period, the predicted cohort fertility is still between ten
and 20 percent higher than the fertility rate.

Another indication of the fertility rate bias is that it is increasing over time for some 
groups.
For example, urban women with eight to eleven years of education showed 1.84, 1.81, and
1.87 over the last three surveys.
These increases likely arise from the stabilization of the age of first birth and the 
spacing between births.

%
% \footnote{
% Increases in TFR over time, even in the face of declining cohort fertility, are not 
% unusual, as shown by Columbia in the early 1990s \citep{Bongaarts1999}.
% }

Despite the declines in the predicted cohort fertility, it is still mostly above the 
generally accepted replacement level of 2.1.
Only for urban women with twelve or more years of education is the predicted cohort
fertility clearly below 2.1, and even then, it is still more than 0.3 children higher than 
the fertility rate estimate of 1.5.
Finally, the predicted cohort fertility numbers are likely too low,
especially for the lower education groups, since it relies only on the
first four births and because the model ignores births that took place past the end
of the imposed 105 month birth interval censoring.

\section{Mortality and the Changing Birth Spacing\label{sec:mortality}}

The final question I address is whether there is an association between infant mortality 
and increases in birth spacing and sex selection. 
Starting with the sample used for estimating birth spacing, I select children born more 
than 12 months before the survey month. 
I restrict the analyses to parities two and three because of the small number of births 
and deaths for parity four. 
Furthermore, I do not show the results for women with twelve or more years of education 
for the 1972--84 period because of the small number of women.

The dependent variable is whether the child died within the first twelve
months of life.
The main set of explanatory variables consists of dummies for the spacing from
the prior birth.
Starting nine months after the prior birth, the birth interval dummies cover 12 
month periods until 57 months, which covers until 105 months after the prior birth.
I use dummies for sex of the index child and the sex composition 
of the prior children.
The birth spacing dummies, the sex of index child, and the sex composition dummies
are all interacted.
Since the actual number of abortions is unobserved, the interactions between
the sex composition of prior children and the sex of the index child serve
as proxies for the use of sex selection.
The other explanatory variables are the same as above, and estimations
are done separately by education level and parity.

I estimate the probability of infant mortality using a Logit model.
Figures \ref{fig:mortality_low_med} and \ref{fig:mortality_high_highest} present 
the predicted probability of the second child dying within the first year at the 
possible combinations of index child sex, sex composition of prior children, and 
the birth spacing dummied.%
\footnote{
The online Appendix shows the corresponding graphs for the third child.
}
All other variables take their average values.
The graphs do not show confidence intervals to ease their legibility.

An important caveat is that the estimations do not 
address potential selection problems.
For example, if women who have difficulties conceiving or carrying a pregnancy 
to term also have a higher mortality risk for their offspring, a spurious correlation 
between long birth spacing and mortality may arise \citep{Kozuki2013}. 
Unfortunately, methods used previously to address selection, such
as family fixed effects, do not work well when the number of births is
low as it is for better-educated women \citep{Kozuki2013,Molitoris2019}.
However, the fixed effects and linear probability results did not 
deviate substantially in prior research.

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{1972-1984}}
% \setcounter{subfigure}{-1}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \captionsetup[subfigure]{font=footnotesize,labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[No Eduction]{\includegraphics[width=\textwidth]{mortality_s2_p1_low_dummies}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{-0} 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \captionsetup[subfigure]{font=footnotesize,labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[One to Seven Years of Education]{\includegraphics[width=\textwidth]{mortality_s2_p1_med_dummies}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \\
% \rotatebox[origin=c]{90}{\small{1985-1994}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p2_low_dummies}
%     \end{minipage}
% } 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p2_med_dummies} 
%     \end{minipage}
% } 
% \\
% \rotatebox[origin=c]{90}{\small{1995-2004}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p3_low_dummies}
%     \end{minipage}
% } 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p3_med_dummies} 
%     \end{minipage}
% } 
% \\
% \rotatebox[origin=c]{90}{\small{2005-2016}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p4_low_dummies}
%     \end{minipage}
% } 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p4_med_dummies} 
%     \end{minipage}
% } 
% \caption{Predicted Probability of Second Child's Infant Mortality 
% for Women with No Education
% and Women with One to Seven Years of Education}
% \label{fig:mortality_low_med}
% \end{figure}

% \begin{figure}[htpb]
% \centering
% \rotatebox[origin=c]{90}{\small{1972-1984}}
% \setcounter{subfigure}{-1}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \captionsetup[subfigure]{font=footnotesize,labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Eight to Eleven Years of Education]{\includegraphics[width=\textwidth]{mortality_s2_p1_high_dummies}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \begin{minipage}[t]{0.39\textwidth}\vspace{-2.68cm}\centering\footnotesize{Twelve or More Years of Education}\end{minipage}
% \\
% \rotatebox[origin=c]{90}{\small{1985-1994}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p2_high_dummies}
%     \end{minipage}
% } 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p2_highest_dummies}
%     \end{minipage}
% } 
% \\
% \rotatebox[origin=c]{90}{\small{1995-2004}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p3_high_dummies}
%     \end{minipage}
% } 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p3_highest_dummies} 
%     \end{minipage}
% } 
% \\
% \rotatebox[origin=c]{90}{\small{2005-2016}}
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p4_high_dummies}
%     \end{minipage}
% } 
% \subfloat[Second Spell]{
%     \begin{minipage}{0.37\textwidth}
%         \includegraphics[width=\textwidth]{mortality_s2_p4_highest_dummies} 
%     \end{minipage}
% } 
% \caption{Predicted Probability of Second Child's Infant Mortality 
% for Women with Eight to Eleven of Education
% and Women with Twelve or More Years of Education}
% \label{fig:mortality_high_highest}
% \end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{mortality_low_med}
\caption{Mortality low medium}
\label{fig:mortality_low_med}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{mortality_high_highest}
\caption{Mortality high highest}
\label{fig:mortality_high_highest}
\end{figure}

There has been substantial convergence in mortality risk across groups over time.
For intervals 21 months or longer, there is now little difference across 
the education groups, with even the no education group showing an infant mortality risk 
below five percent.

Very short birth intervals still exhibit a higher mortality risk, although the 
effect declines with education level.
For the best-educated women, the mortality risk is three to four percent,
whereas women with no education still show a risk that is close to ten percent.

Despite the prior findings of differential mortality by sex, there is little evidence 
that girls have substantially higher mortality risk.
There is some weak evidence that a boy born after a girl has a lower mortality
risk in the earliest periods.
This difference disappears, however, with the general decline in mortality risk.

If sex selective abortions lead to higher mortality risk, boys born after a girl 
(the solid lines)
should have an increased risk with longer spacing for the two highest education groups 
in the last two periods.
There are, however, no apparent consistent differences between these groups and the other 
potential combinations.

% line 3661 in an_infant_mortality.log
The raw numbers for women with the most uneven sex ratio also suggests that even with 
high use of sex selection there is no impact on child mortality. 
A total of 1,004 women with twelve or more years of
education and no boys at the start of the third spell in the last
period, had a third child, of which 685 were boys. 
Of these 685 boys, only six died within the first year of life.
Half of those who died were born in the 9-32 months interval, and none in 57 months 
or longer interval.

\section{Conclusion\label{sec:conclusion}}

% Background
Over the last four decades India saw a dramatic increase in the male-to-female
sex ratio at birth as access to sex selection spread and son preference remained. 
At the same time, economic growth was strong, schooling increased, and the total
fertility rate fell to close to replacement.

% Questions
The main question I address in this paper is how birth spacing responded to the major changes
that occurred in India over the last four decade, in particular the spread of sex selection.
In addition, I examine two related questions.
First, did the changes in birth spacing bias the standard fertility estimates for India? 
Second, what is the relationship between infant mortality and the changes in birth spacing 
and sex selection?

% What has happened to spacing and why
Birth intervals has lengthen, often substantially, over the last four decades.
The increases in spacing are larger, the higher the parity and the higher the mother's
education, and is larger in urban than in rural areas.

Groups with the highest use of sex selection have the most substantial increases in 
birth intervals, with increases in average birth intervals of up to 12 months. 
Some of these increases are so large that we even observe a reversal of the traditional 
spacing pattern: instead of the absence of sons leading to the shortest birth intervals, 
son preference and sex selection lead some groups to have longest birth intervals when
they have no sons.

Despite the longer average birth intervals, a large number of women still have very 
short birth intervals. 
Hence, the changes in birth spacing are driven predominately by the longer intervals 
becoming even longer over time.
This pattern holds generally, but is especially prevalent for those who use sex selection 
and, therefore, may end up with multiple abortions.

% fertility
Fertility has declined for all groups, but the likelihood of having an additional child
still depends strongly on the number of sons, with women with no sons having the highest 
parity progression probabilities.

The combination of the secular increases in spacing and the fact those women most likely to use
sex selection are also those most likely to have a next birth, makes the total fertility rate a 
more biased measure of cohort fertility. 
This bias was most prominent early in the spread of sex selection, when the fertility rate 
was up to one child lower than the predicted cohort fertility. 
It is, however, still present, with the predicted cohort fertility ten to twenty percent 
higher than the fertility rate.

Tempo effects are studied extensively in the literature \citep[See, for example, ][]{Bongaarts1999}.
Still, there are, to my knowledge, no other cases where there has been as 
substantial an increase in birth intervals and associated bias in fertility rates 
as for India.
It is conceivable that we might see increases in the total fertility rate as birth spacing 
stabilizes or, if interventions against sex selection are successful, even shortens again.

The best-educated urban women are the only group for whom the predicted cohort fertility 
is below replacement at 1.8. 
A caveat to the predicted cohort fertility is that only Hindu women are in the sample and 
that the model only includes births up to parity four.
However, almost everybody marries, and Hindu constitute 80 percent of the population.
Furthermore, with only the first fourth births used, these estimates may
be lower-bound, especially for the less educated women.

% Mortality
There has been a substantial reduction in infant mortality over time, and the size of the 
reductions is inversely related to the mother's education.
Hence, there is now little difference in mortality risk across education groups if the 
birth took place more than 21 months from the prior birth.
Short birth spacing is still associated with higher mortality, although the difference is 
small for the best-educated women.
There is no evidence that children born to women who are the most likely to use sex 
selection have higher mortality risk.

An important implication of the results here is that population growth in India
may be higher than expected. 
There has been a rapid reduction in infant mortality risk, the longer spacing should 
further reduce mortality risk, and cohort fertility is higher than previously realized.

% Future research and policy

There are two critical questions that future research should address.
First, sex selection means that girls-only families are less likely to have
very short birth intervals, which may reduce sibling competition. 
Hence, better health outcomes for girls with sex selection could be an
unintended side-effect, rather than a result of girls becoming more
valued as is often assumed. 
Comparing prior children's outcomes across sex composition and the sex of the next child 
could be a way to understand why girls' health outcomes improve in the presence of sex selection.

Second, is there a relationship between female labor force participation
and sex selection? 
Women may be staying out of the labor market precisely because sex selection makes 
them more likely to have a boy and increases the expected birth spacing. 
Better job opportunities for women would affect sex selection for two reasons. 
First, it makes it more expensive to be out of the labor market for long periods. 
Second, it would moderate the differential in potential earnings between husband
and wife and make it more attractive to invest in female compared to
male offspring's human capital. 
This approach could, however, be a double-edged sword. 
If better job opportunities further lower desired fertility, the use of sex selection 
may increase, everything else equal.
Hence, understanding the trade-off between long-term benefits from
improvements in women's labor force participation and short-term costs
from potential increases in sex selection is of paramount importance.

% The higher return to male offspring's human capital combined with access to sex
% selection may also explain another puzzle, why is the labor force participation so
% low when fertility is low and education has increased.
% As shown here, part of the answer may be that the number of children is, in fact, not as 
% low as previously thought, both because of the increased spacing and that more 
% children survive.

\clearpage

\onehalfspacing
\bibliographystyle{aer}
\bibliography{sex_selection_spacing}

\addcontentsline{toc}{section}{References}

\clearpage
\newpage

\appendix

% CHANGING NUMBERING OF FIGURES AND TABLES FOR APPENDIX
\renewcommand\thefigure{\thesection.\arabic{figure}}    
\renewcommand\thetable{\thesection.\arabic{table}}

\section*{Appendices for Online Publication}

These appendices are intended for online publication.
They provide the descriptive statistics, additional
estimated duration tables, and graphs for all 
education groups and spells.

\clearpage
\newpage

\section{Characteristics of Women's Work Experiences}

\setcounter{figure}{0}
\setcounter{table}{0}

Figure \ref{fig:work_by_survey} shows the percent of married women who are 
currently working at the time of the survey by age group and education level.
No other labor force participation question is consistently available across all 
four surveys. 
Because the question refers to currently working, the percentages are lower in previous
studies.

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{currently_working_urban_20} 
    \end{minipage}
}
\caption{Percentage of married women who were working at the time of the 
survey by age group and area of residence}
\label{fig:work_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_cash_urban_20} 
    \end{minipage}
}
\caption{Percentage of women paid cash or cash and in-kind of those women who were 
working at the time of the survey by age group and area of residence}
\label{fig:work_cash_by_survey}
\end{figure}

\begin{figure}[!htpb]
\centering
\rotatebox[origin=c]{90}{\small{40 Years or Older}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_40}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_40} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{30 to 39 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_30}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_30} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{20 to 29 Years Old}}
\subfloat[Rural]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_rural_20}
    \end{minipage}
} 
\subfloat[Urban]{
    \begin{minipage}{0.45\textwidth}
        \includegraphics[width=\textwidth]{work_family_urban_20} 
    \end{minipage}
}
\caption{Percentage women who worked for a family member of those 
working at the time of the survey, by age group and area of residence}
\label{fig:work_family_by_survey}
\end{figure}

\clearpage
\newpage

\section{Recall Error and the Sex Ratio}

\setcounter{figure}{0}
\setcounter{table}{0}

The reliability of the results depends on the correctness of the birth histories
provided by the respondents.
A significant concern here is underreporting of child mortality, especially a systematic
recall error where respondents' likelihood of reporting a deceased child depends on the 
sex of that child. 
This Appendix section assesses the degree of recall error across the surveys and discusses 
methods to address it.

NFHS enumerators probe for any missed births, although the method depends on the survey.
NFHS-1 probe for each calendar birth interval that is four or more years.
NFHS-2 asked for stillbirths, spontaneous and induced abortions and also probed 
for each calendar birth interval four or more years.
NFHS-3 and NFHS-4 did not directly use birth intervals, but asked whether there were any 
other live births between (name of previous birth) and (name), including any children who 
died after birth, and asked for births before the birth listed as first birth and
after the last birth listed as the last birth.

Probing catches many initially missed births, but systematic recall error based on son
preference may still be a problem.
First, son preference leads to significantly higher mortality for girls than boys.
Secondly, son preference makes it more likely that parents will remember deceased boys 
than deceased girls.
Finally, in the absence of sex-selective abortions, parents with a preference for sons may
have the next birth sooner if the last child was a girl than if it was a boy.
If this girl subsequently dies, she is more likely to be missed if probing for missed 
births is only done for long intervals as in NFHS-1 and NFHS-2.

I use two approaches to examine the degree of recall error.
The first approach is to test whether the observed sex ratio is significantly different
from the natural sex ratio.
The natural sex ratio is approximately 105 boys to 100 girls or
51.2\% \citep{ben-porath76b,jacobsen99,Portner2015b}.
Prenatal sex determination techniques did not become widely available until the mid-1980s, 
so any significant deviation from the natural sex ratio before that time is likely the 
result of recall error.
The second approach is to compare births that took place during the same period but
where captured in different surveys.
Recall error is likely to increase with time, so births and deaths that took place earlier 
are more likely to be subject to recall error than more recent events.

Table \ref{tab:recallBirthBO1} shows the sex ratios of children recorded as first-born by 
year of birth, together with tests for whether the observed sex ratio is significantly 
higher than the natural sex ratio and whether more recent surveys have a higher sex ratio 
for the cohort than earlier surveys for the same period births.
Births are combined into five-year cohorts to achieve sufficient power.

\input{../tables/recallBirthBO1.tex}

% \input{../tables/recallBirthBO2.tex}

The ``first-born'' sex ratios illustrate the systematic recall error problem well.
In all four surveys around 55 percent of children reported as first-born are boys
for the first cohort of births observed.
Given that these cohorts cover from 1960-1964 to 1980-1984, which is before sex selection 
techniques became available in India, the most likely explanation for the skewed sex ratio 
is that some children listed as first-borns were not, in fact, the first children born in 
their families.
Instead, for a substantial proportion of families, their first-born was a girl who died 
and went unreported when enumerators asked about birth history.

As expected, the difference between the observed sex ratio and the natural sex ratio is 
less pronounced the closer to the survey date the cohort is.
The observed sex ratio for children born just before the NFHS-1 survey and listed as 
first-born is 0.517, which is not statistically significantly different from the
natural sex ratio.
The same general pattern holds for the other three surveys, with cohorts further away
from the survey date more likely to have a sex ratio skewed male.

% Second births show a pattern very similar to that for first births.
% An interesting difference is that there is evidence of a U-shaped relationship between
% time and sex ratio for second births.
% Cohorts furthest away from the survey year show the highest sex ratios, but declines to the
% natural sex ratio in the mid-1980s, and the sex ratio is then significantly higher again 
% for more recent births.
% This is in line with the results in the main paper that show that sex selective abortions 
% take place on lower parity births as desired fertility declined.

Finally, across surveys, the same cohort tends to show a higher sex ratio the more recent 
the survey (births in the cohort took place earlier relative to the survey date).
Despite this, few cohorts show significantly different sex ratios across surveys, most 
likely because of a lack of power.
The exception is that comparisons involving NFHS-4 are mostly statistically significant
since the number of surveyed households in NFHS-4 were much larger than in prior surveys.

The problem with the above approach is that the year of birth is affected by recall error; 
a second born child listed as first-born is born later than the real first born child.
Year of marriage should, however, be affected neither by parental recall error 
nor the use of sex-selective abortions.
Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2}, therefore, shows sex 
ratios of children recorded as first-born and second-born by year of parents' marriage, 
together with tests for whether the observed sex ratio is significantly higher than the 
natural sex ratio and whether more recent surveys show a higher sex ratio for the cohort 
than earlier surveys.
The basic recall error pattern remains, with women married longer ago more
likely to report that their first-born is a boy.
Similarly, comparing women married in the same five-year period across surveys shows
that women married longer ago are more likely to report having a son.

\input{../tables/recallMarriageBO1.tex}

\input{../tables/recallMarriageBO2.tex}

The relationship between the length of marriage and recall error can also be seen in 
Figures \ref{fig:sex_ratio_recall_rounds_bo1} and \ref{fig:sex_ratio_recall_rounds_bo2}, 
which show the observed sex ratio for children reported as first born as a function of 
the duration of marriage at the time of the survey.
The solid line is the sex ratio of children reported as first-born by the number of years 
between the survey and marriage, while the dashed lines indicate the 95 percent confidence 
interval and the horizontal line the natural sex ratio (approximately 0.512).
To ensure sufficient cell sizes I group years into twos.
In line with the results from Tables \ref{tab:recallMarriageBO1} and 
\ref{tab:recallMarriageBO2}, the observed ratio of boys is increasingly above the expected 
value the longer ago the parents were married.

\begin{figure}
\centering
\subfloat[NFHS-1]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_1}}
\subfloat[NFHS-2]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_2}} \\
\subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_3}} 
\subfloat[NFHS-4]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_4}} 
\caption{Ratio of Boys for ``First'' Births by Survey Round}
\label{fig:sex_ratio_recall_rounds_bo1}
\end{figure}

\begin{figure}
\centering
\subfloat[NFHS-1]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_1_bo2}}
\subfloat[NFHS-2]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_2_bo2}} \\
\subfloat[NFHS-3]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_3_bo2}} 
\subfloat[NFHS-4]{\includegraphics[width=.49\textwidth]{recall_sex_ratio_marriage_round_4_bo2}} 
\caption{Ratio of Boys for ``Second'' Births by Survey Round}
\label{fig:sex_ratio_recall_rounds_bo2}
\end{figure}

The increasingly unequal sex ratio with increasing marriage duration suggests that
a solution to the recall error problem is to drop observations for 
women who were married ``too far'' from the survey year.
The main problem is establishing what the best cut-off point should be, with the
trade-off between retaining enough observations and the correctness of the information.
As Tables \ref{tab:recallMarriageBO1} and \ref{tab:recallMarriageBO2} show, there are 
differences in recall error across the three surveys and between the two birth
orders, although this may be the result of differences in the number of observations 
across surveys.
Furthermore, the recall error pattern is not entirely consistent across observed birth 
orders.
Since most of the surveys start showing significantly biased sex ratio from around 22
years of marriage on, I drop all observations where the marriage took place 22 years
or more.

\clearpage
\newpage

\section{Descriptive Statistics}
\setcounter{figure}{0}
\setcounter{table}{0}

% Descriptive statistics tables
\input{../tables/des_stat.tex}

\clearpage
\newpage

\section{Additional Results Figures and  Tables}

\setcounter{figure}{0}
\setcounter{table}{0}

Figures \ref{fig:spacing_low_urban} and \ref{fig:spacing_highest_rural} show 25th, 50th,
and 75th percentile birth intervals, the sex ratio, and the probability of parity
progression by spell for urban women with no education and rural women with twelve or
more years of education, respectively.

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_low_urban}
\caption{combined graphs - low urban}
\label{fig:spacing_low_urban}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=\textwidth,height=\textheight,keepaspectration=true]{bs_highest_rural}
\caption{combined graphs - highest rural}
\label{fig:spacing_highest_rural}
\end{figure}

The first set of tables, Tables \ref{tab:p25_p50_p75_low}, \ref{tab:p25_p50_p75_med},
\ref{tab:p25_p50_p75_high}, and \ref{tab:p25_p50_p75_highest} show 25th, 50th, and
75th percentile birth intervals together with their standard errors.
The standard errors for all measures are based on bootstrapping,
where the model is repeatedly estimated using resampling with replacement.

The second set of tables, 
Tables \ref{tab:avg_sex_ratio_low}, \ref{tab:avg_sex_ratio_med}, 
\ref{tab:avg_sex_ratio_high}, and \ref{tab:avg_sex_ratio_highest}, show predicted average 
birth intervals, sex ratios, and probabilities of having a birth by decade, spell, and sex 
composition for the four education levels separated by the area of residence, together 
with bootstrapped standard errors for all three outcomes.
To find the average birth interval, I calculate, for each woman, the probability of 
giving birth in each t, and her expected spell length from these probabilities. 
I then average the individual expected spell lengths across women using their parity 
progression probabilities as weights. Finally, I add nine months because spells begin 
nine months after the previous birth.

I also show whether durations for sex composition other than only girls are statistically 
significantly different from the duration with only girls based on bootstrapped 
differences. 
The cleanest test is comparing durations after only boys with durations after
only girls, but the number of births to women with only sons becomes small 
in the later periods.
Hence, it is possible to have substantial differences in spacing that are
not statistically significant because of low power, especially for the third 
and fourth spell.

Each predicted percent of boys is tested against the natural percentage of
boys using the bootstrapped standard errors.
The natural sex ratio is approximately 105 boys to 100 girls or
51.2\% \citep{ben-porath76b,jacobsen99,Portner2015b}.
The predicted percentage boys may differ from the natural rate because of 
natural variation, any remaining recall error not corrected for, or 
sex selection.

\input{../tables/bootstrap_duration_p25_p75_low_all.tex}

\input{../tables/bootstrap_duration_p25_p75_med_all.tex}

\input{../tables/bootstrap_duration_p25_p75_high_all.tex}

\input{../tables/bootstrap_duration_p25_p75_highest_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_low_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_med_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_high_all.tex}

\input{../tables/bootstrap_duration_avg_sex_ratio_highest_all.tex}

\clearpage

% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_low_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_low_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_low_urban}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_med_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_med_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_med_urban}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_high_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_high_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_high_urban}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_highest_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_highest_urban}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 25th Percentile Birth Intervals for Urban Women by Spell and Education
% }
% \label{fig:p25_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_low_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_low_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_low_rural}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_med_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_med_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_med_rural}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_high_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_high_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p25_spell4_high_rural}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p25_spell2_highest_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p25_spell3_highest_rural}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 25th Percentile Birth Intervals for Rural Women by Spell and Education
% }
% \label{fig:p25_rural}
% \end{figure}
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_low_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_low_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_low_urban}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_med_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_med_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_med_urban}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_high_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_high_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_high_urban}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_highest_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_highest_urban}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 50th Percentile Birth Intervals for Urban Women by Spell and Education
% }
% \label{fig:p50_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_low_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_low_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_low_rural}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_med_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_med_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_med_rural}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_high_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_high_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{p50_spell4_high_rural}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{p50_spell2_highest_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{p50_spell3_highest_rural}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in 50th Percentile Birth Intervals for Rural Women by Spell and Education
% }
% \label{fig:p50_rural}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_low_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_low_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_low_urban}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_med_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_med_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_med_urban}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_high_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_high_urban}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_high_urban}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_highest_urban}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_highest_urban}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in Average Birth Intervals for Urban Women by Spell and Education
% }
% \label{fig:avg_urban}
% \end{figure}
% 
% 
% \begin{figure}[htpb]
% \captionsetup[subfigure]{labelformat=empty,position=top}
% \captionsetup{font=footnotesize,skip=2pt}
% \centering
% \caption*{No education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_low_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_low_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_low_rural}}
% \\
% \caption*{One to Seven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_med_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_med_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_med_rural}}
% \\
% \caption*{Eight to Eleven Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_high_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_high_rural}}
% \subfloat[][Fourth Spell]{\includegraphics[width=0.32\textwidth]{avg_spell4_high_rural}}
% \\
% \caption*{Twelve or More Years of Education}
% \subfloat[][Second Spell]{\includegraphics[width=0.32\textwidth]{avg_spell2_highest_rural}}
% \subfloat[][Third Spell]{\includegraphics[width=0.32\textwidth]{avg_spell3_highest_rural}}
% \begin{minipage}{0.32\textwidth}\hspace{1cm}\end{minipage}
% \captionsetup{font=normalsize}
% \caption{
% Changes in Average Birth Intervals for Rural Women by Spell and Education
% }
% \label{fig:avg_rural}
% \end{figure}

\clearpage
\newpage

% \section{Survival Curves Conditional on Parity Progression for All Education and Spell Groups}
% 
% \setcounter{figure}{0}
% \setcounter{table}{0}
% 
% 
% % [FIRST SPELL]
% 
% % \subsection{First Spell}
% % 
% % \input{../figures/appendix_spell1_low.tex}
% % 
% % \input{../figures/appendix_spell1_med.tex}
% % 
% % \input{../figures/appendix_spell1_high.tex}
% % 
% % \begin{figure}[htpb]
% % \centering
% % \caption*{No Education}
% % \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_low_urban_pps}} 
% % \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_low_rural_pps}} \\
% % \caption*{1-7 Years of Education}
% % \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_med_urban_pps}} 
% % \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_med_rural_pps}} \\
% % \caption*{8 or more Years of Education}
% % \subfloat[Urban]{\includegraphics[width=0.49\textwidth]{spell1_highest_urban_pps}} 
% % \subfloat[Rural]{\includegraphics[width=0.49\textwidth]{spell1_highest_rural_pps}} 
% % \caption{Survival curves conditional on progression to first birth; start point is month of marriage}
% % \label{fig:results_spell1_pps}
% % \end{figure}
% % 
% % 
% % \clearpage
% % \newpage
% % 
% \subsection{Second Spell}
% 
% 
% % \input{../figures/appendix_spell2_low.tex}
% 
% 
% % PARITY PROGRESSION SURVIVAL CURVES - Low
% 
% % Low education
% \vspace*{-0.75cm}
% \begin{figure}[hp!]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2005]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_low_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with no education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_low_pps}
% \end{figure}
% 
% 
% 
% 
% % SPELL 2 - URBAN - MEDIUM
% 
% % \input{../figures/appendix_spell2_med.tex}
% 
% % PARITY PROGRESSION SURVIVAL - MEDIUM
% 
% 
% % Low education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_med_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.32\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with 1-7 years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_med_pps}
% \end{figure}
% 
% 
% 
% % High education
% 
% % \input{../figures/appendix_spell2_high.tex}
% 
% % PARITY PROGRESSION SURVIVAL - High education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_high_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with eight to eleven years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_high_pps}
% \end{figure}
% 
% 
% % Highest education
% 
% % PARITY PROGRESSION SURVIVAL - Highest education
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g1_highest_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g2_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g3_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell2_g4_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with twelve or more years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell2_highest_pps}
% \end{figure}
% 
% 
% \clearpage
% \newpage
% 
% \subsection{Third Spell}
% 
% % low education
% % \input{../figures/appendix_spell3_low.tex}
% 
% % PARITY PROGRESSION SURVIVAL - LOW
% \vspace*{-0.75cm}
% \begin{figure}[hp!]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_low_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with no education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell3_low_pps}
% \end{figure}
% 
% 
% 
% 
% % Medium education
% 
% % \input{../figures/appendix_spell3_med.tex}
% 
% % PARITY PROGRESSION SURVIVAL - MEDIUM
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_med_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with 1 to 7 years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell3_med_pps}
% \end{figure}
% 
% 
% 
% 
% % High education
% 
% % \input{../figures/appendix_spell3_high.tex}
% 
% % PARITY PROGRESSION SURVIVAL - high
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_high_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with eight to eleven years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell3_high_pps}
% \end{figure}
% 
% 
% % Highest education
% 
% % PARITY PROGRESSION SURVIVAL - highest
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_highest_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_highest_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g1_highest_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g2_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g3_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell3_g4_highest_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with twelve or more years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell3_highest_pps}
% \end{figure}
% 
% 
% 
% 
% 
% \clearpage
% \newpage
% 
% \subsection{Fourth Spell}
% 
% % low education
% 
% % \input{../figures/appendix_spell4_low.tex}
% 
% % PARITY PROGRESSION SURVIVAL - LOW
% \vspace*{-0.75cm}
% \begin{figure}[hp!]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_low_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_low_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_low_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with no education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell4_low_pps}
% \end{figure}
% 
% 
% 
% 
% % Medium education
% 
% % \input{../figures/appendix_spell4_med.tex}
% 
% % PARITY PROGRESSION SURVIVAL - MEDIUM
% 
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_med_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_med_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_med_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with 1 to 7 years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell4_med_pps}
% \end{figure}
% 
% 
% 
% % High education
% 
% % \input{../figures/appendix_spell4_high.tex}
% 
% % PARITY PROGRESSION SURVIVAL - high
% 
% \begin{figure}[htpb]
% \centering
% \caption*{Urban}
% \setcounter{subfigure}{-1}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob.\ no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_urban_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{-0}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{1}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{2}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_high_urban_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption*{Rural}
% \setcounter{subfigure}{3}
% \subfloat[1972--1984]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g1_high_rural_pps}} 
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } 
% \setcounter{subfigure}{4}
% \subfloat[1985--1994]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g2_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% } \\
% \setcounter{subfigure}{5}
% \subfloat[1995--2004]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g3_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \setcounter{subfigure}{6}
% \subfloat[2005--2016]{
%     \begin{minipage}{0.31\textwidth}
%         \captionsetup[subfigure]{labelformat=empty,position=top,captionskip=-1pt,farskip=-0.5pt}
%         \subfloat[Prob. no birth yet]{\includegraphics[width=\textwidth]{spell4_g4_high_rural_pps}}
%         \captionsetup[subfigure]{labelformat=parens}
%     \end{minipage}
% }
% \caption{Survival curves conditional on parity progression
% for women with eight to eleven years of education by month beginning 9 months after prior birth.
% }
% \label{fig:results_spell4_high_pps}
% \end{figure}
% 
% 
% \clearpage
% \newpage
%

\section{Infant Mortality Graphs}

\setcounter{figure}{0}
\setcounter{table}{0}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_low_dummies} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_low_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_low_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_low_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with no education}
\label{fig:mortality_low}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_med_dummies} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_med_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_med_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_med_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with one to
seven years of education}
\label{fig:mortality_med}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1972-1984}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p1_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p1_high_dummies} 
    \end{minipage}
}
\\
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_high_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_high_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_high_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with eight
to eleven years of education}
\label{fig:mortality_high}
\end{figure}

\begin{figure}[htpb]
\centering
\rotatebox[origin=c]{90}{\small{1985-1994}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p2_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p2_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{1995-2004}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p3_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p3_highest_dummies} 
    \end{minipage}
} 
\\
\rotatebox[origin=c]{90}{\small{2005-2016}}
\subfloat[Second Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s2_p4_highest_dummies}
    \end{minipage}
} 
\subfloat[Third Spell]{
    \begin{minipage}{0.40\textwidth}
        \includegraphics[width=\textwidth]{mortality_s3_p4_highest_dummies} 
    \end{minipage}
} 
\caption{Predicted Probability of Infant Mortality for women with twelve or
more years of education}
\label{fig:mortality_highest}
\end{figure}

\end{document}