### 2021-05-04 final/tables/fertility.tex
Demography needs figures in PDF and separate paper and appendices

\begin{table}[hp!]
\begin{center}
\begin{footnotesize}
\begin{threeparttable}
\caption{Four-parity fertility rate versus predicted cohort fertility based on hazard model}
\label{tab:fertility}
\begin{tabular}{@{} l D{.}{.}{2.2} D{.}{.}{2.2} D{.}{.}{2.2} D{.}{.}{2.2} D{.}{.}{2.2}  @{}}
\toprule
                       &            \mct{NFHS--1}          & \mco{NFHS--2}   & \mco{NFHS--3}   & \mco{NFHS--4}   \\
Fertility Rate Period  & \mco{1987--1988}  & \mco{1992--1993}  & \mco{1998--1999}  & \mco{2005--2006}  & \mco{2015--2016}  \\
Hazard Model Period    & \mco{1972--1984}  &                 & \mco{1985--1994}  & \mco{1995--2004}  & \mco{2004--2016}  \\
\midrule
 & \multicolumn{5}{c}{Urban} \\ \cmidrule(lr){2-6}
 & \multicolumn{5}{c}{No Education} \\
Fertility Rate\tnote{a}   &      3.55       &      3.06       &      2.80       &      2.54       &      2.45       \\
Hazard Model\tnote{b}     &      3.44       &                 &      3.29       &      3.06       &      2.79       \\
\addlinespace 
 & \multicolumn{5}{c}{1--7 Years of Education} \\
Fertility Rate\tnote{a}   &      2.85       &      2.29       &      2.09       &      1.99       &      2.04       \\
Hazard Model\tnote{b}     &      3.18       &                 &      2.88       &      2.62       &      2.42       \\
\addlinespace 
 & \multicolumn{5}{c}{8--11 Years of Education} \\
Fertility Rate\tnote{a}   &      2.43       &      2.04       &      1.84       &      1.81       &      1.87       \\
Hazard Model\tnote{b}     &      2.72       &                 &      2.41       &      2.28       &      2.07       \\
\addlinespace 
 & \multicolumn{5}{c}{12 or More Years of Education} \\
Fertility Rate\tnote{a}   &      2.05       &      1.68       &      1.57       &      1.55       &      1.51       \\
Hazard Model\tnote{b}     &      2.29       &                 &      2.06       &      1.94       &      1.80       \\
\addlinespace 
 & \multicolumn{5}{c}{Rural} \\ \cmidrule(lr){2-6}
 & \multicolumn{5}{c}{No Education} \\
Fertility Rate\tnote{a}   &      3.57       &      2.93       &      2.63       &      2.74       &      2.81       \\
Hazard Model\tnote{b}     &      3.55       &                 &      3.38       &      3.26       &      3.09       \\
\addlinespace 
 & \multicolumn{5}{c}{1--7 Years of Education} \\
Fertility Rate\tnote{a}   &      3.01       &      2.52       &      2.39       &      2.25       &      2.37       \\
Hazard Model\tnote{b}     &      3.29       &                 &      3.08       &      2.83       &      2.70       \\
\addlinespace 
 & \multicolumn{5}{c}{8--11 Years of Education} \\
Fertility Rate\tnote{a}   &      2.56       &      2.21       &      2.22       &      2.16       &      2.19       \\
Hazard Model\tnote{b}     &      2.93       &                 &      2.68       &      2.49       &      2.31       \\
\addlinespace 
 & \multicolumn{5}{c}{12 or More Years of Education} \\
Fertility Rate\tnote{a}   &      1.95       &      1.68       &      2.13       &      2.08       &      1.96       \\
Hazard Model\tnote{b}     &      2.64       &                 &      2.39       &      2.25       &      2.11       \\
\addlinespace 
\bottomrule
\end{tabular}
\begin{tablenotes} \scriptsize
\item \hspace*{-0.5em} \textbf{Note.}
All predictions based on births up to and including parity four births
for both fertility rate and model predictions.
NFHS-1 was collected in 1992--1993, and model results for 1972--1984 were
applied for the predictions.
NFHS-2 was collected in 1998--1999, and model results for 1985--1994 were
applied for the predictions.
NFHS-3 was collected in 2005--2006, and model results for 1995--2004 were
applied for the predictions.
NFHS-4 was collected in 2015--2016, and model results for 2005--2016 were
applied for the predictions.
\item[a] 
The fertility rate is based on five-year age groups, counting births that 
occurred 1--36 months before the survey months.
For NFHS-1 and NFHS-2, the total number of women in the five-year age
groups is based on the household roster because only ever-married women
are in the individual recode sample.
For NFHS-3 and NFHS-4, the total number of women is based on the individual
recode sample because all women were interviewed.
\item[b] 
The model predictions for fertility are the average predicted fertility
across all women in a given sample, using their age of marriage as the
starting point and adding three years for each spell.
Observed births are not taken into account for the predictions.
For each spell, the predicted probability is the likelihood of having a
next birth given sex composition multiplied with the probability of that
sex composition and the likelihood of getting to the spell,
corrected for the probability of sterilization.
\end{tablenotes}
\end{threeparttable}
\end{footnotesize}
\end{center}
\end{table}